### 2014-05-21 Manuscript/dark.tex
Initial commit of draft manuscript

\documentclass[a4paper,12pt]{article}
\usepackage[osf]{mathpazo} % palatino
\usepackage{ms}     % load the template
\usepackage[round]{natbib} % author-year citations
\usepackage{graphicx}
\pagenumbering{arabic}

\title{The "dark-side" of phylogenetic comparative methods}
\author{
Richard G. FitzJohn$^{1}$, Matthew W. Pennell$^{2}$\\, some others and Natalie Cooper$^{3,*}$
}
\date{}
\affiliation{\noindent{\footnotesize
$^1$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia \\
$^2$ Institute for Bioinformatics and Evolutionary Studies, University
of Idaho, Moscow, ID 83844, U.S.A.\\
$^3$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
$^4$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
$^*$ ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland.\\ Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
}}

\vfill
%\paragraph{Word-count:} X words

\runninghead{The dark-side of PCMs}
\keywords{some words}

\begin{document}
%\pagestyle{empty}
\modulolinenumbers[1]   % Line numbering on every line

\mstitlepage
\parindent=1.5em
\addtolength{\parskip}{.3em}

\begin{abstract}
\raggedright
\singlespacing
An abstract abstract
\end{abstract}

\newpage
\raggedright
\doublespacing
\section{Introduction}

\section{Example of the problem}

\section{Future directions}

\section{Acknowledgments}

NC is supported by Marie Curie CIG grant (proposal no: )

%\bibliographystyle{systbiol}
%\bibliography{darkside.bib}

\end{document}

### 2014-05-22 Manuscript/dark.tex
Initial commit of bibliography files

@@
 
-\runninghead{The dark-side of PCMs}
+\runninghead{The dark-side of PCMs} % "Point of view" for Syst Biol
 \keywords{some words}
 
+%%%
+% From Syst Biol template
+\linespread{1.66}
+\raggedright
+\setlength{\parindent}{0.5in}
+
+\pagestyle{empty}
+
+\renewcommand{\section}[1]{%
+\bigskip
+\begin{center}
+\begin{Large}
+\normalfont\scshape #1
+\medskip
+\end{Large}
+\end{center}}
+
+\renewcommand{\subsection}[1]{%
+\bigskip
+\begin{center}
+\begin{large}
+\normalfont\itshape #1
+\end{large}
+\end{center}}
+
+\renewcommand{\subsubsection}[1]{%
+\vspace{2ex}
+\noindent
+\textit{#1.}---}
+
+\renewcommand{\tableofcontents}{}
+
+\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
+
+% Ignoring their title page setup as I like Rich's template better :)
+
+%%%%%
+
 \begin{document}
-%\pagestyle{empty}
 \modulolinenumbers[1]   % Line numbering on every line
@@
 
-\begin{abstract}
-\raggedright
-\singlespacing
-An abstract abstract
-\end{abstract}
+\section{abstract}
+An abstract abstract % If "Point of View" there's no abstract
 
@@
 \doublespacing
-\section{Introduction}
+%\section{Introduction} % No intro heading for Syst biol
 
 \section{Example of the problem}
+
+We extracted yearly citation counts from the Web of Science on 22nd May 2013 for the following papers: \citet{boettiger2012your,blomberg2012independent,diaz1996testing,felsenstein1988phylogenies,felsenstein1985phylogenies,freckleton2009seven,freckleton2000phylogenetic,freckleton2002phylogenetic,freckleton2006detecting,grafen1989phylogenetic,garland1992procedures,hansen1997stabilizing,hansen1996translating,hansen2005assessing,jones1997optimum,garland2000using,garland1992rate,martins1997phylogenies,nee1994extinction,price1997correlated,rohle2006comment,ricklefs1996applications,rohlf2001comparative,schluter1997likelihood,westoby1995misinterpreting,westoby1995further}.
+We also obtained citation counts for \citet{harvey1991comparative} using Google Scholar, as Web of Science does not provide citation counts for books.
 
@@
 \section{Acknowledgments}
+NC is supported by The European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant, proposal number: 321696.
 
-NC is supported by Marie Curie CIG grant (proposal no: )
-
-%\bibliographystyle{systbiol}
-%\bibliography{darkside.bib}
+\bibliographystyle{sysbio}
+\bibliography{darkside}
 

### 2014-05-22 Analyses/Figure_CitationsPerYear.R
First attempt at citations through time graph for all papers

@@ -1,0 +1,24 @@
+library(ggplot2)
+library(reshape)
+
+ds <- read.csv("CitesThruTime.csv")
+
+# Reshape the data so it's in long format
+data_long <- melt(ds, id = "Year")
+
+pdf("CitationsPerYear_AllPapers")
+
+# Graph showing how citations change through time
+# for key PCM papers
+myplot <- ggplot(data = data_long,
+                 aes(x = Year, y = value, colour = variable))
+    
+# Modifications to graph including legend, y-axis label etc.
+myplot + 
+geom_line() + 
+guides(col = guide_legend(nrow = 14, title = "Paper"), size = 0.5) +
+ylab("Citations")
+
+dev.off()
+
+# Need to limit to fewer papers to get a clearer graph

### 2014-05-25 Manuscript/dark.tex
Beginning of introduction draft

--- 2014-05-21 Manuscript/dark.tex
@@
 
-\title{The "dark-side" of phylogenetic comparative methods}
+\title{The ``Dark Side'' of Phylogenetic Comparative Methods}
 \author{
-Richard G. FitzJohn$^{1}$, Matthew W. Pennell$^{2}$\\, some others and Natalie Cooper$^{3,*}$
+Richard G. FitzJohn$^{1}$, Matthew W. Pennell$^{2}$, others,\\ and Natalie Cooper$^{3,4,*}$
 }
@@
 $^4$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-$^*$ ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland.\\ Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+$^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
 }}
@@
 
-\runninghead{The dark-side of PCMs}
-\keywords{some words}
+\runninghead{The "dark side" of PCMs} % "Point of view" for Syst Biol
+\keywords{PCM, assumption, etc.}
+
+%%%%%
+% From Syst Biol template
+\linespread{1.66}
+\raggedright
+\setlength{\parindent}{0.5in}
+
+\pagestyle{empty}
+
+\renewcommand{\section}[1]{%
+\bigskip
+\begin{center}
+\begin{Large}
+\normalfont\scshape #1
+\medskip
+\end{Large}
+\end{center}}
+
+\renewcommand{\subsection}[1]{%
+\bigskip
+\begin{center}
+\begin{large}
+\normalfont\itshape #1
+\end{large}
+\end{center}}
+
+\renewcommand{\subsubsection}[1]{%
+\vspace{2ex}
+\noindent
+\textit{#1.}---}
+
+\renewcommand{\tableofcontents}{}
+
+\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
+
+% Ignoring their title page setup as I like Rich's template better :)
+
+%%%%%
 
 \begin{document}
-%\pagestyle{empty}
 \modulolinenumbers[1]   % Line numbering on every line
@@
 
-\begin{abstract}
-\raggedright
-\singlespacing
-An abstract abstract
-\end{abstract}
+%\section{abstract}
+% If Syst Biol "Point of View" there's no abstract
 
@@
 \doublespacing
-\section{Introduction}
+%\section{Introduction} % No intro heading for Syst Biol
 
-\section{Example of the problem}
+Phylogenetic comparative methods (PCMs) were initially developed in the 1980s to deal with the statistical non-independence of species in comparative analyses.
+Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, rates of trait evolution, and modes of evolution for morphological and ecological traits %list of citations or see Pennell & Harmon or O'Meara).
 
-\section{Future directions}
+PCMs have become extremely popular over recent years; in 2013 alone \citet{harvey1991comparative}'s seminal book on the comparative method in evolutionary biology received 195 citations (Google Scholar search, 22nd May 2014), and at the 2014 Evolution meeting in Raleigh NC, we estimate that over X talks used PCMs.
+This increase in use is at least partially the result of the number of available R packages to run these analyses (for example ape, GEIGER, diversitree, caper - citations) and the community's increasing familiarity with R \citep{R-Core-Team:2014aa}.
+Unfortunately with this increase in popularity, comes an increase in misuse!
+% need to work on this paragraph
+
+% From proposal: PCMs make numerous assumptions and suffer from biases in the same way as any statistical method, but these issues are often inadequately assessed in empirical studies.
+Such issues are the responsibility of end users but also of methods developers: the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
+To address these issues, we propose this symposium to discuss issues in both classical and recent PCMs, along with new research on detecting for these issues and accounting for them.
+We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
+
+\subsection{Example of the problem}
+
+We extracted yearly citation counts from the Web of Science on 22nd May 2013 for the following papers: \citet{diaz1996testing,felsenstein1988phylogenies,felsenstein1985phylogenies,freckleton2000phylogenetic,freckleton2002phylogenetic,freckleton2006detecting,grafen1989phylogenetic,garland1992procedures,hansen1996translating,hansen2005assessing,jones1997optimum,garland2000using,garland1992rate,martins1997phylogenies,price1997correlated,rohle2006comment,ricklefs1996applications,rohlf2001comparative,schluter1997likelihood,westoby1995misinterpreting,westoby1995further}.
+We also obtained citation counts for \citet{harvey1991comparative} using Google Scholar, as Web of Science does not provide citation counts for books.
+
+\subsection{Recommendations}
+
+\section{Funding}
+This work was supported by The European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant, proposal number: 321696 (NC)
 
 \section{Acknowledgments}
+Thanks to Star Wars for the hilarious jokes.
 
-NC is supported by Marie Curie CIG grant (proposal no: )
-
-%\bibliographystyle{systbiol}
-%\bibliography{darkside.bib}
+\bibliographystyle{sysbio}
+\bibliography{darkside}
 

### 2014-05-26 Analyses/Figure_PackagesThruTime.R
Initial commit for figure of cumulative packages through time

@@ -1,0 +1,33 @@
+library(ggplot2)
+library(reshape)
+
+ds <- read.csv("PackagesThruTime.csv")
+
+# Group packages by year using aggregate then fix names
+dsYear <- with(ds, aggregate(Program, by = list(YearReleased, RPackage), 
+	                          FUN = length))
+
+names(dsYear) <- c("YearReleased", "R", "NPackages")
+
+# Order the data by year so cumulative number of packages
+# per year can be calculated
+dsYear <- dsYear[order(dsYear$YearReleased), ]
+
+# Calculate cumulative number of packages per year
+dsYear$TotalPkg <- cumsum(dsYear$NPackages)
+dsR$TotalPkg <- cumsum(dsR$NPackages)
+dsX$TotalPkg <- cumsum(dsX$NPackages)
+
+# pdf("PackagesPerYear.pdf")
+
+# Plot cumulative number of packages through time
+myplot <- ggplot(data = dsYear,
+                 aes(x = YearReleased, y = TotalPkg, color = R))
+    
+# Modifications to graph including legend, y-axis label etc.
+myplot + 
+geom_smooth(method = "loess", se = FALSE) + 
+ylab("Total number of packages") +
+xlab("Year released")
+
+# dev.off()

### 2014-05-26 Analyses/Figure_PackagesThruTime.R
Modified to separate out R and non R packages

@@ -14,9 +14,9 @@ names(dsYear) <- c("YearReleased", "R", "NPackages")
 dsYear <- dsYear[order(dsYear$YearReleased), ]
 
 # Calculate cumulative number of packages per year
-dsYear$TotalPkg <- cumsum(dsYear$NPackages)
-dsR$TotalPkg <- cumsum(dsR$NPackages)
-dsX$TotalPkg <- cumsum(dsX$NPackages)
+# separately for non R and R packages
+dsYear$TotalPkg[dsYear$R == "other"] <- cumsum(dsYear$NPackages[dsYear$R == "other"])
+dsYear$TotalPkg[dsYear$R == "R"] <- cumsum(dsYear$NPackages[dsYear$R == "R"])
 
 # pdf("PackagesPerYear.pdf")
 
@@ -27,7 +27,7 @@ myplot <- ggplot(data = dsYear,
 # Modifications to graph including legend, y-axis label etc.
 myplot + 
 geom_smooth(method = "loess", se = FALSE) + 
-ylab("Total number of packages") +
+ylab("Cumulative number of packages") +
 xlab("Year released")
 
 # dev.off()

### 2014-05-29 Manuscript/dark.tex
Finished introduction draft

@@
 
-Phylogenetic comparative methods (PCMs) were initially developed in the 1980s to deal with the statistical non-independence of species in comparative analyses.
-Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, rates of trait evolution, and modes of evolution for morphological and ecological traits %list of citations or see Pennell & Harmon or O'Meara).
+Phylogenetic comparative methods (PCMs) were initially developed in the 1980s to deal with the statistical non-independence of species in comparative analyses (e.g.
+\citet{felsenstein1985phylogenies,grafen1989phylogenetic}.
+Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary} and \citealp{pennell2013integrative}).
 
-PCMs have become extremely popular over recent years; in 2013 alone \citet{harvey1991comparative}'s seminal book on the comparative method in evolutionary biology received 195 citations (Google Scholar search, 22nd May 2014), and at the 2014 Evolution meeting in Raleigh NC, we estimate that over X talks used PCMs.
-This increase in use is at least partially the result of the number of available R packages to run these analyses (for example ape, GEIGER, diversitree, caper - citations) and the community's increasing familiarity with R \citep{R-Core-Team:2014aa}.
+PCMs have become extremely popular over recent years; in 2013 alone \citet{harvey1991comparative}'s book on the comparative method in evolutionary biology received 195 citations (Google Scholar, 22nd May 2014), and at the 2014 Evolution meeting in Raleigh NC, we estimate that over XX %need to count these!
+talks used PCMs of one form or another.
+Unfortunately, PCMs make numerous assumptions and suffer from biases in the same way as any statistical method, but the more popular these methods become, the less awareness methods users seem to have of this ``dark side''.
+Thus assumptions and biases are often inadequately assessed in empirical studies, leading to poor model fits and mis- or over-interpreted results.
+
+%Here we explore this problem in more detail.
+
+%Such issues are the responsibility of end users but also of methods developers: the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
+To address these issues, we propose this symposium to discuss issues in both classical and recent PCMs, along with new research on detecting for these issues and accounting for them.
+We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
+
+%In this paper we discuss why we think this has happened, and suggest ways to bridge the knowledge gap between methods users and methods developers.
+
+%This increase in use is at least partially the result of the number of available R packages to run these analyses (for example ape, GEIGER, diversitree, caper - citations) and the community's increasing familiarity with R \citep{R-Core-Team:2014aa}.
 Unfortunately with this increase in popularity, comes an increase in misuse!
@@
 We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
+
+%Why are PCMs increasingly being misused? All evidence points to a knowledge-gap between methods users and methods developers.
+
+\subsection{}
+
+To apply a method correctly, methods users must first be aware of the assumptions and biases of the method in question.
+
+- and how to detect whether their model adequately meets these assumptions - as well as details of the number of taxa required for sufficient power etc.
+If failure to account for the assumptions of the method is the result of a knowledge gap then there are several possible explanations:
+
+\begin{enumerate}
+\item The assumptions are not mentioned in the literature.
+\item The assumptions are mentioned in the literature, but the literature is too dense and technical, or pertinent details are hidden within large sections of text.
+\item The assumptions are mentioned in the literature but methods users have gone straight to the implementation of the method and the manual does not mention assumptions
+
+Above we have assumed a knowledge gap, we are not considering cases where The method user is aware of the assumptions and how to test for them, but chooses not to because it means they cannot publish.
 
@@
 \section{Funding}
-This work was supported by The European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant, proposal number: 321696 (NC)
+This work was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696 (NC)
 
 \section{Acknowledgments}
-Thanks to Star Wars for the hilarious jokes.
+Thanks to Star Wars for the ``hilarious'' jokes.
 

### 2014-05-29 Manuscript/dark.tex
Edits to intro and first section

@@
 Phylogenetic comparative methods (PCMs) were initially developed in the 1980s to deal with the statistical non-independence of species in comparative analyses (e.g.
-\citet{felsenstein1985phylogenies,grafen1989phylogenetic}.
-Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary} and \citealp{pennell2013integrative}).
+\citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
+Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
+PCMs have become extremely popular over recent years; in 2013 alone Harvey and Pagel's \citeyearpar{harvey1991comparative} book on \textit{The Comparative Method in Evolutionary Biology} was cited 195 times (Google Scholar, 22nd May 2014), and at the 2014 Evolution meeting in Raleigh NC, we estimate that over XX %need to count these!
+talks used PCMs of one form or another.
 
-PCMs have become extremely popular over recent years; in 2013 alone \citet{harvey1991comparative}'s book on the comparative method in evolutionary biology received 195 citations (Google Scholar, 22nd May 2014), and at the 2014 Evolution meeting in Raleigh NC, we estimate that over XX %need to count these!
-talks used PCMs of one form or another.
-Unfortunately, PCMs make numerous assumptions and suffer from biases in the same way as any statistical method, but the more popular these methods become, the less awareness methods users seem to have of this ``dark side''.
-Thus assumptions and biases are often inadequately assessed in empirical studies, leading to poor model fits and mis- or over-interpreted results.
+PCMs also have a ``dark side''; they make various assumptions and suffer from biases in the same way as any statistical method.
+Unfortunately, the more popular these methods become, the less awareness methods users seem to have of this ``dark side''.
+Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and mis- or over-interpreted results.
+Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers, but who should take responsibility for such actions?
+And why has this gap developed in the first place?
+Here we explore these issues and make suggestions for making everything better.
+%needs some tightening!!!
 
-%Here we explore this problem in more detail.
+\subsection{Why aren't more methods-users aware of the ``dark side'' of PCMs?}
 
-%Such issues are the responsibility of end users but also of methods developers: the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
-To address these issues, we propose this symposium to discuss issues in both classical and recent PCMs, along with new research on detecting for these issues and accounting for them.
-We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
+To apply a method correctly, methods-users must be aware of the assumptions and biases of the method in question, and know how to identify whether their model adequately meets these assumptions.
+If methods users are unaware of these factors there are several possible explanations:
 
-%In this paper we discuss why we think this has happened, and suggest ways to bridge the knowledge gap between methods users and methods developers.
+\begin{enumerate}
+\item Assumptions and biases are not mentioned in the literature.
+\item Assumptions and biases are mentioned in the literature, but the literature is too technical for the novice user and/or pertinent details are hidden within the text.
+\item Methods users have bypassed the literature gone straight to the implementation of the method.
+\end{enumerate}
+
+Note that we assume that methods-users ignore assumptions and biases because they are unaware of them.
+Of course, there may be methods-users who are perfectly aware of these issues but choose to sweep them under the carpet to get their work published.
+Although we believe there is no excuse for doing this, we also recognize that the ``publish or perish'' culture of academia may pressurize nontenured scientists into this.
+Hopefully some of the changes we suggest below will also help with this problem.
+
+\subsubsection{1. Assumptions and biases are not mentioned in the literature}
+
+\subsubsection{2. Literature is too technical or dense}
+
+\subsubsection{3. Users jump straight to the implementation of the method}
 
@@
 % need to work on this paragraph
-
-% From proposal: PCMs make numerous assumptions and suffer from biases in the same way as any statistical method, but these issues are often inadequately assessed in empirical studies.
-Such issues are the responsibility of end users but also of methods developers: the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
-To address these issues, we propose this symposium to discuss issues in both classical and recent PCMs, along with new research on detecting for these issues and accounting for them.
-We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
-
-%Why are PCMs increasingly being misused? All evidence points to a knowledge-gap between methods users and methods developers.
-
-\subsection{}
-
-To apply a method correctly, methods users must first be aware of the assumptions and biases of the method in question.
-
-- and how to detect whether their model adequately meets these assumptions - as well as details of the number of taxa required for sufficient power etc.
-If failure to account for the assumptions of the method is the result of a knowledge gap then there are several possible explanations:
-
-\begin{enumerate}
-\item The assumptions are not mentioned in the literature.
-\item The assumptions are mentioned in the literature, but the literature is too dense and technical, or pertinent details are hidden within large sections of text.
-\item The assumptions are mentioned in the literature but methods users have gone straight to the implementation of the method and the manual does not mention assumptions
-
-Above we have assumed a knowledge gap, we are not considering cases where The method user is aware of the assumptions and how to test for them, but chooses not to because it means they cannot publish.
 
@@
 \subsection{Recommendations}
+
+%Such issues are the responsibility of end users but also of methods developers: the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
+To address these issues, we propose this symposium to discuss issues in both classical and recent PCMs, along with new research on detecting for these issues and accounting for them.
+We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
 

### 2014-05-29 Manuscript/dark.tex
Added notes for completion of other sections

@@
 PCMs also have a ``dark side''; they make various assumptions and suffer from biases in the same way as any statistical method.
-Unfortunately, the more popular these methods become, the less awareness methods users seem to have of this ``dark side''.
-Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and mis- or over-interpreted results.
+Unfortunately, the more popular these methods become, the less awareness methods-users seem to have of this ``dark side''.
+Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results.
 Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers, but who should take responsibility for such actions?
 And why has this gap developed in the first place?
-Here we explore these issues and make suggestions for making everything better.
-%needs some tightening!!!
+Here we explore these issues and make suggestions for improvements.\\ %needs some rewording
 
@@
 To apply a method correctly, methods-users must be aware of the assumptions and biases of the method in question, and know how to identify whether their model adequately meets these assumptions.
-If methods users are unaware of these factors there are several possible explanations:
+If methods-users are unaware of these factors there are several possible explanations:
 
@@
 
+%Distinction between "warning" papers, and papers that just mention assumptions in passing.
+
+%Yes they are!
+Pictures of warnings papers and some highlighted text from Harvey and Pagel.
+Seems to be disconnect between users and literature, esp older stuff – graph for Felsenstein 85 showing increased citations but without caveats papers cited.
+Caveats papers often cited as but(t) see or fly by citations with no depth.
+
+%Note that there is a time lag though and a settling in period for a new method (show example with PICs: 1991 before real critiques appeared)
+
+%Some stuff isn’t published, or is published and isn’t justified - “folklore” and rules of thumb (±3 from Jones&Purvis97).
+
+%Notes on how citation counts were extracted:
+%We extracted yearly citation counts from the Web of Science on 22nd May 2013 for the following papers: We also obtained citation counts for \citet{harvey1991comparative} using Google Scholar, as Web of Science does not provide citation counts for books.
+
 \subsubsection{2. Literature is too technical or dense}
+
+%OK maybe a fair point.
+Can we expect methods users to read every paper before using a method?
+(Yes – wouldn’t let someone work on your microscope unless you were sure they knew what they were doing!!!).
+Perhaps some work to be done in summarizing this – needs incentives – no benefits to wikis and workshop teaching cf to writing high impact papers, or papers at all.
+However, people willing to share skills are the people we really want as colleagues, not the selfish high achievers!
+Collaborate with someone who does know this stuff?
+Again problematic as highly skilled technical people end up with too much to do and still end up only being middle author.
+Maybe something like in physics with “theoretical lead” and “technical lead” authors???
+Perhaps write a less technical summary for methods papers, or a “Caveats box” to put this up front – i.e.
+will only work if you have X taxa and this kind of data.
+Check these assumptions.
+Incentivize accompanying blogs to explain this stuff?
+Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
+Speed up publication by archiving.
+Put code on github so knowledgeable users can access at an earlier stage if they want to.
+Helpful for finding bugs etc.
+Methods developers could also benefit from collaboration with methods users who could help them find corner cases and make their methods more useful with real data?
+[Again I’m torn with some of these ideas because is it really the developer’s job to stop people from being fuckwits?
+But if this info isn’t included here, *where* would it come from and *who* would provide it?
+Users can’t be trusted.]
 
@@
 
-\subsection{Example of the problem}
+%Huge increase in number of PCM packages since 1980s, and especially since 2008 [Graph].
+Now people familiar with R they can just jump in and try stuff on their data with very little knowledge of what is going on or what the results mean [Though code sharing is cool and I’m not dissing that].
 
-We extracted yearly citation counts from the Web of Science on 22nd May 2013 for the following papers: \citet{diaz1996testing,felsenstein1988phylogenies,felsenstein1985phylogenies,freckleton2000phylogenetic,freckleton2002phylogenetic,freckleton2006detecting,grafen1989phylogenetic,garland1992procedures,hansen1996translating,hansen2005assessing,jones1997optimum,garland2000using,garland1992rate,martins1997phylogenies,price1997correlated,rohle2006comment,ricklefs1996applications,rohlf2001comparative,schluter1997likelihood,westoby1995misinterpreting,westoby1995further}.
-We also obtained citation counts for \citet{harvey1991comparative} using Google Scholar, as Web of Science does not provide citation counts for books.
+%I’m torn as to what I think here – part of me thinks there should be a mention of assumptions in the manual/vignette (though not if this replicates an accompanying blog or book or something) but part of me thinks it’s the users responsibility to not be a fuckwit.
+Caper, diversitree and CAIC are good examples where assumptions are clearly laid out and methods for testing assumptions are available in caper and CAIC (not sure about diversitree ...).
+Ape book has none of this in it at all!
+
+%Again this needs to be incentivized as it’s adding work for already hard working and underappreciated methods developers.
+Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
 
@@
 We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
+
+%Big issue = who is responsible? Can't expect empiricists to be an expert in every method, but do want to add to workload of overloaded developers.
+
+%Improvements in methods, and ways of detecting biases etc should be sufficient for publication (rather than MEE's weird focus on novelty now it has a huge IF)
+
+%Incentives - funding for pure methods. Different authorship rules.
+
+%Change the definition of success in academia. Quality not quantity. Remove rush to publish. This would help WOMEN too!
+
+%Collaborate! But make collaboration valuable to non first authors somehow
+
+%We really *do not* want to lose all the technical people from science.
+We need them more than ever as evo bio becomes more and more computationally intensive.
+Need to make it worth their while (can't match wages of industry)
+
+%Add methods to test assumptions as standard to packages.
+
+%Grouping assumptions, biases and caveats more carefully in papers. Caveats box???
+
+%Accompanying blogs with methods papers to explain to non-technical audience. Could be Supp Matt too. Lavin et al 2008 is a beautiful example.
+
+%Some website with a list of key papers, and key summary points, for each method. Who would host? Who would write? Who would peer review?
+
+%Users need to take responsibility too.
+Work harder to understand.
+Read widely.
+Don't take results at face value.
+Become less trusting!
+Simulate data to check method does what you think it should do.
+Don't be afraid to question standard practice - sometimes it's just folklore.
 

### 2014-05-30 Manuscript/dark.tex
Added reminder to cite other similar papers
Reworking introduction

@@
 
-Phylogenetic comparative methods (PCMs) were initially developed in the 1980s to deal with the statistical non-independence of species in comparative analyses (e.g.
+Phylogenetic comparative methods (PCMs) were initially developed a long time ago in a galaxy far, far away - also know as the 1980s - to deal with the statistical non-independence of species in comparative analyses (e.g.
 \citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
@@
 PCMs have become extremely popular over recent years; in 2013 alone Harvey and Pagel's \citeyearpar{harvey1991comparative} book on \textit{The Comparative Method in Evolutionary Biology} was cited 195 times (Google Scholar, 22nd May 2014), and at the 2014 Evolution meeting in Raleigh NC, we estimate that over XX %need to count these!
-talks used PCMs of one form or another.
+talks used PCMs of one form or another. With advances in blah balh, there has never been a better time to be a comparative biologist.
 
-PCMs also have a ``dark side''; they make various assumptions and suffer from biases in the same way as any statistical method.
-Unfortunately, the more popular these methods become, the less awareness methods-users seem to have of this ``dark side''.
+However, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in the same way as any statistical method \citet{freckleton2009seven} Unfortunately, the more popular these methods become, the less awareness methods-users seem to have of this ``dark side''.
 Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results.
+Additionally, little consideration is given to whether using a PCM is really appropriate for the question \citet{losos2011seeing}.
+Although methods-developers (and advanced methods-users) are aware of these issues, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
 Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers, but who should take responsibility for such actions?
@@
 Here we explore these issues and make suggestions for improvements.\\ %needs some rewording
+
+%some comments showing we're not the only people to think this, nor the first to say it.
+%cite Boettiger etal 2012
+%cite Losos 2011
+%cite Freckleton 2009
+%cite Blomberg et al 2012
 

### 2014-08-21 Manuscript/dark.tex
Reordered introduction

--- 2014-05-29 Manuscript/dark.tex
@@
+% Preamble
 \documentclass[a4paper,12pt]{article}
+
 \usepackage[osf]{mathpazo} % palatino
@@
 
-\title{The ``Dark Side'' of Phylogenetic Comparative Methods}
-\author{
-Richard G. FitzJohn$^{1}$, Matthew W. Pennell$^{2}$, others,\\ and Natalie Cooper$^{3,4,*}$
-}
-\date{}
-\affiliation{\noindent{\footnotesize
-$^1$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia \\
-$^2$ Institute for Bioinformatics and Evolutionary Studies, University
-of Idaho, Moscow, ID 83844, U.S.A.\\
-$^3$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
-$^4$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-$^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
-}}
-
-\vfill
-%\paragraph{Word-count:} X words
-
-\runninghead{The "dark side" of PCMs} % "Point of view" for Syst Biol
-\keywords{PCM, assumption, etc.}
-
-%%%%%
 % From Syst Biol template
@@
 
-\renewcommand{\section}[1]{%
+\renewcommand{\section}[1]{
   \bigskip
@@
   \end{Large}
-\end{center}}
+  \end{center}
+}
 
-\renewcommand{\subsection}[1]{%
+\renewcommand{\subsection}[1]{
   \bigskip
@@
   \end{large}
-\end{center}}
+  \end{center}
+}
 
@@
 
-%%%%%
+% Title page information
+
+\title{The ``Dark Side'' of Phylogenetic Comparative Methods}
+\author{
+  Richard G. FitzJohn$^{1}$, Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\ Matthew W. Pennell$^{4}$, Gavin H. Thomas$^{5}$, April Wright$^{6}$,\\ and Natalie Cooper$^{7,8,*}$
+}
+\date{}
+\affiliation{\noindent{\footnotesize
+$^1$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
+$^2$ Wisconsin\\
+$^3$ UBC\\
+$^4$ Institute for Bioinformatics and Evolutionary Studies, University
+of Idaho, Moscow, ID 83844, U.S.A.\\
+$^5$ Sheffield\\
+$^6$ Austin\\
+$^7$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
+$^8$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
+$^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+}}
+
+\vfill
+%\paragraph{Word-count:} X words
+
+\runninghead{The ``Dark Side'' of PCMs}
+\keywords{PCM, assumption, etc.}
+
+% End of preamble
 
@@
 \mstitlepage
-\parindent=1.5em
+\parindent = 1.5em
 \addtolength{\parskip}{.3em}
@@
 
-Phylogenetic comparative methods (PCMs) were initially developed in the 1980s to deal with the statistical non-independence of species in comparative analyses (e.g.
+A long time ago in a galaxy far, far away - also known as the 1980s - phylogenetic comparative methods (PCMs) were developed to deal with the statistical non-independence of species in comparative analyses (e.g.
 \citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
-PCMs have become extremely popular over recent years; in 2013 alone Harvey and Pagel's \citeyearpar{harvey1991comparative} book on \textit{The Comparative Method in Evolutionary Biology} was cited 195 times (Google Scholar, 22nd May 2014), and at the 2014 Evolution meeting in Raleigh NC, we estimate that over XX %need to count these!
-talks used PCMs of one form or another.
+PCMs have become extremely popular over recent years; the number of papers featured the phrase `phylogenetic comparative' has increased exponentially since the 1980s (Figure \ref{PCMCitations}). 
+With new methods being published almost weekly, there has never been a better time to be a comparative biologist.
 
-PCMs also have a ``dark side''; they make various assumptions and suffer from biases in the same way as any statistical method.
-Unfortunately, the more popular these methods become, the less awareness methods-users seem to have of this ``dark side''.
+\begin{figure}[h]
+\centering
+\includegraphics[width = 8cm]{PCMCitations.png}
+\caption{PCM papers through time}
+\label{PCMCitations}
+\end{figure}
+
+Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any statistical method \citep{freckleton2009seven,boettiger2012your}. 
+The more popular these methods become, the less awareness methods-users seem to have of this dark side.
 Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results.
+Additionally, little consideration is given to whether using a PCM is really appropriate for the question at hand \citep{losos2011seeing}. 
+To paraphrase \citet{blomberg2012independent}, the enthusiastic application of PCMs has greatly outpaced our theoretical understanding of the methods and their relations to the rest of statistical theory.
+
+% Do we need more explicit comments showing we're not the only people to think this, nor the first to say it?
+
+One major issue is that two groups of people are involved with PCMs. 
+At one extreme, methods-developers (wookies) are involved with developing new methods and implementing them. They are aware of the limitations of their methods, and the assumptions that underly them. 
+% Bit of a stretch but we can pretend right? :)
+At the other extreme, naive methods-users (ewoks) have no knowledge of these complexities, they just want to use PCMs on their data. 
+They may be graduate students or more established scientists who have been asked to apply a phylogenetic correction by a reviewer. 
+Of course the majority of PCM users fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use.
+Although methods-developers (and advanced methods-users) are aware of the problems with PCMs, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
 Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers, but who should take responsibility for such actions?
@@
 
-\subsection{Why aren't more methods-users aware of the ``dark side'' of PCMs?}
+\section{Why aren't more methods-users aware of the ``dark side'' of PCMs?}
 
@@
 \begin{enumerate}
-\item Assumptions and biases are not mentioned in the literature.
-\item Assumptions and biases are mentioned in the literature, but the literature is too technical for the novice user and/or pertinent details are hidden within the text.
+\item Not everything is mentioned in the literature.
+\item The literature is too technical for the novice user and/or pertinent details are hidden within the text.
 \item Methods users have bypassed the literature gone straight to the implementation of the method.
@@
 
-Note that we assume that methods-users ignore assumptions and biases because they are unaware of them.
+%Note that we assume that methods-users ignore assumptions and biases because they are unaware of them.
 Of course, there may be methods-users who are perfectly aware of these issues but choose to sweep them under the carpet to get their work published.
@@
 
-\subsubsection{1. Assumptions and biases are not mentioned in the literature}
+\subsection{1. Not everything is mentioned in the literature}
+
+Program defaults are another thing which become folklore although tey are often arbitrary.
 
@@
 
-\subsubsection{2. Literature is too technical or dense}
+\subsubsection{2. Literature is technical, dense and numerous}
 
@@
 
-\subsubsection{3. Users jump straight to the implementation of the method}
+\subsubsection{3. Important information is hard to locate}
+
+\subsubsection{4. Users jump straight to the implementation of the method}
 
@@
 Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
+
+\subsection{An example using contrasts}
+Probably the most well-known, PCM is Felsenstein's \citeyearpar{felsenstein1985phylogenies} phylogenetically independent contrasts (PIC).
+shortly afterwards Grafen introduced the phylogenetic regression also known as phylogenetic generalised least squares (PGLS).
+Later papers confirmed that these methods are actually equivalent if a model of Brownian motion evolution is assumed (although the mathematical proof for this is "awkward"; \citet{rohlf2001comparative,blomberg2012independent})- with PIC being a special case of PGLS \citet{blomberg2012independent,rohle2006comment,rohlf2001comparative,garland2000using}.
+
+Many papers have discussed the assumptions of this method at length.\citealp{diaz1996testing} list these assumptions as 1) topology is perfect; 2) branch lengths are perfect and 3) the model of evolution is Brownian.
+Felsenstein points out the assumption of Brownian motion may not be very useful in his original paper \citet[p.13]{felsenstein1985phylogenies}.
+This was followed up by numerous publications in the 1990s discussing this assumption and providing multiple tests to see if the assumption was correct.
+Many of these papers are extremely explicit about the need to investigate these assumptions.
+
+Assumptions of tree accuracy are less commonly discussed, partly because evidence suggests they make very little difference. Polytomies are dealt with though.
 
@@
 
+%A New Hope joke...
+
 \section{Funding}
-This work was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696 (NC)
+This work was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696 (NC).
 
 \section{Acknowledgments}
-Thanks to Star Wars for the ``hilarious'' jokes.
+Thanks to the Society of Systematic Biologists for funding our symposium at Evolution 2014. We also thank George Lucas for the ``hilarious'' Star Wars jokes.
 

### 2014-08-21 Manuscript/dark.tex
Added section on phylogenetic folklore

@@
 
-\section{Why aren't more methods-users aware of the ``dark side'' of PCMs?}
+\section{Why don't all methods-users know about the ``Dark Side'' of PCMs?}
 
@@
 
-Program defaults are another thing which become folklore although tey are often arbitrary.
+A lot of the information needed to properly apply PCMs is not found in the literature. 
+We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students and over time these myths and rules-of-thumb become set in stone. 
+One example of such folklore is in the defaults of programs that perform PCMs. 
+These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.
 
-%Distinction between "warning" papers, and papers that just mention assumptions in passing.
+% Do we want a specific example here? Theta estimation in OU models? 
+% Could mention ±3 from Jones&Purvis97 - totally unjustified rule of thumb now = canon.
 
-%Yes they are!
-Pictures of warnings papers and some highlighted text from Harvey and Pagel.
-Seems to be disconnect between users and literature, esp older stuff – graph for Felsenstein 85 showing increased citations but without caveats papers cited.
-Caveats papers often cited as but(t) see or fly by citations with no depth.
+Other information about the limitations of a method may be absent from the literature due to the time lag between a new method being published and others having time to test and critique it. 
+For example, \citealp{felsenstein1985phylogenies} published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published %add citations.
+This time lag is shorter with more recent methods because more simulations testing the method are now required by journals.
 
-%Note that there is a time lag though and a settling in period for a new method (show example with PICs: 1991 before real critiques appeared)
-
-%Some stuff isn’t published, or is published and isn’t justified - “folklore” and rules of thumb (±3 from Jones&Purvis97).
-
-%Notes on how citation counts were extracted:
-%We extracted yearly citation counts from the Web of Science on 22nd May 2013 for the following papers: We also obtained citation counts for \citet{harvey1991comparative} using Google Scholar, as Web of Science does not provide citation counts for books.
-
-\subsubsection{2. Literature is technical, dense and numerous}
+\subsection{2. Literature is technical, dense and numerous}
 
@@
 
-\subsubsection{3. Important information is hard to locate}
+%Distinction between "warning" papers, and papers that just mention assumptions in passing.
 
-\subsubsection{4. Users jump straight to the implementation of the method}
+%Yes they are!
+Pictures of warnings papers and some highlighted text from Harvey and Pagel.
+Seems to be disconnect between users and literature, esp older stuff – graph for Felsenstein 85 showing increased citations but without caveats papers cited.
+Caveats papers often cited as but(t) see or fly by citations with no depth.
+
+%Note that there is a time lag though and a settling in period for a new method (show example with PICs: 1991 before real critiques appeared)
+
+%Some stuff isn’t published, or is published and isn’t justified - “folklore” and rules of thumb (
+
+\subsection{3. Important information is hard to locate}
+
+\subsection{4. Users jump straight to the implementation of the method}
 
@@
 
-\subsection{Recommendations}
+\section{How can we solve some of these problems?}
+
+\subsection{1. Simplify, summarize and share}
+\subsection{2. Read, think and learn}
+\subsection{3. Collaboration}
+\subsection{4. Incentives}
 
@@
 Don't be afraid to question standard practice - sometimes it's just folklore.
+My pglm vs pgls example.
 

### 2014-08-21 Manuscript/dark.tex
Added section on literature technicality

@@
 \section{Why don't all methods-users know about the ``Dark Side'' of PCMs?}
+% Or something related to the break down in communication
 
@@
 \item Not everything is mentioned in the literature.
-\item The literature is too technical for the novice user and/or pertinent details are hidden within the text.
+\item The literature is too technical for the novice user and/or important details are hidden within the text.
 \item Methods users have bypassed the literature gone straight to the implementation of the method.
@@
 Hopefully some of the changes we suggest below will also help with this problem.
+%This could be a footnote, or left out!
 
@@
 % Could mention ±3 from Jones&Purvis97 - totally unjustified rule of thumb now = canon.
+% Mention hidden assumptions?
 
@@
 
-\subsection{2. Literature is technical, dense and numerous}
+\subsection{2. The literature is technical, dense and important information is hidden}
 
-%OK maybe a fair point.
-Can we expect methods users to read every paper before using a method?
-(Yes – wouldn’t let someone work on your microscope unless you were sure they knew what they were doing!!!).
-Perhaps some work to be done in summarizing this – needs incentives – no benefits to wikis and workshop teaching cf to writing high impact papers, or papers at all.
-However, people willing to share skills are the people we really want as colleagues, not the selfish high achievers!
-Collaborate with someone who does know this stuff?
-Again problematic as highly skilled technical people end up with too much to do and still end up only being middle author.
-Maybe something like in physics with “theoretical lead” and “technical lead” authors???
-Perhaps write a less technical summary for methods papers, or a “Caveats box” to put this up front – i.e.
-will only work if you have X taxa and this kind of data.
-Check these assumptions.
-Incentivize accompanying blogs to explain this stuff?
-Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
-Speed up publication by archiving.
-Put code on github so knowledgeable users can access at an earlier stage if they want to.
-Helpful for finding bugs etc.
-Methods developers could also benefit from collaboration with methods users who could help them find corner cases and make their methods more useful with real data?
-[Again I’m torn with some of these ideas because is it really the developer’s job to stop people from being fuckwits?
-But if this info isn’t included here, *where* would it come from and *who* would provide it?
-Users can’t be trusted.]
+Although a lot of information is not found in the literature (see above), the majority of assumptions and biases are documented somewhere. 
+A big issue for novice methods-users is that this information is often extremely technical and dense. 
+It is not unusual for papers to be full of equations, over 20 pages long, and written in a way that gives even advanced users a headache. 
+Additionally there is a lot of literature to wade through before getting all of the required information. 
+For example, one of us (NC) reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies}. 
+Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book.
+
+Another issue is that it can sometimes be hard to find the information required.
+Assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper.
+They are rarely neatly corralled.
 
@@
 
-%Note that there is a time lag though and a settling in period for a new method (show example with PICs: 1991 before real critiques appeared)
-
-%Some stuff isn’t published, or is published and isn’t justified - “folklore” and rules of thumb (
-
-\subsection{3. Important information is hard to locate}
-
-\subsection{4. Users jump straight to the implementation of the method}
+\subsection{3. Users jump straight to the implementation of the method}
 
@@
 Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
-
-\subsection{An example using contrasts}
-Probably the most well-known, PCM is Felsenstein's \citeyearpar{felsenstein1985phylogenies} phylogenetically independent contrasts (PIC).
-shortly afterwards Grafen introduced the phylogenetic regression also known as phylogenetic generalised least squares (PGLS).
-Later papers confirmed that these methods are actually equivalent if a model of Brownian motion evolution is assumed (although the mathematical proof for this is "awkward"; \citet{rohlf2001comparative,blomberg2012independent})- with PIC being a special case of PGLS \citet{blomberg2012independent,rohle2006comment,rohlf2001comparative,garland2000using}.
-
-Many papers have discussed the assumptions of this method at length.\citealp{diaz1996testing} list these assumptions as 1) topology is perfect; 2) branch lengths are perfect and 3) the model of evolution is Brownian.
-Felsenstein points out the assumption of Brownian motion may not be very useful in his original paper \citet[p.13]{felsenstein1985phylogenies}.
-This was followed up by numerous publications in the 1990s discussing this assumption and providing multiple tests to see if the assumption was correct.
-Many of these papers are extremely explicit about the need to investigate these assumptions.
-
-Assumptions of tree accuracy are less commonly discussed, partly because evidence suggests they make very little difference. Polytomies are dealt with though.
 
@@
 
+%Can we expect methods users to read every paper before using a method?
+(Yes – wouldn’t let someone work on your microscope unless you were sure they knew what they were doing!!!).
+Perhaps some work to be done in summarizing this – needs incentives – no benefits to wikis and workshop teaching cf to writing high impact papers, or papers at all.
+However, people willing to share skills are the people we really want as colleagues, not the selfish high achievers!
+Collaborate with someone who does know this stuff?
+Again problematic as highly skilled technical people end up with too much to do and still end up only being middle author.
+Maybe something like in physics with “theoretical lead” and “technical lead” authors???
+Perhaps write a less technical summary for methods papers, or a “Caveats box” to put this up front – i.e.
+will only work if you have X taxa and this kind of data.
+Check these assumptions.
+Incentivize accompanying blogs to explain this stuff?
+Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
+Speed up publication by archiving.
+Put code on github so knowledgeable users can access at an earlier stage if they want to.
+Helpful for finding bugs etc.
+Methods developers could also benefit from collaboration with methods users who could help them find corner cases and make their methods more useful with real data?
+[Again I’m torn with some of these ideas because is it really the developer’s job to stop people from being fuckwits?
+But if this info isn’t included here, *where* would it come from and *who* would provide it?
+Users can’t be trusted.]
+
 %A New Hope joke...

### 2014-08-21 Analyses/Figure_PCMCitations.R
Fixed draft figure 1

@@ -1,12 +1,10 @@
 ds <- read.csv("PCMCitesThruTime.csv")
 
-# Plot for talk slides
-newred <- rgb(red = 235, green = 81, blue = 83, maxColorValue = 255)
-
-png("PCMCitations.png", bg = "transparent")
+png("PCMCitations.png")
 
 # Plot citations
+par(bty = "l")
 plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
-     col = newred, lwd = 5)
+	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2)
 
 dev.off()

### 2014-08-22 Manuscript/dark.tex
Fixed sizing ofig 1
Added to section on folklore

@@
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
-PCMs have become extremely popular over recent years; the number of papers featured the phrase `phylogenetic comparative' has increased exponentially since the 1980s (Figure \ref{PCMCitations}). 
+PCMs have become extremely popular over recent years; the number of papers featured the phrase ``phylogenetic comparative'' has increased exponentially since the 1980s (Figure \ref{PCMCitations}). 
 With new methods being published almost weekly, there has never been a better time to be a comparative biologist.
@@
 \centering
-\includegraphics[width = 8cm]{PCMCitations.png}
+\includegraphics[width = 10cm]{PCMCitations.png}
 \caption{PCM papers through time}
@@
 Additionally, little consideration is given to whether using a PCM is really appropriate for the question at hand \citep{losos2011seeing}. 
-To paraphrase \citet{blomberg2012independent}, the enthusiastic application of PCMs has greatly outpaced our theoretical understanding of the methods and their relations to the rest of statistical theory.
+To paraphrase \citet{blomberg2012independent}, the enthusiastic application of PCMs has greatly outpaced our theoretical understanding of the methods and their relations to the rest of statistical theory.\\
 
@@
 One major issue is that two groups of people are involved with PCMs. 
-At one extreme, methods-developers (wookies) are involved with developing new methods and implementing them. They are aware of the limitations of their methods, and the assumptions that underly them. 
+At one extreme, methods-developers (``wookies'') develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
 % Bit of a stretch but we can pretend right? :)
-At the other extreme, naive methods-users (ewoks) have no knowledge of these complexities, they just want to use PCMs on their data. 
-They may be graduate students or more established scientists who have been asked to apply a phylogenetic correction by a reviewer. 
-Of course the majority of PCM users fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use.
-Although methods-developers (and advanced methods-users) are aware of the problems with PCMs, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
-Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers, but who should take responsibility for such actions?
-And why has this gap developed in the first place?
-Here we explore these issues and make suggestions for improvements.\\ %needs some rewording
+At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data, and may not be aware of issues with the methods. 
+Of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use.
+The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users.
+Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
+Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers.
+Here we explore the causes of this communication gap and suggest some potential solutions.
+%still not happy with this writing
 
-\section{Why don't all methods-users know about the ``Dark Side'' of PCMs?}
-% Or something related to the break down in communication
+\section{What impedes information transfer from methods-developers to methods-users?}
+% Or something similar
 
-To apply a method correctly, methods-users must be aware of the assumptions and biases of the method in question, and know how to identify whether their model adequately meets these assumptions.
-If methods-users are unaware of these factors there are several possible explanations:
+As scientists we mainly communicate our ideas and through the literature.
+Therefore, any difficulty transferring information between methods-developers and methods-users must begin with the literature.
+If this is the case we suggest three possible reasons:
+%Yuk fix this
 
@@
 \item Not everything is mentioned in the literature.
-\item The literature is too technical for the novice user and/or important details are hidden within the text.
-\item Methods users have bypassed the literature gone straight to the implementation of the method.
+\item The literature is too technical and/or important details are hidden within the text.
+\item Methods-users have bypassed the literature and gone straight to the implementation of the method.
 \end{enumerate}
@@
 A lot of the information needed to properly apply PCMs is not found in the literature. 
-We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students and over time these myths and rules-of-thumb become set in stone. 
+We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students.
+Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb. 
+Other times folklore is more opinion-based, but over time these myths become set in stone. 
 One example of such folklore is in the defaults of programs that perform PCMs. 
 These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.
+Useful PCM folklore is often shared among methods-developers, and among collaborating groups, but is rarely shared outside of these circles. 
+When it is shared, it tends to be in the form of email exchanges between methods-developers and struggling methods-users.
+This advice is extremely useful to the methods-user in question, but forms a kind of ``dark advice'' that isn't accessible to the rest of the community.\\
 
@@
 Other information about the limitations of a method may be absent from the literature due to the time lag between a new method being published and others having time to test and critique it. 
-For example, \citealp{felsenstein1985phylogenies} published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published %add citations.
-This time lag is shorter with more recent methods because more simulations testing the method are now required by journals.
+For example, \citet{felsenstein1985phylogenies} published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published.
+%add citations.
+This time lag is shorter with more recent methods because simulations testing the method are now required by journals.
+However, we suspect there are still many hidden assumptions and biases in virtually all PCMs.
 
-\subsection{2. The literature is technical, dense and important information is hidden}
+\subsection{2. The literature is too technical and/or important details are hidden}
 
-Although a lot of information is not found in the literature (see above), the majority of assumptions and biases are documented somewhere. 
+Although a lot of information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
 A big issue for novice methods-users is that this information is often extremely technical and dense.
@@
 Additionally there is a lot of literature to wade through before getting all of the required information. 
-For example, one of us (NC) reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies}. 
-Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book.
+For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies}. 
+Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. 
+This seems like an excessive amount of reading to expect from a methods-user trying to run a simple empirical analysis.\\
 
-Another issue is that it can sometimes be hard to find the information required.
-Assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper.
-They are rarely neatly corralled.
-
-%Distinction between "warning" papers, and papers that just mention assumptions in passing.
-
-%Yes they are!
-Pictures of warnings papers and some highlighted text from Harvey and Pagel.
-Seems to be disconnect between users and literature, esp older stuff – graph for Felsenstein 85 showing increased citations but without caveats papers cited.
-Caveats papers often cited as but(t) see or fly by citations with no depth.
+Another issue is that it can sometimes be hard to find the information required. Assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
+They are rarely neatly corralled in one place, making it really easy to miss pertinent details. 
+This places an additional burden on methods-users who must read every paper very carefully.
 
 \subsection{3. Users jump straight to the implementation of the method}
+
+In the early days of PCMs there was
+
+Increasingly new PCMs are published with an accompanying R package, or at least some code
 

### 2014-08-24 Manuscript/dark.tex
Minor edits

--- 2014-08-21 Manuscript/dark.tex
@@
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
-PCMs have become extremely popular over recent years; the number of papers featured the phrase `phylogenetic comparative' has increased exponentially since the 1980s (Figure \ref{PCMCitations}). 
+PCMs have also become extremely popular over recent years; the number of papers containing the phrase `phylogenetic comparative' has increased exponentially since the 1980s (Figure \ref{PCMCitations}).
 With new methods being published almost weekly, there has never been a better time to be a comparative biologist.
@@
 
-Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any statistical method \citep{freckleton2009seven,boettiger2012your}. 
+Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method \citep{freckleton2009seven,boettiger2012your}.
 The more popular these methods become, the less awareness methods-users seem to have of this dark side.
@@
 One major issue is that two groups of people are involved with PCMs. 
-At one extreme, methods-developers (wookies) are involved with developing new methods and implementing them. They are aware of the limitations of their methods, and the assumptions that underly them. 
+At one extreme, methods-developers (``wookies'') are involved with developing new methods and implementing them.
+They are aware of the limitations of their methods, and the assumptions that underly them.
 % Bit of a stretch but we can pretend right? :)
-At the other extreme, naive methods-users (ewoks) have no knowledge of these complexities, they just want to use PCMs on their data. 
+At the other extreme, naive methods-users (``ewoks'') have no knowledge of these complexities, they just want to use PCMs on their data. 
 They may be graduate students or more established scientists who have been asked to apply a phylogenetic correction by a reviewer. 

### 2014-08-24 Manuscript/dark.tex

--- 2014-08-22 Manuscript/dark.tex
@@
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
-PCMs have become extremely popular over recent years; the number of papers featured the phrase ``phylogenetic comparative'' has increased exponentially since the 1980s (Figure \ref{PCMCitations}). 
+PCMs have also become extremely popular over recent years; the number of papers containing the phrase `phylogenetic comparative' has increased exponentially since the 1980s (Figure \ref{PCMCitations}).
 With new methods being published almost weekly, there has never been a better time to be a comparative biologist.

### 2014-08-24 Manuscript/dark.tex
Added section on jumping to implementation

@@
 
-In the early days of PCMs there was
+In the early days of PCMs, some methods-developers provided stand alone packages to run their methods, others provided code in whichever language they were comfortable programming in, and still others provided no way of implementing their methods at all.
+This resulted in many frustrating hours (and days and months!) trying to implement any new method you wanted to use.\\
 
-Increasingly new PCMs are published with an accompanying R package, or at least some code
+More recently however, the PCM community has moved towards mostly implementing their methods in R (cite), and code sharing has become almost ubiquitous. 
+The number of R packages for PCMs has increased exponentially since 2005 when ape (A=P=E=, cite) was released, and has increased particularly sharply since 2008 (Figure \ref{x}).
+Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
+We are all big proponents of Open Science so we think this is a wonderful thing for the field, and we do not mean to criticize 
+However, this has led to an increasing number of methods-users jumping straight to the implementation of a method, without really carefully trying to understand what the method is doing, or what it's assumptions might be.
+Additionally, many methods-users seem to try to retrofit a question so they can use the newest shiniest methods available. 
+All of us regularly review papers where the authors have clearly not read the original papers very carefully.
+
+This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit
 

### 2014-08-28 Manuscript/dark.tex
completed draft of problems section

@@
 % Bit of a stretch but we can pretend right? :)
-At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data, and may not be aware of issues with the methods. 
-Of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use.
+At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data, and may not be aware of issues with the methods 
+%when the next sentence is not in brackets it breaks the flow. But needs to be here. Brackets feel wrong too. Will Syst Biol allow footnotes
+(of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
 The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users. 
@@
 
-As scientists we mainly communicate our ideas and through the literature.
-Therefore, any difficulty transferring information between methods-developers and methods-users must begin with the literature.
+As scientists we mainly communicate our ideas through the literature.
+Therefore, any difficulty transferring information between methods-developers and methods-users must begin there.
 If this is the case we suggest three possible reasons:
@@
 A lot of the information needed to properly apply PCMs is not found in the literature. 
-We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students.
+We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and it is occasionally closer to fiction than fact.
 Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb. 
-Other times folklore is more opinion-based, but over time these myths become set in stone. 
+Other times folklore is more opinion-based, but over time these opinions become rules. 
 One example of such folklore is in the defaults of programs that perform PCMs. 
@@
 Other information about the limitations of a method may be absent from the literature due to the time lag between a new method being published and others having time to test and critique it. 
-For example, \citet{felsenstein1985phylogenies} published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published.
+For example, Felsenstein published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published.
 %add citations.
 This time lag is shorter with more recent methods because simulations testing the method are now required by journals.
-However, we suspect there are still many hidden assumptions and biases in virtually all PCMs.
+However, we suspect there are still many hidden assumptions and biases in virtually all new PCMs that have yet to be properly explored in the literature. 
+%We can cite your paper on Pagel 1994 here if you like - as an example of old methods that are still being investigated.
 
@@
 More recently however, the PCM community has moved towards mostly implementing their methods in R (cite), and code sharing has become almost ubiquitous. 
-The number of R packages for PCMs has increased exponentially since 2005 when ape (A=P=E=, cite) was released, and has increased particularly sharply since 2008 (Figure \ref{x}).
+The number of R packages for PCMs has increased exponentially since 2005 when APE (cite) was released, and has increased particularly sharply since 2008 (Figure \ref{x}).
 Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
-We are all big proponents of Open Science so we think this is a wonderful thing for the field, and we do not mean to criticize 
-However, this has led to an increasing number of methods-users jumping straight to the implementation of a method, without really carefully trying to understand what the method is doing, or what it's assumptions might be.
+This has led to an increasing number of methods-users jumping straight to the implementation of a method, without really carefully trying to understand what the method is doing, what its assumptions might be, or what the results they extract really mean in a biological context.
+This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
+The APE book (cite) for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation - cite).\\
+We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology.
+It is extremely frustrating reading/reviewing papers where the authors have clearly not carefully read or understood the primary literature on the method they are using.
 Additionally, many methods-users seem to try to retrofit a question so they can use the newest shiniest methods available.
-All of us regularly review papers where the authors have clearly not read the original papers very carefully.
-
-This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit
-
-%This increase in use is at least partially the result of the number of available R packages to run these analyses (for example ape, GEIGER, diversitree, caper - citations) and the community's increasing familiarity with R \citep{R-Core-Team:2014aa}.
-Unfortunately with this increase in popularity, comes an increase in misuse!
-% need to work on this paragraph
-
-%Huge increase in number of PCM packages since 1980s, and especially since 2008 [Graph].
-Now people familiar with R they can just jump in and try stuff on their data with very little knowledge of what is going on or what the results mean [Though code sharing is cool and I’m not dissing that].
-
-%I’m torn as to what I think here – part of me thinks there should be a mention of assumptions in the manual/vignette (though not if this replicates an accompanying blog or book or something) but part of me thinks it’s the users responsibility to not be a fuckwit.
-Caper, diversitree and CAIC are good examples where assumptions are clearly laid out and methods for testing assumptions are available in caper and CAIC (not sure about diversitree ...).
-Ape book has none of this in it at all!
-
-%Again this needs to be incentivized as it’s adding work for already hard working and underappreciated methods developers.
-Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
+This is all driven by our ``publish or perish'' culture, and the pressure to produce as many papers as possible out of a dataset.
+The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have made the situation worse rather than better (\citealp{freckleton2009seven}, for example, mentions R as a way of improving things).
+%needs a bit of finessing
 
@@
 
+The problems described above require action by methods-developers, methods-users and the community as a whole.
+%Can't expect empiricists to be an expert in every method, but do want to add to workload of overloaded developers.
+
 \subsection{1. Simplify, summarize and share}
-\subsection{2. Read, think and learn}
-\subsection{3. Collaboration}
-\subsection{4. Incentives}
 
-%Such issues are the responsibility of end users but also of methods developers: the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
-To address these issues, we propose this symposium to discuss issues in both classical and recent PCMs, along with new research on detecting for these issues and accounting for them.
-We hope that this will both increase awareness of these problems and encourage further research and careful thought in the area, along with better dialogue between method developers and method users.
-
-%Big issue = who is responsible? Can't expect empiricists to be an expert in every method, but do want to add to workload of overloaded developers.
-
-%Improvements in methods, and ways of detecting biases etc should be sufficient for publication (rather than MEE's weird focus on novelty now it has a huge IF)
-
-%Incentives - funding for pure methods. Different authorship rules.
-
-%Change the definition of success in academia. Quality not quantity. Remove rush to publish. This would help WOMEN too!
-
-%Collaborate! But make collaboration valuable to non first authors somehow
-
-%We really *do not* want to lose all the technical people from science.
-We need them more than ever as evo bio becomes more and more computationally intensive.
-Need to make it worth their while (can't match wages of industry)
-
-%Add methods to test assumptions as standard to packages.
+a) dealing with
 
@@
 %Some website with a list of key papers, and key summary points, for each method. Who would host? Who would write? Who would peer review?
+
+May not be doable... List too many assumptions...
+Discourse pages – forums for particular things. Dark “advice”.
+Reduces time needed. 
+Phylobabble
+
+\subsection{2. Read, think and learn}
 
@@
 My pglm vs pgls example.
+
+\subsection{3. Collaboration}
+
+A obvious solution to all these problems is for methods-developers and methods-users to collaborate more.
+The benefits to methods-users is obvious, but methods-developers can gain extra datasets to test their ideas on, people who will discover corner cases and bugs in their software before it get released, and gain an appreciation of .
+A big difficulty with collaboration
+
+%Collaborate! But make collaboration valuable to non first authors somehow
+
+\subsection{4. Incentives}
+
+%Improvements in methods, and ways of detecting biases etc should be sufficient for publication (rather than MEE's weird focus on novelty now it has a huge IF)
+
+%Incentives - funding for pure methods. Different authorship rules. 
+%Change the definition of success in academia. Quality not quantity. Remove rush to publish. This would help WOMEN too!
+
+%We really *do not* want to lose all the technical people from science.
+We need them more than ever as evo bio becomes more and more computationally intensive.
+Need to make it worth their while (can't match wages of industry)
+
+%Add methods to test assumptions as standard to packages. 
+%I’m torn as to what I think here – part of me thinks there should be a mention of assumptions in the manual/vignette (though not if this replicates an accompanying blog or book or something) but part of me thinks it’s the users responsibility to not be a fuckwit.
+Caper, diversitree and CAIC are good examples where assumptions are clearly laid out and methods for testing assumptions are available in caper and CAIC (not sure about diversitree ...).
+Ape book has none of this in it at all!
+
+%Again this needs to be incentivized as it’s adding work for already hard working and underappreciated methods developers.
+Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
+
+Encourage quality over quantity % this is how I'm apparently allowed to justify my publication record
+Additionally, many methods-users seem to try to retrofit a question so they can use the newest shiniest methods available.
 

### 2014-08-28 Manuscript/dark.tex
Added sections on solutions 1-3

@@
 
-The problems described above require action by methods-developers, methods-users and the community as a whole.
+%The problems described above require action by methods-developers, methods-users and the community as a whole.
 %Can't expect empiricists to be an expert in every method, but do want to add to workload of overloaded developers.
+%Can we expect methods users to read every paper before using a method? (Yes – wouldn’t let someone work on your microscope unless you were sure they knew what they were doing!!!).
 
@@
 
-a) dealing with
+Many of the problems above are caused by methods papers being difficult to read. 
+This isn't entirely the fault of methods-developers; most journals have strict word limits so it's often easier to use a one line equation rather than a paragraph of description. 
+However, it's hard to complain about novice users failing to understand a method when the paper is incomprehensible. %cough cough Hansen. Ow my eyes are bleeding.
+One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 
+Some journals already encourage this (e.g. Methods in Ecology and Evolution), and even if they do not, there is no harm in doing this and hosting it on a personal website. 
+The ability to share ideas with a non-technical audience is a key skill to develop, and it may even encourage more people to use the method. 
+%Another more %knitr reports in supp mat (e.g. cite that nice paper )
+At the very least, having a simple list of the main assumptions, biases and caveats of the method somewhere would reduce misuse and provide a place to point methods-users to when they begin using a method.
+Of course this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible.\\
 
-%Grouping assumptions, biases and caveats more carefully in papers. Caveats box???
+%Want to mention adding ways of testing assumptions to packages
+%follow up papers (hard to publish coz they are incremental - see incentives)
+%or should this go in incentives...?
 
-%Accompanying blogs with methods papers to explain to non-technical audience. Could be Supp Matt too. Lavin et al 2008 is a beautiful example.
+The glut of literature that already exists for established methods is a bigger problem. 
+One solution would be to establish a wiki-style website where methods-developers and advanced methods-users could post summaries of commonly used methods, along with lists of key papers to read. 
+The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
+Eric Matsen's Phylobabble discourse page (link) is an excellent example of a similar approach for phylogenetics.
+Unfortunately this would be hard to arrange as there is little incentive for methods-developers and advanced methods-users to contribute to such a project at the cost of time pursuing their own research.\\
 
-%Some website with a list of key papers, and key summary points, for each method. Who would host? Who would write? Who would peer review?
-
-May not be doable... List too many assumptions...
-Discourse pages – forums for particular things. Dark “advice”.
-Reduces time needed. 
-Phylobabble
+One solution we believe will be of use is establishing a discourse page for providing R help. 
+Currently most methods-users either email the R-sig-phylo help list or the package developer for assistance. 
+Both approaches are suboptimal. 
+R-sig-phylo is impossible to search effectively, leading to the same questions being dealt with repeatedly. 
+The same is true in email exchanges between methods-users and package developers with the added issue of this being ``dark advice'' that isn't shared. 
+A discourse page would collate all of this information and reduce pressure on package developers by allowing advanced methods-users to help novice users. 
+This would be a valuable resource for methods-users and also reduce the amount of time methods-developers spend dealing with queries.
+We floated this idea at Evolution 2014 and hope to have a version of the discourse page up and running soon.
 
 \subsection{2. Read, think and learn}
+%This section is currently a bit personal. You should do this and that, rather than methods-users should do this and that.
+If methods-developers are going to put time and effort into making their methods easier to understand, then methods-users need to take advantage of this and work harder to understand the methods they are using.
+At a minimum, methods-users should read the original papers describing a method, and any recent updates. 
+Jumping straight in to an R implementation is a fine way to start a project, but it needs to be followed up by careful reading and thought. 
+Playing with shiny new methods is always fun, but don't retrofit your question to the new method. 
+Think carefully about the question you want to ask, and whether the method really answers that question. 
+Also consider whether you really need to use a phylogeny for your analysis at all \citep{losos2011seeing}.\\
 
-%Users need to take responsibility too.
-Work harder to understand.
-Read widely.
-Don't take results at face value.
-Become less trusting!
-Simulate data to check method does what you think it should do.
-Don't be afraid to question standard practice - sometimes it's just folklore.
-My pglm vs pgls example.
+Another key skill to develop is distrust! 
+Never take results from PCMs (or any other statistical analysis) at face value. 
+Always think carefully about whether there may be hidden assumptions or biases that have not been explored in the accompanying papers. 
+A good way to check this is to simulate some data and see if you get the results you expect. 
+Also consider whether the method will work on your dataset; how many species does it need to have reasonable power, how does it deal with polytomies, can it cope with non-ultrametric trees etc.
+Finally, never be afraid to question standard practice, sometimes it's just PCM folklore
+%Could add my pglm vs pgls example.
 
@@
 A obvious solution to all these problems is for methods-developers and methods-users to collaborate more. 
-The benefits to methods-users is obvious, but methods-developers can gain extra datasets to test their ideas on, people who will discover corner cases and bugs in their software before it get released, and gain an appreciation of .
-A big difficulty with collaboration
-
-%Collaborate! But make collaboration valuable to non first authors somehow
+The benefits to methods-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
+A big difficulty with collaboration is that it is often asymmetrical in terms of benefit to methods-users and methods-developers. 
+Methods-developers and advanced methods-users can often put a lot of time and effort into running analyses for less experienced colleagues, yet at best they come out of the experience with a middle authored paper.
+Until these author positions are considered equal to first and last author positions, there is little incentive to collaborate extensively.
 
@@
 Encourage quality over quantity % this is how I'm apparently allowed to justify my publication record
-Additionally, many methods-users seem to try to retrofit a question so they can use the newest shiniest methods available.
+Discourage retrofitting a question so they can use the newest shiniest methods available.
 
-%Can we expect methods users to read every paper before using a method?
-(Yes – wouldn’t let someone work on your microscope unless you were sure they knew what they were doing!!!).
-Perhaps some work to be done in summarizing this – needs incentives – no benefits to wikis and workshop teaching cf to writing high impact papers, or papers at all.
+%
+
+%Perhaps some work to be done in summarizing this – needs incentives – no benefits to wikis and workshop teaching cf to writing high impact papers, or papers at all.
 However, people willing to share skills are the people we really want as colleagues, not the selfish high achievers!
-Collaborate with someone who does know this stuff?
-Again problematic as highly skilled technical people end up with too much to do and still end up only being middle author.
-Maybe something like in physics with “theoretical lead” and “technical lead” authors???
-Perhaps write a less technical summary for methods papers, or a “Caveats box” to put this up front – i.e.
-will only work if you have X taxa and this kind of data.
-Check these assumptions.
-Incentivize accompanying blogs to explain this stuff?
-Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
-Speed up publication by archiving.
-Put code on github so knowledgeable users can access at an earlier stage if they want to.
-Helpful for finding bugs etc.
-Methods developers could also benefit from collaboration with methods users who could help them find corner cases and make their methods more useful with real data?
-[Again I’m torn with some of these ideas because is it really the developer’s job to stop people from being fuckwits?
-But if this info isn’t included here, *where* would it come from and *who* would provide it?
-Users can’t be trusted.]
 
-%A New Hope joke...
+%Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
+%Speed up publication by archiving. 
+%Put code on github so knowledgeable users can access at an earlier stage if they want to. Helpful for finding bugs etc.
+
+\section{Conclusion}
+
+Hopefully we can banish the ``Dark Side'' of PCMs and create a ``New Hope'' for the whole PCM community, whether they be wookies, ewoks or somewhere in between.
 
@@
 \section{Acknowledgments}
-Thanks to the Society of Systematic Biologists for funding our symposium at Evolution 2014. We also thank George Lucas for the ``hilarious'' Star Wars jokes.
+Thanks to the Society of Systematic Biologists for funding our symposium at Evolution 2014. 
+We also thank George Lucas for the ``hilarious'' Star Wars jokes.
+%Should possible thank Brian O'Meara for funding discourse page
 

### 2014-08-29 Analyses/Figure_PCMRpackages.R
Added draft figure for R packages over time

@@ -1,0 +1,27 @@
+library(reshape)
+
+ds <- read.csv("PackagesThruTime.csv")
+
+# Group packages by year using aggregate then fix names
+dsYear <- with(ds, aggregate(Program, by = list(YearReleased, RPackage), 
+	                          FUN = length))
+
+names(dsYear) <- c("YearReleased", "R", "NPackages")
+
+# Order the data by year so cumulative number of packages
+# per year can be calculated
+dsYear <- dsYear[order(dsYear$YearReleased), ]
+
+# Calculate cumulative number of packages per year
+# separately for non R and R packages
+dsYear$TotalPkg[dsYear$R == "other"] <- cumsum(dsYear$NPackages[dsYear$R == "other"])
+dsYear$TotalPkg[dsYear$R == "R"] <- cumsum(dsYear$NPackages[dsYear$R == "R"])
+
+png("PCMRpackages.png")
+
+par(bty = "l")
+plot(TotalPkg[R == "R"] ~ YearReleased[R == "R"], data = dsYear, 
+     type = "l" , las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2,
+	 xlim = c(2000,2014), xlab = "Year Released", ylab = "Total Packages")
+
+dev.off()

### 2014-08-29 Manuscript/dark.tex
Added page numbers
Added small part ot incentives section

@@
 \usepackage{graphicx}
+\usepackage{parskip}      
 \pagenumbering{arabic}
@@
 \linespread{1.66}
-\raggedright
-\setlength{\parindent}{0.5in}
-
-\pagestyle{empty}
 
@@
   $^3$ UBC\\
-$^4$ Institute for Bioinformatics and Evolutionary Studies, University
-of Idaho, Moscow, ID 83844, U.S.A.\\
+  $^4$ Institute for Bioinformatics and Evolutionary Studies, University of Idaho, Moscow, ID 83844, U.S.A.\\
   $^5$ Sheffield\\
@@
   $^8$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-$^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+  $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. 
+       Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
 }}
@@
 \doublespacing
+\setlength{\parindent}{1cm}
+
 %\section{Introduction} % No intro heading for Syst Biol
 
+\noindent
 A long time ago in a galaxy far, far away - also known as the 1980s - phylogenetic comparative methods (PCMs) were developed to deal with the statistical non-independence of species in comparative analyses (e.g.
@@
 % Or something similar
-
 As scientists we mainly communicate our ideas through the literature.
@@
 \subsection{1. Not everything is mentioned in the literature}
-
 A lot of the information needed to properly apply PCMs is not found in the literature. 
@@
 \subsection{2. The literature is too technical and/or important details are hidden}
-
 Although a lot of information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
@@
 \subsection{3. Users jump straight to the implementation of the method}
-
 In the early days of PCMs, some methods-developers provided stand alone packages to run their methods, others provided code in whichever language they were comfortable programming in, and still others provided no way of implementing their methods at all.
@@
 More recently however, the PCM community has moved towards mostly implementing their methods in R (cite), and code sharing has become almost ubiquitous. 
-The number of R packages for PCMs has increased exponentially since 2005 when APE (cite) was released, and has increased particularly sharply since 2008 (Figure \ref{x}).
+The number of R packages for PCMs has increased exponentially since 2005 when APE (cite) was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
 Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
@@
 
+\begin{figure}[h]
+\centering
+\includegraphics[width = 10cm]{PCMRpackages.png}
+\caption{Increase in R packages through time}
+\label{PCMRpackages}
+\end{figure}
+
 \section{How can we solve some of these problems?}
@@
 \subsection{1. Simplify, summarize and share}
-
 Many of the problems above are caused by methods papers being difficult to read. 
@@
 \subsection{3. Collaboration}
-
 A obvious solution to all these problems is for methods-developers and methods-users to collaborate more. 
@@
 
-%Improvements in methods, and ways of detecting biases etc should be sufficient for publication (rather than MEE's weird focus on novelty now it has a huge IF)
+For any of the solutions above to work we need to incentivize more thoughtful methods development and use.
+
+First we need to stop insisting that methods papers must be entirely novel.
+Improvements to existing methods, and ways of detecting biases or testing model fit should be sufficient for publication.
+These kinds of papers are arguably more useful to the community than constantly publishing new methods as this encourages methods-users to use the shiniest new methods, rather than carefully thinking about the question they are trying to answer.
+Journals also have a part to play in encouraging methods-developers to develop ways of testing the assumptions of their methods and building this into the original publications.
+
+Second we need to fund pure methods development, including incremental methods.
+Currently it is difficult to get funding for purely methods driven research; and empirical component is always needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
+This unfairly prejudices funding bodies, and hiring committees, away from methods-developers.
+Our ability to do science often relies on methods-developers we can't lose them!!!
+
+Finally we need to re-evaluate our criteria for success in academia. The ``publish or perish'' mentality is not conducive to careful method use or attempts to fully understand a method prior
 

### 2014-08-31 Manuscript/dark.tex
Added citations

@@
 
-More recently however, the PCM community has moved towards mostly implementing their methods in R (cite), and code sharing has become almost ubiquitous. 
-The number of R packages for PCMs has increased exponentially since 2005 when APE (cite) was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
+More recently however, the PCM community has moved towards mostly implementing their methods in R \citep{R-Core-Team:2014aa}, and code sharing has become almost ubiquitous. 
+The number of R packages for PCMs has increased exponentially since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
 Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
@@
 This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
-The APE book (cite) for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation - cite).\\
+The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citep{purvis1995comparative,Orme:2013aa}).\\
 We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology.

### 2014-09-01 Manuscript/dark.tex
Small edits

@@
 
-% Ignoring their title page setup as I like Rich's template better :)
+% NC: Ignoring their title page setup as I like Rich's template better :)
 
@@
 %\section{abstract}
-% If Syst Biol "Point of View" there's no abstract
+% NC: If Syst Biol "Point of View" there's no abstract
 
@@
 
-%\section{Introduction} % No intro heading for Syst Biol
+%\section{Introduction} 
+% NC: No intro heading for Syst Biol
 
@@
 
-% Do we need more explicit comments showing we're not the only people to think this, nor the first to say it?
+% NC: Do we need more explicit comments showing we're not the only people to think this, nor the first to say it?
 
@@
 At one extreme, methods-developers (``wookies'') develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
-% Bit of a stretch but we can pretend right? :)
-At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data, and may not be aware of issues with the methods 
-%when the next sentence is not in brackets it breaks the flow. But needs to be here. Brackets feel wrong too. Will Syst Biol allow footnotes
-(of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
+At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data, and may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
 The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users. 
@@
 Here we explore the causes of this communication gap and suggest some potential solutions.
-%still not happy with this writing
+% NC: still not happy with this final line or two.
 
 \section{What impedes information transfer from methods-developers to methods-users?}
-% Or something similar
+% NC: Or something similar
 As scientists we mainly communicate our ideas through the literature. 
-Therefore, any difficulty transferring information between methods-developers and methods-users must begin there.
-If this is the case we suggest three possible reasons:
-%Yuk fix this
+Below we propose three possible reasons for difficulties transferring information between methods-developers and methods-users via the literature.
 
@@
 Hopefully some of the changes we suggest below will also help with this problem.
-%This could be a footnote, or left out!
+% NC: This could be left out!
 
@@
 
-% Do we want a specific example here? Theta estimation in OU models? 
-% Could mention ±3 from Jones&Purvis97 - totally unjustified rule of thumb now = canon.
-% Mention hidden assumptions?
+% NC: Do we want a specific example here? Theta estimation in OU models? 
+% NC: Could mention ±3 from Jones&Purvis97 - totally unjustified rule of thumb now = canon.
+% NC: Mention hidden assumptions? I mention them later
 
@@
 However, we suspect there are still many hidden assumptions and biases in virtually all new PCMs that have yet to be properly explored in the literature. 
-%We can cite your paper on Pagel 1994 here if you like - as an example of old methods that are still being investigated.
+% NC: We can cite your paper on Pagel 1994 here if you like - as an example of old methods that are still being investigated.
 
@@
 For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies}. 
-Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. 
+Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. % add numbers
 This seems like an excessive amount of reading to expect from a methods-user trying to run a simple empirical analysis.\\
@@
 Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
-This has led to an increasing number of methods-users jumping straight to the implementation of a method, without really carefully trying to understand what the method is doing, what its assumptions might be, or what the results they extract really mean in a biological context.
+This has led to an increasing number of methods-users jumping straight to the implementation of a method, without understanding what the method is doing, what its assumptions might be, or what the results they extract mean in a biological context.
 This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
@@
 It is extremely frustrating reading/reviewing papers where the authors have clearly not carefully read or understood the primary literature on the method they are using.
-Additionally, many methods-users seem to try to retrofit a question so they can use the newest shiniest methods available.
+Additionally, many methods-users seem to try to retrofit a question so they can use the shiniest new methods available.
 This is all driven by our ``publish or perish'' culture, and the pressure to produce as many papers as possible out of a dataset.
-The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have made the situation worse rather than better (\citealp{freckleton2009seven}, for example, mentions R as a way of improving things).
-%needs a bit of finessing
+The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have made the situation worse rather than better.
+Particularly when the flexibility of R should improve the abilit methods \citealp{freckleton2009seven}.
+%check Freck
 
@@
 
-%The problems described above require action by methods-developers, methods-users and the community as a whole.
-%Can't expect empiricists to be an expert in every method, but do want to add to workload of overloaded developers.
-%Can we expect methods users to read every paper before using a method? (Yes – wouldn’t let someone work on your microscope unless you were sure they knew what they were doing!!!).
-
 \subsection{1. Simplify, summarize and share}
@@
 The ability to share ideas with a non-technical audience is a key skill to develop, and it may even encourage more people to use the method. 
-%Another more %knitr reports in supp mat (e.g. cite that nice paper )
+% NC: Another more %knitr reports in supp mat (e.g. cite that nice paper )
 At the very least, having a simple list of the main assumptions, biases and caveats of the method somewhere would reduce misuse and provide a place to point methods-users to when they begin using a method.
@@
 The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
-Eric Matsen's Phylobabble discourse page (link) is an excellent example of a similar approach for phylogenetics.
+Erick Matsen's Phylobabble discourse page (link) is an excellent example of a similar approach for phylogenetics.
 Unfortunately this would be hard to arrange as there is little incentive for methods-developers and advanced methods-users to contribute to such a project at the cost of time pursuing their own research.\\
@@
 \subsection{2. Read, think and learn}
-%This section is currently a bit personal. You should do this and that, rather than methods-users should do this and that.
 If methods-developers are going to put time and effort into making their methods easier to understand, then methods-users need to take advantage of this and work harder to understand the methods they are using.
@@
 
-For any of the solutions above to work we need to incentivize more thoughtful methods development and use.
+Our ability to do science often relies on methods-developers and other highly-skilled technical people, especially as evolutionary biology becomes ever more computationally intensive. 
+We cannot afford to lose these people to industry, nor can we pay industry wages, thus we need to make it worthwhile for methods-developers to remain in (or at least interact with) academia.
 
@@
 Improvements to existing methods, and ways of detecting biases or testing model fit should be sufficient for publication. 
-These kinds of papers are arguably more useful to the community than constantly publishing new methods as this encourages methods-users to use the shiniest new methods, rather than carefully thinking about the question they are trying to answer.
-Journals also have a part to play in encouraging methods-developers to develop ways of testing the assumptions of their methods and building this into the original publications.
+These kinds of papers are arguably more useful to the community than constantly publishing new methods as this encourages methods-users to retrofit questions so they can use the newest shiniest methods available, rather than carefully thinking about the question they are trying to answer.
+Journals also have a part to play in encouraging methods-developers to develop ways of testing the assumptions of their methods and building this into the original publications and implementations.
 
 Second we need to fund pure methods development, including incremental methods. 
-Currently it is difficult to get funding for purely methods driven research; and empirical component is always needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
+Currently it is difficult to get funding for purely methods driven research; an empirical component is always needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
 This unfairly prejudices funding bodies, and hiring committees, away from methods-developers.
-Our ability to do science often relies on methods-developers we can't lose them!!!
 
-Finally we need to re-evaluate our criteria for success in academia. The ``publish or perish'' mentality is not conducive to careful method use or attempts to fully understand a method prior
-
-%Incentives - funding for pure methods. Different authorship rules. 
-%Change the definition of success in academia. Quality not quantity. Remove rush to publish. This would help WOMEN too!
-
-%We really *do not* want to lose all the technical people from science.
-We need them more than ever as evo bio becomes more and more computationally intensive.
-Need to make it worth their while (can't match wages of industry)
-
-%Add methods to test assumptions as standard to packages. 
-%I’m torn as to what I think here – part of me thinks there should be a mention of assumptions in the manual/vignette (though not if this replicates an accompanying blog or book or something) but part of me thinks it’s the users responsibility to not be a fuckwit.
-Caper, diversitree and CAIC are good examples where assumptions are clearly laid out and methods for testing assumptions are available in caper and CAIC (not sure about diversitree ...).
-Ape book has none of this in it at all!
-
-%Again this needs to be incentivized as it’s adding work for already hard working and underappreciated methods developers.
-Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
-
-Encourage quality over quantity % this is how I'm apparently allowed to justify my publication record
-Discourage retrofitting a question so they can use the newest shiniest methods available.
-
-%
-
-%Perhaps some work to be done in summarizing this – needs incentives – no benefits to wikis and workshop teaching cf to writing high impact papers, or papers at all.
-However, people willing to share skills are the people we really want as colleagues, not the selfish high achievers!
-
-%Funding of pure methods to encourage methods developers to make methods more user friendly, and to take the time to add methods for detecting deviations from assumptions rather than rushing to publish.
-%Speed up publication by archiving. 
-%Put code on github so knowledgeable users can access at an earlier stage if they want to. Helpful for finding bugs etc.
+Finally we need to re-evaluate our criteria for success in academia. 
+The ``publish or perish'' mentality is not conducive to careful method use or attempts to fully understand a method prior to using it. 
+Nor does it encourage methods-developers to spend more time summarizing their work and making their methods more user-friendly. 
+If publication quality were considered more important that quantity this would improve career prospects for methods-developers, careful methods-users and also for groups that traditionally publish fewer papers, including women.
+This is particularly important as the people willing to take the time to share their skills are the people we really want as colleagues, not the selfish high achievers with hundreds of papers!
 
 \section{Conclusion}
+
+We feel that phylogenetic comparative methods have fallen into a bit of a rut in recent years. Their increased popularity and easier implementation have led to numerous sloppy publications.
 

### 2014-09-01 Manuscript/dark.tex
Fixed up to incentives

@@
 
-Other information about the limitations of a method may be absent from the literature due to the time lag between a new method being published and others having time to test and critique it. 
-For example, Felsenstein published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published.
-%add citations.
-This time lag is shorter with more recent methods because simulations testing the method are now required by journals.
-However, we suspect there are still many hidden assumptions and biases in virtually all new PCMs that have yet to be properly explored in the literature. 
-% NC: We can cite your paper on Pagel 1994 here if you like - as an example of old methods that are still being investigated.
+Other information about the limitations of a method may be absent from the literature due to the time-lag between a new method being published and others having time to test and critique it. 
+For example, Felsenstein published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published (e.g.
+\citealp{garland1992procedures})
+This time-lag is shorter with more recent methods because simulations testing the method are now required by journals.
+However, we suspect there are still many hidden assumptions and biases in all new PCMs that have yet to be properly explored in the literature. 
+% NC: Rich, we can cite your paper on Pagel 1994 here if you like - as an example of old methods that are still being investigated.
 
@@
 For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies}. 
-Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. % add numbers
+Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. % NC: add numbers
 This seems like an excessive amount of reading to expect from a methods-user trying to run a simple empirical analysis.\\
@@
 Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
-This has led to an increasing number of methods-users jumping straight to the implementation of a method, without understanding what the method is doing, what its assumptions might be, or what the results they extract mean in a biological context.
+\citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve the ability of methods-users to implement methods correctly. 
+However, this has instead led to an increasing number of methods-users jumping straight to the implementation of a method, without understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
 This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
-The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citep{purvis1995comparative,Orme:2013aa}).\\
+The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citep{purvis1995comparative,Orme:2013aa}.\\
 We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology.
-It is extremely frustrating reading/reviewing papers where the authors have clearly not carefully read or understood the primary literature on the method they are using.
+It is extremely frustrating reading/reviewing papers where the authors have clearly failed to carefully read or understand the primary literature on the method they are using.
 Additionally, many methods-users seem to try to retrofit a question so they can use the shiniest new methods available.
@@
 The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have made the situation worse rather than better.
-Particularly when the flexibility of R should improve the abilit methods \citealp{freckleton2009seven}.
-%check Freck
 
@@
 The ability to share ideas with a non-technical audience is a key skill to develop, and it may even encourage more people to use the method. 
-% NC: Another more %knitr reports in supp mat (e.g. cite that nice paper )
+% NC: Perhaps add something about having reproducible %knitr reports in supp mat (e.g. cite that nice recent paper)
 At the very least, having a simple list of the main assumptions, biases and caveats of the method somewhere would reduce misuse and provide a place to point methods-users to when they begin using a method.
 Of course this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible.\\
-
-%Want to mention adding ways of testing assumptions to packages
-%follow up papers (hard to publish coz they are incremental - see incentives)
-%or should this go in incentives...?
 
@@
 The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
-Erick Matsen's Phylobabble discourse page (link) is an excellent example of a similar approach for phylogenetics.
-Unfortunately this would be hard to arrange as there is little incentive for methods-developers and advanced methods-users to contribute to such a project at the cost of time pursuing their own research.\\
+Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) is an excellent example of a similar approach for phylogenetics.
+Unfortunately there is currently little incentive for methods-developers and advanced methods-users to contribute to such a project at the cost of time pursuing their own research.\\
 
-One solution we believe will be of use is establishing a discourse page for providing R help. 
+One solution is to establish a discourse page for providing R help. 
 Currently most methods-users either email the R-sig-phylo help list or the package developer for assistance. 
@@
 The same is true in email exchanges between methods-users and package developers with the added issue of this being ``dark advice'' that isn't shared. 
-A discourse page would collate all of this information and reduce pressure on package developers by allowing advanced methods-users to help novice users. 
-This would be a valuable resource for methods-users and also reduce the amount of time methods-developers spend dealing with queries.
-We floated this idea at Evolution 2014 and hope to have a version of the discourse page up and running soon.
+A discourse page would collate all of this information and reduce pressure on package developers by allowing advanced methods-users to help novice users with simple queries. 
+We are in the process of setting up a discourse page for this purpose (it will be called ``phyloBootstrap'' because it will provide ``support'' for PCMs).
+At Evolution 2014 many methods-developers agreed to move their R package help to this forum.
 
@@
 At a minimum, methods-users should read the original papers describing a method, and any recent updates. 
-Jumping straight in to an R implementation is a fine way to start a project, but it needs to be followed up by careful reading and thought. 
-Playing with shiny new methods is always fun, but don't retrofit your question to the new method. 
-Think carefully about the question you want to ask, and whether the method really answers that question. 
-Also consider whether you really need to use a phylogeny for your analysis at all \citep{losos2011seeing}.\\
+Jumping straight into an R implementation is a tolerable way to start a project, but it needs to be followed up by careful reading and thought. 
+Methods-users also need to avoid retrofitting their question to shiny new methods, instead they should think carefully about the question they want to ask, whether the method is appropriate for the question, and whether PCMs are
+needed at all \citep{losos2011seeing}.\\
 
-Another key skill to develop is distrust! 
-Never take results from PCMs (or any other statistical analysis) at face value. 
-Always think carefully about whether there may be hidden assumptions or biases that have not been explored in the accompanying papers. 
-A good way to check this is to simulate some data and see if you get the results you expect. 
-Also consider whether the method will work on your dataset; how many species does it need to have reasonable power, how does it deal with polytomies, can it cope with non-ultrametric trees etc.
-Finally, never be afraid to question standard practice, sometimes it's just PCM folklore
-%Could add my pglm vs pgls example.
+Another key skill to develop is distrust; never take results from PCMs (or any other statistical analysis) at face value. 
+A good way to check a method is to simulate some data and see if the results are as expected. 
+This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
+It is also important for methods-users to consider whether the method will work on their dataset; for example, how many species does it need to have reasonable power, how does it deal with polytomies, can it cope with non-ultrametric trees etc.
+Finally, methods-users should never be afraid to question standard practice, sometimes it's just PCM folklore!
+% NC: Could add my pglm vs pgls example.
 
@@
 We also thank George Lucas for the ``hilarious'' Star Wars jokes.
-%Should possible thank Brian O'Meara for funding discourse page
+%NC: Should possibly thank Brian O'Meara for funding discourse page
 

### 2014-09-01 Manuscript/dark.tex
Fixed incentives section

@@
 
-\title{The ``Dark Side'' of Phylogenetic Comparative Methods}
+\title{The ``Dark Side'' of Phylogenetic Comparative Methods$^\dag$}
 \author{
@@
        Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+  $^\dag$This paper is based on a Society of Systematic Biologists symposium at Evolution 2014, in Raleigh NC.\\
 }}
@@
 For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies}. 
-Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. % NC: add numbers
+Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. % NC: I will add numbers at some point!
 This seems like an excessive amount of reading to expect from a methods-user trying to run a simple empirical analysis.\\
@@
 The ability to share ideas with a non-technical audience is a key skill to develop, and it may even encourage more people to use the method. 
-% NC: Perhaps add something about having reproducible %knitr reports in supp mat (e.g. cite that nice recent paper)
+% NC: Perhaps add something about having reproducible knitr reports in supp mat (e.g. cite that nice recent paper)
 At the very least, having a simple list of the main assumptions, biases and caveats of the method somewhere would reduce misuse and provide a place to point methods-users to when they begin using a method.
@@
 \subsection{4. Incentives}
-
-Our ability to do science often relies on methods-developers and other highly-skilled technical people, especially as evolutionary biology becomes ever more computationally intensive. 
-We cannot afford to lose these people to industry, nor can we pay industry wages, thus we need to make it worthwhile for methods-developers to remain in (or at least interact with) academia.
+It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
+We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for methods-developers to remain in (or at least interact with) academia.
 
@@
 Improvements to existing methods, and ways of detecting biases or testing model fit should be sufficient for publication. 
-These kinds of papers are arguably more useful to the community than constantly publishing new methods as this encourages methods-users to retrofit questions so they can use the newest shiniest methods available, rather than carefully thinking about the question they are trying to answer.
-Journals also have a part to play in encouraging methods-developers to develop ways of testing the assumptions of their methods and building this into the original publications and implementations.
+This is fairly standard in other fields, e.g. phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
+When novel methods are published, journals should encourage methods-developers to include ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.
 
@@
 Currently it is difficult to get funding for purely methods driven research; an empirical component is always needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
-This unfairly prejudices funding bodies, and hiring committees, away from methods-developers.
+This unfairly prejudices funding bodies (and hiring committees) away from methods-developers.
 
 Finally we need to re-evaluate our criteria for success in academia. 
-The ``publish or perish'' mentality is not conducive to careful method use or attempts to fully understand a method prior to using it. 
-Nor does it encourage methods-developers to spend more time summarizing their work and making their methods more user-friendly. 
-If publication quality were considered more important that quantity this would improve career prospects for methods-developers, careful methods-users and also for groups that traditionally publish fewer papers, including women.
-This is particularly important as the people willing to take the time to share their skills are the people we really want as colleagues, not the selfish high achievers with hundreds of papers!
+The ``publish or perish'' mentality does not encourage methods-developers to spend more time on their methods than necessary. 
+Thus there is little incentive to summarize, simplify and share, or to collaborate with methods-users, as we suggest above. 
+This mentality is also not conducive to methods-users attempting to fully understand a method prior to using it. 
+Instead we need to value quality over quantity when it comes to publications.
+This would improve career prospects for both methods-developers and careful methods-users.
 
 \section{Conclusion}
+% NC: Still need a couple of lines of conclusion but I'm drawing a blank. Final sentence is fine I think. We could just skip the conclusion heading and have this as a standalone sentence at the end.
 
-We feel that phylogenetic comparative methods have fallen into a bit of a rut in recent years. Their increased popularity and easier implementation have led to numerous sloppy publications.
-
-Hopefully we can banish the ``Dark Side'' of PCMs and create a ``New Hope'' for the whole PCM community, whether they be wookies, ewoks or somewhere in between.
+Hopefully by trying to implement some of these solutions we can banish the ``Dark Side'' of PCMs and create a ``New Hope'' for the whole PCM community, whether they be wookies, ewoks or somewhere in between.
 

### 2014-09-02 Manuscript/dark.tex
Fixed author order as requested

@@
 \author{
-  Richard G. FitzJohn$^{1}$, Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\ Matthew W. Pennell$^{4}$, Gavin H. Thomas$^{5}$, April Wright$^{6}$,\\ and Natalie Cooper$^{7,8,*}$
+  Natalie Cooper$^{1,2,*}$, Matthew W. Pennell$^{3}$ \\
+  and Richard G. FitzJohn$^{4}$ 
+  %Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\  Gavin H. Thomas$^{5}$, April Wright$^{6}$,\\
 }
@@
 \affiliation{\noindent{\footnotesize
-  $^1$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
-  $^2$ Wisconsin\\
-  $^3$ UBC\\
-  $^4$ Institute for Bioinformatics and Evolutionary Studies, University of Idaho, Moscow, ID 83844, U.S.A.\\
-  $^5$ Sheffield\\
-  $^6$ Austin\\
-  $^7$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
-  $^8$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
+
+ % $^2$ Wisconsin\\
+ % $^3$ UBC\\
+ % $^5$ Sheffield\\
+ % $^6$ Austin\\
+  $^1$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
+  $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
+  $^3$ Institute for Bioinformatics and Evolutionary Studies, University of Idaho, Moscow, ID 83844, U.S.A.\\
+  $^4$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
   $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. 

### 2014-09-18 Analyses/Figure_PCMCitations.R
Altered y axis label of draft Fig 1

@@ -5,6 +5,7 @@ png("PCMCitations.png")
 # Plot citations
 par(bty = "l")
 plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
-	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2)
+	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2, 
+	 ylab = "Papers", xlab = "Year")
 
 dev.off()

### 2014-09-18 Manuscript/dark.tex
Dealt with Rich's comments

@@
 \citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
+% RF: I think we need a definition of PCMs here, around the second sentence?
+% NC: I can't fit one in here without totally breaking the flow of the paragraph...
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
-PCMs have also become extremely popular over recent years; the number of papers containing the phrase `phylogenetic comparative' has increased exponentially since the 1980s (Figure \ref{PCMCitations}).
+PCMs have also become extremely popular over recent years; the number of papers containing the phrase ``phylogenetic comparative'' has increased exponentially since the 1980s (Figure \ref{PCMCitations}).
 With new methods being published almost weekly, there has never been a better time to be a comparative biologist.
@@
 Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method \citep{freckleton2009seven,boettiger2012your}.
-The more popular these methods become, the less awareness methods-users seem to have of this dark side.
+The more popular these methods become, the less awareness methods-users seem to have of these issues.
 Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results.
@@
 At one extreme, methods-developers (``wookies'') develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
-At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data, and may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
+At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data and, although they have expertise in other biological domains, they may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
 The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users. 
@@
 \citealp{garland1992procedures})
-This time-lag is shorter with more recent methods because simulations testing the method are now required by journals.
-However, we suspect there are still many hidden assumptions and biases in all new PCMs that have yet to be properly explored in the literature. 
-% NC: Rich, we can cite your paper on Pagel 1994 here if you like - as an example of old methods that are still being investigated.
+This time-lag is shorter with more recent methods because simulations testing the method are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions).
+However, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature (for example, see Maddison and FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method).
 
@@
 More recently however, the PCM community has moved towards mostly implementing their methods in R \citep{R-Core-Team:2014aa}, and code sharing has become almost ubiquitous. 
-The number of R packages for PCMs has increased exponentially since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
+The number of R packages for PCMs has increased sharply since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
 Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
@@
 This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
-The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citep{purvis1995comparative,Orme:2013aa}.\\
+The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).\\
 We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology.
 It is extremely frustrating reading/reviewing papers where the authors have clearly failed to carefully read or understand the primary literature on the method they are using.
-Additionally, many methods-users seem to try to retrofit a question so they can use the shiniest new methods available.
+Unfortunately, when methods-users are chastised for making errors it is often just via an assertion that they have used the method ``incorrectly'' and to ``please see the following references''.
+This makes it very difficult for methods-users to fully understand where they went wrong.
+More cynically one could accuse some methods-users of trying to retrofit a question so they can use the newest methods available.
 This is all driven by our ``publish or perish'' culture, and the pressure to produce as many papers as possible out of a dataset.
-The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have made the situation worse rather than better.
+The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have caused additional problems (though on balance, many things have been greatly improved).
 
@@
 Jumping straight into an R implementation is a tolerable way to start a project, but it needs to be followed up by careful reading and thought. 
-Methods-users also need to avoid retrofitting their question to shiny new methods, instead they should think carefully about the question they want to ask, whether the method is appropriate for the question, and whether PCMs are
+Methods-users also need to avoid retrofitting their question to the newest methods, instead they should think carefully about the question they want to ask, whether the method is appropriate for the question, and whether PCMs are
 needed at all \citep{losos2011seeing}.\\
@@
 Improvements to existing methods, and ways of detecting biases or testing model fit should be sufficient for publication. 
-This is fairly standard in other fields, e.g. phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
+This is fairly standard in other fields, e.g. statistical phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
 When novel methods are published, journals should encourage methods-developers to include ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.

### 2015-04-14 Manuscript/dark.tex
Minor edits

@@
 \author{
-  Natalie Cooper$^{1,2,*}$, Matthew W. Pennell$^{3}$ \\
+  Natalie Cooper$^{1,2,*}$, Gavin H. Thomas$^{3}$ \\
   and Richard G. FitzJohn$^{4}$ 
-  %Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\  Gavin H. Thomas$^{5}$, April Wright$^{6}$,\\
+  %Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\  $^{5}$, April Wright$^{6}$,\\
 }
@@
 
- % $^2$ Wisconsin\\
- % $^3$ UBC\\
- % $^5$ Sheffield\\
- % $^6$ Austin\\
   $^1$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
   $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-  $^3$ Institute for Bioinformatics and Evolutionary Studies, University of Idaho, Moscow, ID 83844, U.S.A.\\
+  $^3$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield S10 2TN, UK.\\
   $^4$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
@@
 One major issue is that two groups of people are involved with PCMs. 
-At one extreme, methods-developers (``wookies'') develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
-At the other extreme, na\"{i}ve methods-users (``ewoks'') just want to use PCMs to answer empirical questions with their data and, although they have expertise in other biological domains, they may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on the ``wookie-ewok'' spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
+At one extreme, methods-developers develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
+At the other extreme, na\"{i}ve methods-users just want to use PCMs to answer empirical questions with their data and, although they have expertise in other biological domains, they may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on this spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
 The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users. 
@@
 \end{enumerate}
-
-%Note that we assume that methods-users ignore assumptions and biases because they are unaware of them.
-Of course, there may be methods-users who are perfectly aware of these issues but choose to sweep them under the carpet to get their work published.
-Although we believe there is no excuse for doing this, we also recognize that the ``publish or perish'' culture of academia may pressurize nontenured scientists into this.
-Hopefully some of the changes we suggest below will also help with this problem.
-% NC: This could be left out!
 
@@
 
-Hopefully by trying to implement some of these solutions we can banish the ``Dark Side'' of PCMs and create a ``New Hope'' for the whole PCM community, whether they be wookies, ewoks or somewhere in between.
-
 \section{Funding}

### 2015-04-15 Analyses/Figure_PCMCitations.R
Updated data and analyses for talk and paper

@@ -1,11 +1,15 @@
-ds <- read.csv("PCMCitesThruTime.csv")
+# For talk
 
-png("PCMCitations.png")
+ds <- read.csv("Data/PCMCitesThruTime.csv", )
+
+newred <- rgb(red = 235, green = 81, blue = 83, maxColorValue = 255)
+
+png("Figures/PCMCitationsTalk.png", bg = "transparent")
 
 # Plot citations
 par(bty = "l")
 plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
-	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2, 
+	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2, col = newred,
 	 ylab = "Papers", xlab = "Year")
 
 dev.off()

### 2015-04-15 Analyses/Figure_PackagesForTalk.R

@@ -1,6 +1,6 @@
 library(reshape)
 
-ds <- read.csv("PackagesThruTime.csv")
+ds <- read.csv("Data/PackagesThruTime.csv")
 
 # Group packages by year using aggregate then fix names
 dsYear <- with(ds, aggregate(Program, by = list(YearReleased, RPackage), 
@@ -23,7 +23,7 @@ newred <- rgb(red = 235, green = 81, blue = 83, maxColorValue = 255)
 newblue <- rgb(red = 75, green = 213, blue = 238, maxColorValue = 255)
 newyellow <- rgb(red = 229, green = 177, blue = 58, maxColorValue = 255)
 
-png("PackagesThruTime.png", bg = "transparent")
+png("Figures/PackagesThruTime_Talk.png", bg = "transparent")
 
 # Plot cumulative number of non-R packages through time
 # plot(TotalPkg[R == "other"] ~ YearReleased[R == "other"], 

### 2015-04-15 Analyses/Figure_BrownianAssumptions.R
Tidied up repo

@@ -1,7 +1,7 @@
 library(reshape)
 
-ds <- read.csv("CitesThruTime.csv")
-type <- read.csv("PaperContents.csv")
+ds <- read.csv("Data/CitesThruTime.csv")
+type <- read.csv("Data/PaperContents.csv")
 
 # Reshape the data so it's in long format
 data_long <- melt(ds, id = "Year")
@@ -28,7 +28,7 @@ newred <- rgb(red = 235, green = 81, blue = 83, maxColorValue = 255)
 newblue <- rgb(red = 75, green = 213, blue = 238, maxColorValue = 255)
 newyellow <- rgb(red = 229, green = 177, blue = 58, maxColorValue = 255)
 
-png("BrownianAssumptionCitations.png")
+png("Manuscript/Figures/BrownianAssumptionCitations.png")
 
 par(bg = "black")
 

### 2015-04-15 Analyses/Figure_CitationsPerYear.R

@@ -1,12 +1,12 @@
 library(ggplot2)
 library(reshape)
 
-ds <- read.csv("CitesThruTime.csv")
+ds <- read.csv("Data/CitesThruTime.csv")
 
 # Reshape the data so it's in long format
 data_long <- melt(ds, id = "Year")
 
-pdf("CitationsPerYear_AllPapers")
+pdf("Manuscript/Figures/CitationsPerYear_AllPapers")
 
 # Graph showing how citations change through time
 # for key PCM papers

### 2015-04-15 Analyses/Figure_PCMCitationsForTalk.R

@@ -1,0 +1,15 @@
+# For talk
+
+ds <- read.csv("Data/PCMCitesThruTime.csv", )
+
+newred <- rgb(red = 235, green = 81, blue = 83, maxColorValue = 255)
+
+png("Manuscript/Figures/PCMCitations_Talk.png", bg = "transparent")
+
+# Plot citations
+par(bty = "l")
+plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
+	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2, col = newred,
+	 ylab = "Papers", xlab = "Year")
+
+dev.off()

### 2015-04-15 Analyses/Figure_PackagesForTalk.R

@@ -23,7 +23,7 @@ newred <- rgb(red = 235, green = 81, blue = 83, maxColorValue = 255)
 newblue <- rgb(red = 75, green = 213, blue = 238, maxColorValue = 255)
 newyellow <- rgb(red = 229, green = 177, blue = 58, maxColorValue = 255)
 
-png("Figures/PackagesThruTime_Talk.png", bg = "transparent")
+png("Manuscript/Figures/PackagesThruTime_Talk.png", bg = "transparent")
 
 # Plot cumulative number of non-R packages through time
 # plot(TotalPkg[R == "other"] ~ YearReleased[R == "other"], 

### 2015-04-15 Analyses/Figure_PackagesThruTime.R

@@ -1,7 +1,7 @@
 library(ggplot2)
 library(reshape)
 
-ds <- read.csv("PackagesThruTime.csv")
+ds <- read.csv("Data/PackagesThruTime.csv")
 
 # Group packages by year using aggregate then fix names
 dsYear <- with(ds, aggregate(Program, by = list(YearReleased, RPackage), 
@@ -18,7 +18,7 @@ dsYear <- dsYear[order(dsYear$YearReleased), ]
 dsYear$TotalPkg[dsYear$R == "other"] <- cumsum(dsYear$NPackages[dsYear$R == "other"])
 dsYear$TotalPkg[dsYear$R == "R"] <- cumsum(dsYear$NPackages[dsYear$R == "R"])
 
-# pdf("PackagesPerYear.pdf")
+# pdf("Manuscript/Figures/PackagesPerYear.pdf")
 
 # Plot cumulative number of packages through time
 myplot <- ggplot(data = dsYear,

### 2015-04-15 Analyses/Figure_PCMCitations.R
Tidying repo

@@ -1,15 +1,0 @@
-# For talk
-
-ds <- read.csv("Data/PCMCitesThruTime.csv", )
-
-newred <- rgb(red = 235, green = 81, blue = 83, maxColorValue = 255)
-
-png("Figures/PCMCitationsTalk.png", bg = "transparent")
-
-# Plot citations
-par(bty = "l")
-plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
-	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2, col = newred,
-	 ylab = "Papers", xlab = "Year")
-
-dev.off()

### 2015-04-15 Analyses/Figure_PCMCitations.R
Code to make figure 1 for paper

@@ -1,0 +1,11 @@
+ds <- read.csv("Data/PCMCitesThruTime.csv", )
+
+png("Manuscript/Figures/PCMCitations.png", bg = "transparent")
+
+# Plot citations
+par(bty = "l")
+plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
+	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2,
+	 ylab = "Papers", xlab = "Year")
+
+dev.off()

### 2015-04-15 Analyses/Figure_PCMCitationsForTalk.R
Thicker line for talk fig 1

@@ -9,7 +9,7 @@ png("Manuscript/Figures/PCMCitations_Talk.png", bg = "transparent")
 # Plot citations
 par(bty = "l")
 plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
-	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2, col = newred,
+	 las = 1, lwd = 5, cex.lab = 1.2, cex.axis = 1.2, col = newred,
 	 ylab = "Papers", xlab = "Year")
 
 dev.off()

### 2015-04-15 Analyses/Figure_RPackages.R
Figure 2 with updated data for paper

@@ -1,0 +1,28 @@
+library(reshape)
+
+ds <- read.csv("Data/PackagesThruTime.csv")
+
+# Group packages by year using aggregate then fix names
+dsYear <- with(ds, aggregate(Program, by = list(YearReleased, RPackage), 
+	                          FUN = length))
+
+names(dsYear) <- c("YearReleased", "R", "NPackages")
+
+# Order the data by year so cumulative number of packages
+# per year can be calculated
+dsYear <- dsYear[order(dsYear$YearReleased), ]
+
+# Calculate cumulative number of packages per year
+# separately for non R and R packages
+dsYear$TotalPkg[dsYear$R == "other"] <- cumsum(dsYear$NPackages[dsYear$R == "other"])
+dsYear$TotalPkg[dsYear$R == "R"] <- cumsum(dsYear$NPackages[dsYear$R == "R"])
+
+png("Manuscript/Figures/PackagesThruTime.png")
+
+par(bty = "l")
+plot(TotalPkg[R == "R"] ~ YearReleased[R == "R"], 
+	 data = dsYear, type = "l", ylim = c(0, 80), 
+	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2,
+	 ylab = "Total R packages", xlab = "Year Released", xlim = c(2000,2014))
+
+dev.off()

### 2015-04-17 Manuscript/dark.tex
Changed locations of figures
Small changes

--- 2014-09-18 Manuscript/dark.tex
@@
 \author{
-  Natalie Cooper$^{1,2,*}$, Matthew W. Pennell$^{3}$ \\
-  and Richard G. FitzJohn$^{4}$ 
+  Natalie Cooper$^{1,2,*}$ and Richard G. FitzJohn$^{4}$ 
   %Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\  Gavin H. Thomas$^{5}$, April Wright$^{6}$,\\
@@
   $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-  $^3$ Institute for Bioinformatics and Evolutionary Studies, University of Idaho, Moscow, ID 83844, U.S.A.\\
-  $^4$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
-  $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. 
-       Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+  $^3$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
+  $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
   $^\dag$This paper is based on a Society of Systematic Biologists symposium at Evolution 2014, in Raleigh NC.\\

### 2015-04-29 Manuscript/dark.tex
edits to first paragraph

--- 2015-04-14 Manuscript/dark.tex
@@
 % End of preamble
-
+%----------------------------
 \begin{document}
@@
 
+%----------------------------
 %\section{Introduction} 
@@
 \noindent
-A long time ago in a galaxy far, far away - also known as the 1980s - phylogenetic comparative methods (PCMs) were developed to deal with the statistical non-independence of species in comparative analyses (e.g.
+A long time ago in a galaxy far, far away - also known as the 1980s - phylogenetic comparative methods (PCMs) were first developed. 
+These methods use information on the evolutionary relationships of organisms to compare species \citep{harvey1991comparative}, and were initially designed to deal with the statistical non-independence of species in comparative analyses (e.g.
 \citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
 % RF: I think we need a definition of PCMs here, around the second sentence?
-% NC: I can't fit one in here without totally breaking the flow of the paragraph...
+% NC: I've had a go, see what you think...
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
@@
     \centering
-\includegraphics[width = 10cm]{PCMCitations.png}
+    \includegraphics[width=12cm]{Figures/PCMCitations.png}
     \caption{PCM papers through time}
@@
 \centering
-\includegraphics[width = 10cm]{PCMRpackages.png}
+\includegraphics[width=12cm]{Figures/PackagesThruTime.png}
 \caption{Increase in R packages through time}

### 2015-04-29 Analyses/Figure_PCMCitations.R
Improved resolution of figures

@@ -1,11 +1,12 @@
 ds <- read.csv("Data/PCMCitesThruTime.csv", )
 
-png("Manuscript/Figures/PCMCitations.png", bg = "transparent")
+png("Manuscript/Figures/PCMCitations.png", height = 1200, width = 1200, res = 300)
 
 # Plot citations
 par(bty = "l")
+par(mar = c(4, 4, 2, 2))
 plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
 	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2,
-	 ylab = "Papers", xlab = "Year")
+	 ylab = "Number of papers", xlab = "Year")
 
 dev.off()

### 2015-04-29 Analyses/Figure_PCMRpackages.R

@@ -17,11 +17,14 @@ dsYear <- dsYear[order(dsYear$YearReleased), ]
 dsYear$TotalPkg[dsYear$R == "other"] <- cumsum(dsYear$NPackages[dsYear$R == "other"])
 dsYear$TotalPkg[dsYear$R == "R"] <- cumsum(dsYear$NPackages[dsYear$R == "R"])
 
-png("Manuscript/Figures/PCMRpackages.png")
+png("Manuscript/Figures/PCMRpackages.png", height = 1200, width = 1200, res = 300)
 
 par(bty = "l")
+par(mar = c(4, 4, 2, 2))
+
 plot(TotalPkg[R == "R"] ~ YearReleased[R == "R"], data = dsYear, 
      type = "l" , las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2,
-	 xlim = c(2000,2014), xlab = "Year Released", ylab = "Total Packages")
+	 xlim = c(2000, 2015), ylim = c(0, 80), 
+	 xlab = "Year released", ylab = "Total number of R packages")
 
 dev.off()

### 2015-04-29 Analyses/Figure_PCMCitations.R
Reduced text size on figures

@@ -6,7 +6,6 @@ png("Manuscript/Figures/PCMCitations.png", height = 1200, width = 1200, res = 30
 par(bty = "l")
 par(mar = c(4, 4, 2, 2))
 plot(Citations ~ Year, data = ds, type = "l", ylim = c(0, 800), 
-	 las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2,
-	 ylab = "Number of papers", xlab = "Year")
+	 las = 1, lwd = 2, ylab = "Number of papers", xlab = "Year")
 
 dev.off()

### 2015-04-29 Analyses/Figure_PCMRpackages.R

@@ -23,7 +23,7 @@ par(bty = "l")
 par(mar = c(4, 4, 2, 2))
 
 plot(TotalPkg[R == "R"] ~ YearReleased[R == "R"], data = dsYear, 
-     type = "l" , las = 1, lwd = 2, cex.lab = 1.2, cex.axis = 1.2,
+     type = "l" , las = 1, lwd = 2, 
 	 xlim = c(2000, 2015), ylim = c(0, 80), 
 	 xlab = "Year released", ylab = "Total number of R packages")
 

### 2015-04-29 Manuscript/dark.tex
edits to first paragraph with merge
Tidyied up figures

@@
 \author{
-  Natalie Cooper$^{1,2,*}$, Gavin H. Thomas$^{3}$ \\
-  and Richard G. FitzJohn$^{4}$ 
-  %Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\  $^{5}$, April Wright$^{6}$,\\
+  Natalie Cooper$^{1,2,*}$ and Richard G. FitzJohn$^{4}$ 
+  %Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\  Gavin H. Thomas$^{5}$, April Wright$^{6}$,\\
 }
@@
 
-  \begin{figure}[h]
+  \begin{figure}[!htbp]
     \centering
       \includegraphics[width=12cm]{Figures/PCMCitations.png}
-    \caption{PCM papers through time}
+      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year (Google Scholar search April 13th 2015).}
       \label{PCMCitations}
@@
 
-\begin{figure}[h]
+  \begin{figure}[!htbp]
     \centering
-\includegraphics[width=12cm]{Figures/PackagesThruTime.png}
-\caption{Increase in R packages through time}
+      \includegraphics[width=12cm]{Figures/PCMRpackages.png}
+      \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time.
+Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.}
       \label{PCMRpackages}

### 2015-04-29 Manuscript/dark.tex
Edits to intro

@@
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
-PCMs have also become extremely popular over recent years; the number of papers containing the phrase ``phylogenetic comparative'' has increased exponentially since the 1980s (Figure \ref{PCMCitations}).
+PCMs have also become extremely popular over recent years; the number of papers containing the phrase ``phylogenetic comparative'' has increased dramatically since the 1980s (Figure \ref{PCMCitations}).
 With new methods being published almost weekly, there has never been a better time to be a comparative biologist.
@@
 Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method \citep{freckleton2009seven,boettiger2012your}.
-The more popular these methods become, the less awareness methods-users seem to have of these issues.
+The more popular these methods become, the less awareness many methods-users seem to have of these issues.
 Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results.
@@
 
-One major issue is that two groups of people are involved with PCMs. 
-At one extreme, methods-developers develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
-At the other extreme, na\"{i}ve methods-users just want to use PCMs to answer empirical questions with their data and, although they have expertise in other biological domains, they may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on this spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
-The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users. 
-Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable. 
-Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers.
+We believe that a major cause of this problem is that although people developing and implementing new PCMs are aware of the limitations of their methods and the assumptions that underly them, this information is not being effectively transferred to methods-users.Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable.
+Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers. Here we explore the causes of this communication gap and suggest some potential solutions.
+
+%One major issue is that two groups of people are involved with PCMs. 
+%At one extreme, methods-developers develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
+%At the other extreme, na\"{i}ve methods-users just want to use PCMs to answer empirical questions with their data and, although they have expertise in other biological domains, they may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on this spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
+%The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users. 
+%Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable. 
+%Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers.
 Here we explore the causes of this communication gap and suggest some potential solutions.

### 2015-06-08 Manuscript/dark.tex
Minor edits to intro section

--- 2015-04-29 Manuscript/dark.tex
@@
 \author{
-  Natalie Cooper$^{1,2,*}$ and Richard G. FitzJohn$^{4}$ 
-  %Cecile An\'{e}$^{2}$, Wayne Maddison$^{3}$,\\  Gavin H. Thomas$^{5}$, April Wright$^{6}$,\\
+  Natalie Cooper$^{1,2,*}$, Gavin H. Thomas$^{3}$\\ and Richard G. FitzJohn$^{4}$
 }
@@
 \affiliation{\noindent{\footnotesize
-
   $^1$ School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\ 
-  $^2$ Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-  $^3$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield S10 2TN, UK.\\
+  $^2$ Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\
+  $^3$ Department of Animal and Plant Sciences, University of Sheffield, Sheffield, S10 2TN, UK.\\
   $^4$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
-  $^*$ Corresponding author: ncooper@tcd.ie; Zoology Building, Trinity College Dublin, Dublin 2, Ireland. 
-       Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+  $^*$ Corresponding author: nhcooper12@gmail.com; Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
   $^\dag$This paper is based on a Society of Systematic Biologists symposium at Evolution 2014, in Raleigh NC.\\
@@
 \vfill
-%\paragraph{Word-count:} X words
 
@@
 \citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
-% RF: I think we need a definition of PCMs here, around the second sentence?
-% NC: I've had a go, see what you think...
 Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
-PCMs have also become extremely popular over recent years; the number of papers containing the phrase ``phylogenetic comparative'' has increased exponentially since the 1980s (Figure \ref{PCMCitations}).
+PCMs have also become extremely popular over recent years; the number of papers containing the phrase ``phylogenetic comparative'' has increased dramatically since the 1980s (Figure \ref{PCMCitations}).
 With new methods being published almost weekly, there has never been a better time to be a comparative biologist.
@@
 
-One major issue is that two groups of people are involved with PCMs. 
-At one extreme, methods-developers develop new methods and implement them. They are (generally) aware of the limitations of their methods, and the assumptions that underly them. 
-At the other extreme, na\"{i}ve methods-users just want to use PCMs to answer empirical questions with their data and, although they have expertise in other biological domains, they may not be aware of issues with the methods (of course the majority of people fall somewhere between these two extremes on this spectrum; many methods-developers are developing methods with a specific empirical question in mind, while there are many advanced methods-users who have detailed knowledge of the limitations of the methods they use).
-The problem is that although methods-developers (and advanced methods-users) are aware of the problems with PCMs, this information is not being effectively transferred to methods-users. 
+We believe that a major cause of this problem is that although people developing and implementing new PCMs are aware of the limitations of these methods and the assumptions that underly them, this information is not being effectively transferred to methods-users.
 Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable. 
@@
 Here we explore the causes of this communication gap and suggest some potential solutions.
-% NC: still not happy with this final line or two.
 
@@
 Thanks to the Society of Systematic Biologists for funding our symposium at Evolution 2014. 
-We also thank George Lucas for the ``hilarious'' Star Wars jokes.
 %NC: Should possibly thank Brian O'Meara for funding discourse page

### 2015-06-08 Manuscript/dark.tex
Edits to folklore

@@
     A lot of the information needed to properly apply PCMs is not found in the literature. 
-    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and it is occasionally closer to fiction than fact.
+    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
     Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb. 
     Other times folklore is more opinion-based, but over time these opinions become rules. 
-    One example of such folklore is in the defaults of programs that perform PCMs. 
+    % actually this is a bad example as it is in the literature
+    One example of such folklore is the suggestion in \citet{jones1997optimum} that to avoid outliers overly influencing the results of regressions using phylogenetically independent contrasts, species with phylogenetic residuals $\pm 3$ should be omitted from the analyses.
+% check exact wording
+    The $\pm 3$ cut-off is arbitrary, but has become a rule-of-thumb for running these kinds of analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
+
+    Another example is in the defaults of programs that perform PCMs. 
     These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.
+
     Useful PCM folklore is often shared among methods-developers, and among collaborating groups, but is rarely shared outside of these circles. 
@@
     Other information about the limitations of a method may be absent from the literature due to the time-lag between a new method being published and others having time to test and critique it. 
-For example, Felsenstein published the phylogenetic independent contrasts method in 1985, but it wasn't until the early 1990s that real critiques of the method and its assumptions began to be published (e.g.
-\citealp{garland1992procedures})
+    For example, Felsenstein published the phylogenetic independent contrasts method in 1985, but it was not until the early 1990s that thorough critiques of the method and its assumptions began to be published \citep[e.g.][]{garland1992procedures}.
     This time-lag is shorter with more recent methods because simulations testing the method are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions).
@@
 A big issue for novice methods-users is that this information is often extremely technical and dense.
-It is not unusual for papers to be full of equations, over 20 pages long, and written in a way that gives even advanced users a headache. 
+It is not unusual for papers to be full of equations, \textgreater 20 pages long, and written in a way that gives even advanced users a headache. 
 Additionally there is a lot of literature to wade through before getting all of the required information. 

### 2015-10-06 Manuscript/dark.tex
edits
Formatting for MEE

@@
   $^4$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
-  $^*$ Corresponding author: nhcooper12@gmail.com; Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. Fax: +353 1 677 8094; Tel: +353 1 896 1926.\\
+  $^*$ Corresponding author: natalie.cooper@nhm.ac.uk; Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. Fax: +353 1 677 8094; Tel: +353 1 896 5083.\\
   $^\dag$This paper is based on a Society of Systematic Biologists symposium at Evolution 2014, in Raleigh NC.\\
@@
   \subsection{1. Not everything is mentioned in the literature}
-    A lot of the information needed to properly apply PCMs is not found in the literature. 
+    Some of the information needed to properly apply PCMs is not found in the literature. 
     We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
@@
 The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have caused additional problems (though on balance, many things have been greatly improved).
-
-  \begin{figure}[!htbp]
-    \centering
-      \includegraphics[width=12cm]{Figures/PCMRpackages.png}
-      \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time.
-Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.}
-      \label{PCMRpackages}
-  \end{figure}
 
@@
 
-\bibliographystyle{sysbio}
+\bibliographystyle{mee}
 \bibliography{darkside}
 
+\newpage
+\section{Tables}
+
+\newpage
+\section{Figures}
+
+  \begin{figure}[!htbp]
+    \centering
+      \includegraphics[width=12cm]{Figures/PCMCitations.png}
+      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year (Google Scholar search April 13th 2015).}
+      \label{PCMCitations}
+  \end{figure}
+
+\newpage
+  \begin{figure}[!htbp]
+    \centering
+      \includegraphics[width=12cm]{Figures/PCMRpackages.png}
+      \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time.
+Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.}
+      \label{PCMRpackages}
+  \end{figure}
+
 \end{document}

### 2015-10-06 Manuscript/dark.tex
Formatting for MEE

@@
 \pagenumbering{arabic}    
-
-% From Syst Biol template
 \linespread{1.66}
-
-\renewcommand{\section}[1]{
-  \bigskip
-  \begin{center}
-  \begin{Large}
-  \normalfont\scshape #1
-  \medskip
-  \end{Large}
-  \end{center}
-}
-
-\renewcommand{\subsection}[1]{
-  \bigskip
-  \begin{center}
-  \begin{large}
-  \normalfont\itshape #1
-  \end{large}
-  \end{center}
-}
-
-\renewcommand{\subsubsection}[1]{%
-\vspace{2ex}
-\noindent
-\textit{#1.}---}
-
-\renewcommand{\tableofcontents}{}
-
-\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
-
-% NC: Ignoring their title page setup as I like Rich's template better :)
 
@@
 
-\title{The ``Dark Side'' of Phylogenetic Comparative Methods$^\dag$}
+\title{The ``Dark Side'' of Phylogenetic Comparative Methods}
 \author{
@@
   $^*$ Corresponding author: natalie.cooper@nhm.ac.uk; Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. Fax: +353 1 677 8094; Tel: +353 1 896 5083.\\
-  $^\dag$This paper is based on a Society of Systematic Biologists symposium at Evolution 2014, in Raleigh NC.\\
 }}
@@
 \runninghead{The ``Dark Side'' of PCMs}
-\keywords{PCM, assumption, etc.}
+\keywords{PCM, assumption,}
 
 % End of preamble
-%----------------------------
+
 \begin{document}
@@
 
-%\section{abstract}
-% NC: If Syst Biol "Point of View" there's no abstract
+% Abstract
+\section{Abstract}
+
+% Need to call it summary
 
@@
 %----------------------------
-%\section{Introduction} 
-% NC: No intro heading for Syst Biol
+\section{Introduction}
 
-\noindent
 A long time ago in a galaxy far, far away - also known as the 1980s - phylogenetic comparative methods (PCMs) were first developed. 
@@
 
-\section{Funding}
+\section{Acknowledgments}
+
+Thanks to the Society of Systematic Biologists for funding the symposium at Evolution 2014, in Raleigh, NC, where the ideas in this paper were first consolidated.
 This work was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696 (NC).
-
-\section{Acknowledgments}
-Thanks to the Society of Systematic Biologists for funding our symposium at Evolution 2014. 
-%NC: Should possibly thank Brian O'Meara for funding discourse page
 

### 2015-10-09 Manuscript/dark.tex
Add examples to manuscript

@@
 % Title page information
-
 \title{The ``Dark Side'' of Phylogenetic Comparative Methods}
@@
 
-\runninghead{The ``Dark Side'' of PCMs}
-\keywords{PCM, assumption,}
+\runninghead{Limitations of PCMs}
+\keywords{PCM, assumption,\\
+\textbf{Word count}: XXX (including references, tables and figure legends)}
 
@@
 \mstitlepage
+
 \parindent = 1.5em
@@
 
-% Abstract
-\section{Abstract}
+% Abstract 350 words max
+\abstract
 
-% Need to call it summary
+Point 1: set the context and purpose for the work; 
+Point 2: indicate the approach and methods used; 
+Point 3: outline the main results; 
+Point 4: identify the conclusions, the wider implications and the relevance to management or policy.
 
@@
 \citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
-Since then PCMs have been extended to investigate evolutionary pattern and process, and include methods for investigating drivers of diversification, the tempo and mode of trait evolution, and models of speciation and extinction (see reviews in \citealp{o2012evolutionary, pennell2013integrative}).
+Since then PCMs have been extended to investigate evolutionary pattern and process (see reviews in \citealp{o2012evolutionary, pennell2013integrative}), and include methods for investigating drivers of diversification \citep[e.g.][]{maddison2007estimating}, the tempo and mode of trait evolution \citep[e.g.][]{o2012evolutionary}, and models of speciation and extinction \citep[e.g.][]{nee1994extinction}.
 PCMs have also become extremely popular over recent years; the number of papers containing the phrase ``phylogenetic comparative'' has increased dramatically since the 1980s (Figure \ref{PCMCitations}).
@@
 
-  \begin{figure}[!htbp]
-    \centering
-      \includegraphics[width=12cm]{Figures/PCMCitations.png}
-      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year (Google Scholar search April 13th 2015).}
-      \label{PCMCitations}
-  \end{figure}
+Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method, and this is well established in the literature \citep[e.g.][]{freckleton2009seven,boettiger2012your,losos2011seeing,blomberg2012independent}.
+Increasingly however, assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results (see examples below). % ADD EXAMPLES
+Additionally, little consideration is given to whether using a PCM is really appropriate for the question at hand \citep{losos2011seeing,westoby1995misinterpreting}.
 
-Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method \citep{freckleton2009seven,boettiger2012your}.
-The more popular these methods become, the less awareness many methods-users seem to have of these issues.
-Increasingly assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results.
-Additionally, little consideration is given to whether using a PCM is really appropriate for the question at hand \citep{losos2011seeing}. 
-To paraphrase \citet{blomberg2012independent}, the enthusiastic application of PCMs has greatly outpaced our theoretical understanding of the methods and their relations to the rest of statistical theory.\\
-
-% NC: Do we need more explicit comments showing we're not the only people to think this, nor the first to say it?
-
-We believe that a major cause of this problem is that although people developing and implementing new PCMs are aware of the limitations of these methods and the assumptions that underly them, this information is not being effectively transferred to methods-users.
+We suggest that one cause of this problem is that although people developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to methods-users.
 Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable. 
@@
 Here we explore the causes of this communication gap and suggest some potential solutions.
+Note that many of these issues are applicable across methods in ecology and evolution, not just PCMs.
 
-\section{What impedes information transfer from methods-developers to methods-users?}
-  % NC: Or something similar
+\subsection{Examples of the problem}
+\subsubsection{Trait-dependent diversification}
+
+Analyses of trait-dependent diversification are used to detect whether particular traits promote high rates of diversification, leading to some clades becoming more diverse than others. 
+These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{}. 
+Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSEE paper \citealp{maddison2007estimating} has been cited 384 times; Google Scholar search 9th October 2015)
+However, this method has several caveats. 
+Rabosky and Goldberg BUT all mentioned in Maddison and FitzJohn papers. Not unknown but definitely misused.
+
+a strong correlation
+between character states and diversification can be
+inferred from merely a single diversification rate
+shift within a phylogenetic tree, even if the shift is
+unrelated to the modeled character
+
+This illustrates one
+form of model inadequacy: speciation rate heterogeneity
+tied to a character not under study can drive
+mistaken inference of state-dependent diversification
+
+\subsubsection{Ornstein-Uhlenbeck models of trait evolution}
+Most models of trait evolution are based on the Brownian constant variance model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
+The Ornstein-Uhlenbeck (OU) model is a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{Butler:2004aa,hansen1997stabilizing}.
+OU models have become increasingly popular as they tend to fit the data better than Brownian motion models, and have attractive biological interpretations \citep{cooper2016}.
+For example, fit to an OU model has been seen as evidence of evolutionary constraints, stabilizing selection, niche conservatism, and selective regimes \citep{}.
+However, the OU model has several well-known caveats \citep[see][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your}. 
+First, it is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, particularly for small datasets that are commonly used in these analyses (the median number of taxa used for OU studies is 58; \citealp{cooper2016}).
+Second, very small amounts of error in datasets can result in an OU model being favoured over Brownian motion simply because OU can accommodate more variance towards the tips of the phylogeny, rather than due to any interesting biological process \citep{boettiger2012your,pennell2015model}.
+Finally, the literature describing the OU model is clear that a simple explanation of clade-wide stabilizing selection is unlikely to account for data fitting an OU model \citep[e.g.][]{hansen1997stabilizing,hansen2005assessing}, but implementations of the model often state that this is the case.
+Unfortunately, these limitations are rarely taken into account in empirical studies.
+
+\subsubsection{Phylogenetically independent contrasts}
+Even the most commonly used phylogenetic comparative method, phylogentically independent contrasts has weel-known assumptions that are rarely tested.
+
+\subsection{What impedes information transfer about the limitations of phylogenetic comparative  methods?}
+These three brief examples illustrate that although the field as a whole is aware of the limitations of PCMs, this information is not filtering through to everyone using the methods. 
+Why might this be the case?
+
+  \subsubsection{1. Not everything is mentioned in the literature}
     As scientists we mainly communicate our ideas through the literature. 
-  Below we propose three possible reasons for difficulties transferring information between methods-developers and methods-users via the literature.
+    Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
+    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
+    Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb, other times it is more opinion-based, but over time these opinions become rules. 
+    Useful PCM folklore is often shared among methods-developers, and among collaborating groups, but is not always shared outside of these circles. 
+    When it is shared, it tends to be in the form of email exchanges that form a kind of ``dark advice'' that is not accessible to the rest of the community.
 
-    \begin{enumerate}
-      \item Not everything is mentioned in the literature.
-      \item The literature is too technical and/or important details are hidden within the text.
-      \item Methods-users have bypassed the literature and gone straight to the implementation of the method.
-    \end{enumerate}
-
-  \subsection{1. Not everything is mentioned in the literature}
-    Some of the information needed to properly apply PCMs is not found in the literature. 
-    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
-    Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb. 
-    Other times folklore is more opinion-based, but over time these opinions become rules. 
-    % actually this is a bad example as it is in the literature
-    One example of such folklore is the suggestion in \citet{jones1997optimum} that to avoid outliers overly influencing the results of regressions using phylogenetically independent contrasts, species with phylogenetic residuals $\pm 3$ should be omitted from the analyses.
+    One example of PCM folklore is the suggestion in \citet{jones1997optimum} that to avoid outliers overly influencing the results of regressions using phylogenetically independent contrasts, species with phylogenetic residuals $\pm 3$ should be omitted from the analyses.
 % check exact wording
     The $\pm 3$ cut-off is arbitrary, but has become a rule-of-thumb for running these kinds of analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
-
     Another example is in the defaults of programs that perform PCMs. 
     These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.
+    For example, % Anyone got a good example here?
 
-    Useful PCM folklore is often shared among methods-developers, and among collaborating groups, but is rarely shared outside of these circles. 
-    When it is shared, it tends to be in the form of email exchanges between methods-developers and struggling methods-users.
-    This advice is extremely useful to the methods-user in question, but forms a kind of ``dark advice'' that isn't accessible to the rest of the community.\\
+    Other information about the limitations of a method may be absent from the literature due to the time-lag between a new method being published and others having time to test it. 
+    For example, Felsenstein published the phylogenetic independent contrasts method in 1985 \citep{felsenstein1985phylogenies}, but it was not until the early 1990s that thorough critiques of the method and its assumptions began to be published \citep[e.g.][]{garland1992procedures}.
+    This time-lag is shorter with more recent methods because simulations testing the method are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions).
+    However, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature
+    For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation modelss \citep{maddison2007estimating,nee1994reconstructed}.
 
-% NC: Do we want a specific example here? Theta estimation in OU models? 
-% NC: Could mention ±3 from Jones&Purvis97 - totally unjustified rule of thumb now = canon.
-% NC: Mention hidden assumptions? I mention them later
-
-    Other information about the limitations of a method may be absent from the literature due to the time-lag between a new method being published and others having time to test and critique it. 
-    For example, Felsenstein published the phylogenetic independent contrasts method in 1985, but it was not until the early 1990s that thorough critiques of the method and its assumptions began to be published \citep[e.g.][]{garland1992procedures}.
-    This time-lag is shorter with more recent methods because simulations testing the method are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions).
-    However, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature (for example, see Maddison and FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method).
-
-\subsection{2. The literature is too technical and/or important details are hidden}
-Although a lot of information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
-A big issue for novice methods-users is that this information is often extremely technical and dense.
-It is not unusual for papers to be full of equations, \textgreater 20 pages long, and written in a way that gives even advanced users a headache. 
+  \subsection{2. The literature is too technical and/or important details are difficult to locate}
+    Although some information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
+    A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
+    It is not unusual for papers to be full of equations, \textgreater 20 pages long, and written in a way that gives one a headache. 
     Additionally there is a lot of literature to wade through before getting all of the required information. 
@@
     Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. % NC: I will add numbers at some point!
-This seems like an excessive amount of reading to expect from a methods-user trying to run a simple empirical analysis.\\
-
-Another issue is that it can sometimes be hard to find the information required. Assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
-They are rarely neatly corralled in one place, making it really easy to miss pertinent details. 
-This places an additional burden on methods-users who must read every paper very carefully.
+    This seems like an excessive amount of reading to expect from someone trying to run a simple empirical analysis.
+    Another issue is that it can sometimes be hard to find the information required. 
+    Assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
+    They are rarely neatly corralled in one place, making it easy to miss pertinent details.
 
@@
 
-More recently however, the PCM community has moved towards mostly implementing their methods in R \citep{R-Core-Team:2014aa}, and code sharing has become almost ubiquitous. 
+    More recently however, the community has moved towards mostly implementing methods in R \citep{R-Core-Team:2014aa}, and code sharing has become almost ubiquitous. 
     The number of R packages for PCMs has increased sharply since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
-Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
-\citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve the ability of methods-users to implement methods correctly. 
-However, this has instead led to an increasing number of methods-users jumping straight to the implementation of a method, without understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
+    Simultaneously, more people are able to use R, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
+    \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
+    However, this has instead led to an increasing number of people (myself included) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
     This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
     The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).\\
-We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology.
+
+    %We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology.
 It is extremely frustrating reading/reviewing papers where the authors have clearly failed to carefully read or understand the primary literature on the method they are using.
-Unfortunately, when methods-users are chastised for making errors it is often just via an assertion that they have used the method ``incorrectly'' and to ``please see the following references''.
-This makes it very difficult for methods-users to fully understand where they went wrong.
-More cynically one could accuse some methods-users of trying to retrofit a question so they can use the newest methods available.
+    %Unfortunately, when authors are chastised for making errors it is often just via an assertion that they have used the method ``incorrectly'' and to ``please see the following references''.
+    %This makes it very difficult to completely understand where they went wrong.
+    %More cynically one could accuse some methods-users of trying to retrofit a question so they can use the newest methods available.
 This is all driven by our ``publish or perish'' culture, and the pressure to produce as many papers as possible out of a dataset.
-The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have caused additional problems (though on balance, many things have been greatly improved).
+    %The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have caused additional problems (though on balance, many things have been greatly improved).
 
-\section{How can we solve some of these problems?}
+\subsection{How can we solve some of these problems?}
 
-\subsection{1. Simplify, summarize and share}
+\subsubsection{1. Simplify, summarize and share}
 Many of the problems above are caused by methods papers being difficult to read. 
@@
 
-\subsection{2. Read, think and learn}
+\subsubsection{2. Read, think and learn}
 If methods-developers are going to put time and effort into making their methods easier to understand, then methods-users need to take advantage of this and work harder to understand the methods they are using.
@@
 
-\subsection{3. Collaboration}
+\subsubsection{3. Collaboration}
 A obvious solution to all these problems is for methods-developers and methods-users to collaborate more. 
@@
 
-\subsection{4. Incentives}
+\subsubsection{4. Incentives}
 It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
@@
 
+\section{Data Accessibility}
+All data are available on GitHub at https://github.com/nhcooper123/pcm-darkside.
+
 \bibliographystyle{mee}
 \bibliography{darkside}
-
-\newpage
-\section{Tables}
 
@@
     \centering
-      \includegraphics[width=12cm]{Figures/PCMCitations.png}
-      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year (Google Scholar search April 13th 2015).}
+      \includegraphics[width=12cm]{Figures/PCMCitations.pdf}
+      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).}
       \label{PCMCitations}
@@
     \centering
-      \includegraphics[width=12cm]{Figures/PCMRpackages.png}
-      \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time.
+      %\includegraphics[width=12cm]{Figures/PCMRpackages.pdf}
+      \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time from 1980-2014.
 Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.}

### 2015-10-09 Manuscript/dark.tex
Play with solutions

@@
 
-Point 1: set the context and purpose for the work; 
-Point 2: indicate the approach and methods used; 
-Point 3: outline the main results; 
-Point 4: identify the conclusions, the wider implications and the relevance to management or policy.
+Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
+However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
+Unfortunately, although these limitations are often well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
+Here we explore reasons for this communication gap between those developing new methods and those using them. 
+We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
+We also highlight problems with users jumping straight into implementations (e.g. in R) that may lack documentation of biases and assumptions that are mentioned in the original papers.
+To help solve these problems we propose...
 
@@
 Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method, and this is well established in the literature \citep[e.g.][]{freckleton2009seven,boettiger2012your,losos2011seeing,blomberg2012independent}.
-Increasingly however, assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results (see examples below). % ADD EXAMPLES
+Increasingly however, assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results (see examples below).
 Additionally, little consideration is given to whether using a PCM is really appropriate for the question at hand \citep{losos2011seeing,westoby1995misinterpreting}.
@@
 \subsection{Examples of the problem}
-\subsubsection{Trait-dependent diversification}
+\subsubsection{Trait-dependent diversification} % FINISH and get GAVIN TO CHECK!!!
 
@@
 
-\subsubsection{Phylogenetically independent contrasts}
+\subsubsection{Phylogenetically independent contrasts} % FINISH!!!
 Even the most commonly used phylogenetic comparative method, phylogentically independent contrasts has weel-known assumptions that are rarely tested.
@@
 \subsection{How can we solve some of these problems?}
-
   \subsubsection{1. Simplify, summarize and share}
     Many of the problems above are caused by methods papers being difficult to read. 
-This isn't entirely the fault of methods-developers; most journals have strict word limits so it's often easier to use a one line equation rather than a paragraph of description. 
-However, it's hard to complain about novice users failing to understand a method when the paper is incomprehensible. %cough cough Hansen. Ow my eyes are bleeding.
+    This is not entirely the fault of those writing these papers; most journals have strict word limits so it is often easier to use a one line equation rather than a paragraph of description. 
+    However, it is hard to complain about novice users failing to understand a method when the paper is incomprehensible.
     One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 
-Some journals already encourage this (e.g. Methods in Ecology and Evolution), and even if they do not, there is no harm in doing this and hosting it on a personal website. 
-The ability to share ideas with a non-technical audience is a key skill to develop, and it may even encourage more people to use the method. 
-% NC: Perhaps add something about having reproducible knitr reports in supp mat (e.g. cite that nice recent paper)
-At the very least, having a simple list of the main assumptions, biases and caveats of the method somewhere would reduce misuse and provide a place to point methods-users to when they begin using a method.
-Of course this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible.\\
+    Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there is no harm in doing this and hosting it on a personal website.
+    The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
+    Encouraging increased efforts at reproducibility would also help, by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
+    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).\\
 
     The glut of literature that already exists for established methods is a bigger problem. 
-One solution would be to establish a wiki-style website where methods-developers and advanced methods-users could post summaries of commonly used methods, along with lists of key papers to read. 
+    One solution would be to establish a wiki-style website where people could post summaries of commonly used methods, along with lists of key papers to read. 
     The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
     Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) is an excellent example of a similar approach for phylogenetics. 
-Unfortunately there is currently little incentive for methods-developers and advanced methods-users to contribute to such a project at the cost of time pursuing their own research.\\
+    The British Ecological Society's Quantitative Ecology Special Interest Group is also in the process of creating a ``Field Guide for Ecologists'' (http://bes-qsig.github.io/fge/) that will fill a similar niche in ecological methods.
+    Unfortunately there is currently little incentive for people to contribute to such a project at the cost of time pursuing their own research.\\
 
-One solution is to establish a discourse page for providing R help. 
-Currently most methods-users either email the R-sig-phylo help list or the package developer for assistance. 
-Both approaches are suboptimal. 
-R-sig-phylo is impossible to search effectively, leading to the same questions being dealt with repeatedly. 
-The same is true in email exchanges between methods-users and package developers with the added issue of this being ``dark advice'' that isn't shared. 
-A discourse page would collate all of this information and reduce pressure on package developers by allowing advanced methods-users to help novice users with simple queries. 
-We are in the process of setting up a discourse page for this purpose (it will be called ``phyloBootstrap'' because it will provide ``support'' for PCMs).
+%One solution is to establish a discourse page for providing R help. 
+%Currently most methods-users either email the R-sig-phylo help list or the package developer for assistance. 
+%Both approaches are suboptimal. 
+%R-sig-phylo is impossible to search effectively, leading to the same questions being dealt with repeatedly. 
+%The same is true in email exchanges between methods-users and package developers with the added issue of this being ``dark advice'' that isn't shared. 
+%A discourse page would collate all of this information and reduce pressure on package developers by allowing advanced methods-users to help novice users with simple queries. 
+%We are in the process of setting up a discourse page for this purpose (it will be called ``phyloBootstrap'' because it will provide ``support'' for PCMs).
 At Evolution 2014 many methods-developers agreed to move their R package help to this forum.
 
-\subsubsection{2. Read, think and learn}
-If methods-developers are going to put time and effort into making their methods easier to understand, then methods-users need to take advantage of this and work harder to understand the methods they are using.
-At a minimum, methods-users should read the original papers describing a method, and any recent updates. 
-Jumping straight into an R implementation is a tolerable way to start a project, but it needs to be followed up by careful reading and thought. 
-Methods-users also need to avoid retrofitting their question to the newest methods, instead they should think carefully about the question they want to ask, whether the method is appropriate for the question, and whether PCMs are
-needed at all \citep{losos2011seeing}.\\
+  \subsubsection{2. Be more cynical about the methods you use}
+    A key skill to develop in science is cynicism - never take results from PCMs (or any other statistical analysis) at face value.
+    At a minimum, it is worth reading the original papers describing a method, plus any recent updates, and looking carefully for assumptions and caveats that may affect the analyses at hand.
+    A good way to check a method is to simulate some data and see if the results are as expected \citep[e.g.][]{boettiger2012your}. 
+    This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
+    It is also important to consider whether the method will work on a particular dataset for example.
+    \begin{itemize}
+      \item How many species does it need to have reasonable power? 
+      Often methods require more species than are generally available. 
+      For example, the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily be applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most such analyses have \textless 1000 tips.
+      \item How does it deal with polytomies?
+      Many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution \citep{cooper2010body}.
+      \item Can the method use non-ultrametric trees? 
+      Not all PCMs can use non-ultrametric trees. 
+      For example, some current implementations of the OU model should not be used with non-ultrametric trees because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix (e.g.
+MOTMOT; \citealp{Thomas:2011aa}), and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citep{slater2014correction}.
+    \end{itemize}
+    It is also important to avoid retrofitting questions to the newest methods, instead think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{losos2011seeing}.
+    Finally, methods-users should never be afraid to question standard practice, sometimes it's just PCM folklore!
 
-Another key skill to develop is distrust; never take results from PCMs (or any other statistical analysis) at face value. 
-A good way to check a method is to simulate some data and see if the results are as expected. 
-This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
-It is also important for methods-users to consider whether the method will work on their dataset; for example, how many species does it need to have reasonable power, how does it deal with polytomies, can it cope with non-ultrametric trees etc.
-Finally, methods-users should never be afraid to question standard practice, sometimes it's just PCM folklore!
-% NC: Could add my pglm vs pgls example.
+%\subsubsection{3. Collaboration}
+%A obvious solution to all these problems is for methods-developers and methods-users to collaborate more. 
+%The benefits to methods-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
+%A big difficulty with collaboration is that it is often asymmetrical in terms of benefit to methods-users and methods-developers. 
+%Methods-developers and advanced methods-users can often put a lot of time and effort into running analyses for less experienced colleagues, yet at best they come out of the experience with a middle authored paper.
+%Until these author positions are considered equal to first and last author positions, there is little incentive to collaborate extensively.
 
-\subsubsection{3. Collaboration}
-A obvious solution to all these problems is for methods-developers and methods-users to collaborate more. 
-The benefits to methods-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
-A big difficulty with collaboration is that it is often asymmetrical in terms of benefit to methods-users and methods-developers. 
-Methods-developers and advanced methods-users can often put a lot of time and effort into running analyses for less experienced colleagues, yet at best they come out of the experience with a middle authored paper.
-Until these author positions are considered equal to first and last author positions, there is little incentive to collaborate extensively.
+  \subsubsection{3. Incentives}
+    It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
+    We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for such skilled researchers to remain in (or at least interact with) academia.
 
-\subsubsection{4. Incentives}
-It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
-We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for methods-developers to remain in (or at least interact with) academia.
-
-First we need to stop insisting that methods papers must be entirely novel. 
+    First we need to stop insisting that methods papers are entirely novel. 
     Improvements to existing methods, and ways of detecting biases or testing model fit should be sufficient for publication.
     This is fairly standard in other fields, e.g. statistical phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
-When novel methods are published, journals should encourage methods-developers to include ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.
+    When novel methods are published, journals should encourage researchers to include ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.
 
@@
     Currently it is difficult to get funding for purely methods driven research; an empirical component is always needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
-This unfairly prejudices funding bodies (and hiring committees) away from methods-developers.
+    This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.
 
-Finally we need to re-evaluate our criteria for success in academia. 
-The ``publish or perish'' mentality does not encourage methods-developers to spend more time on their methods than necessary. 
-Thus there is little incentive to summarize, simplify and share, or to collaborate with methods-users, as we suggest above. 
-This mentality is also not conducive to methods-users attempting to fully understand a method prior to using it. 
-Instead we need to value quality over quantity when it comes to publications.
-This would improve career prospects for both methods-developers and careful methods-users.
+    Finally, an obvious solution to many of these problems is for methods-developers and methods-users to collaborate more. 
+    The benefits to methods-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
+    A big difficulty with collaboration is that it is often asymmetrical in terms of benefit; some researchers can put a lot of time and effort into running analyses for less experienced colleagues, yet at best they come out of the experience with a middle authored paper.
+    Until these author positions are considered equal to first and last author positions, there is little incentive to collaborate extensively.
 
-\section{Conclusion}
-% NC: Still need a couple of lines of conclusion but I'm drawing a blank. Final sentence is fine I think. We could just skip the conclusion heading and have this as a standalone sentence at the end.
+  \subsection{Conclusion}
 
-\section{Acknowledgments}
+%I guess I’d close by saying it’s our responsbilitiy as part of the methods developing and using community to find innovative ways to tackle these challenges.
 
-Thanks to the Society of Systematic Biologists for funding the symposium at Evolution 2014, in Raleigh, NC, where the ideas in this paper were first consolidated.
+\subsection{Acknowledgments}
+This paper is based on a talk given by NC at the Methods in Ecology and Evolution 5th Anniversary Symposium.
+Thanks to the Society of Systematic Biologists for funding the symposium at Evolution 2014 where the ideas in this paper were first consolidated. 
 This work was supported by The European Commission CORDIS Seventh Framework Program (FP7) Marie Curie CIG grant, proposal number: 321696 (NC).
 
-\section{Data Accessibility}
+\subsection{Data Accessibility}
 All data are available on GitHub at https://github.com/nhcooper123/pcm-darkside.

### 2015-10-11 Manuscript/dark.tex
Added details to all but examples section

@@
 % Abstract 350 words max
-\abstract
+\section{Abstract}
 
@@
 We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
-We also highlight problems with users jumping straight into implementations (e.g. in R) that may lack documentation of biases and assumptions that are mentioned in the original papers.
-To help solve these problems we propose...
+We also highlight problems with users jumping straight into software implementations of the methods (e.g.
+in R) that may lack documentation of biases and assumptions that are mentioned in the original papers.
+To help solve these problems we make a number of suggestions including providing
+blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, incentives for collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
+Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
 
@@
 
-Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method, and this is well established in the literature \citep[e.g.][]{freckleton2009seven,boettiger2012your,losos2011seeing,blomberg2012independent}.
+Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method - a fact that is well established in the literature \citep[e.g.][]{freckleton2009seven,boettiger2012your,losos2011seeing,blomberg2012independent}.
 Increasingly however, assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results (see examples below).
@@
 
-We suggest that one cause of this problem is that although people developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to methods-users.
+We suggest that one cause of this problem is that although researchers developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to end-users.
 Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable. 
-Clearly more effort is needed to bridge the widening gap between methods-users and methods-developers. 
+Clearly more effort is needed to bridge the widening gap between those developing methods and end-users. 
 Here we explore the causes of this communication gap and suggest some potential solutions.
@@
 
-\subsection{What impedes information transfer about the limitations of phylogenetic comparative  methods?}
-These three brief examples illustrate that although the field as a whole is aware of the limitations of PCMs, this information is not filtering through to everyone using the methods. 
+\subsection{What impedes information transfer about the limitations of PCMs?}
+These three brief examples illustrate that although the PCM community is aware of the limitations of PCMs, this information is not filtering through to everyone using the methods. 
 Why might this be the case?
@@
     We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
-    Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb, other times it is more opinion-based, but over time these opinions become rules. 
-    Useful PCM folklore is often shared among methods-developers, and among collaborating groups, but is not always shared outside of these circles. 
-    When it is shared, it tends to be in the form of email exchanges that form a kind of ``dark advice'' that is not accessible to the rest of the community.
+    Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules. 
+    Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
+    When it is shared, it tends to be as email exchanges of ``dark advice'' that is not accessible to the rest of the community.
 
-    One example of PCM folklore is the suggestion in \citet{jones1997optimum} that to avoid outliers overly influencing the results of regressions using phylogenetically independent contrasts, species with phylogenetic residuals $\pm 3$ should be omitted from the analyses.
-% check exact wording
-    The $\pm 3$ cut-off is arbitrary, but has become a rule-of-thumb for running these kinds of analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
+    One example of PCM folklore is that species with Studentized residuals $\pm 3$ are often omitted from regressions of phylogenetic independent contrasts, to avoid highly influential outliers affecting the results \citep[e.g.][]{GEB:GEB355}.
+    The rationale for this comes from \citet{jones1997optimum}, however, the $\pm 3$ cut-off is arbitrary, and barely mentioned in the original paper, but has become a rule-of-thumb for running these analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
     Another example is in the defaults of programs that perform PCMs. 
@@
     This time-lag is shorter with more recent methods because simulations testing the method are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions).
-    However, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature
-    For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation modelss \citep{maddison2007estimating,nee1994reconstructed}.
+    However, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
+    For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation models \citep{maddison2007estimating,nee1994reconstructed}.
 
@@
     Additionally there is a lot of literature to wade through before getting all of the required information. 
-    For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies}.
-    Even with prior knowledge of the key papers and authors to focus on, this resulted in XX pages, XX papers and a book to read. % NC: I will add numbers at some point!
-    This seems like an excessive amount of reading to expect from someone trying to run a simple empirical analysis.
+    For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies} (including how they are related to phylogenetic generalised least squares models; \citealp{blomberg2012independent,rohle2006comment}).
+    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ pages, 23 papers and a book to read.
+    This seems like an excessive amount of reading to run a simple empirical analysis.
     Another issue is that it can sometimes be hard to find the information required. 
@@
   \subsection{3. Users jump straight to the implementation of the method}
-    In the early days of PCMs, some methods-developers provided stand alone packages to run their methods, others provided code in whichever language they were comfortable programming in, and still others provided no way of implementing their methods at all.
+    In the early days of PCMs, some researchers provided stand alone packages to run their methods (e.g.
+PDAP; \citealp{diaz1996testing}), others provided code in whichever language they chose to program in (e.g.
+MATLAB code; \citealp{rohlf2001comparative}), and still others provided no way of implementing their methods at all.
     This resulted in many frustrating hours (and days and months!) trying to implement any new method you wanted to use.\\
 
-    More recently however, the community has moved towards mostly implementing methods in R \citep{R-Core-Team:2014aa}, and code sharing has become almost ubiquitous. 
+    More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' http://ftp.iitm.ac.in/cran/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous.
     The number of R packages for PCMs has increased sharply since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
-    Simultaneously, more people are able to use R, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
+    Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
     \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
-    However, this has instead led to an increasing number of people (myself included) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
+    However, this has instead led to an increasing number of people (including the authors) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
     This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
@@
     Encouraging increased efforts at reproducibility would also help, by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
-    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).\\
+    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).\\
 
     The glut of literature that already exists for established methods is a bigger problem.
+    The Oxford Bibliographies Evolutionary biology pages have lists of key papers (http://www.oxfordbibliographies.com/obo/page/evolutionary-biology), but these are curated by just a few individuals and tend to contain a lot of papers.
     One solution would be to establish a wiki-style website where people could post summaries of commonly used methods, along with lists of key papers to read. 
@@
 
-  \subsubsection{2. Be more cynical about the methods you use}
-    A key skill to develop in science is cynicism - never take results from PCMs (or any other statistical analysis) at face value.
+  \subsubsection{2. Be more cynical about methods}
+    A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
     At a minimum, it is worth reading the original papers describing a method, plus any recent updates, and looking carefully for assumptions and caveats that may affect the analyses at hand.
@@
     This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
-    It is also important to consider whether the method will work on a particular dataset for example.
-    \begin{itemize}
-      \item How many species does it need to have reasonable power? 
-      Often methods require more species than are generally available. 
-      For example, the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily be applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most such analyses have \textless 1000 tips.
-      \item How does it deal with polytomies?
-      Many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution \citep{cooper2010body}.
-      \item Can the method use non-ultrametric trees? 
-      Not all PCMs can use non-ultrametric trees. 
-      For example, some current implementations of the OU model should not be used with non-ultrametric trees because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix (e.g.
-MOTMOT; \citealp{Thomas:2011aa}), and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citep{slater2014correction}.
-    \end{itemize}
-    It is also important to avoid retrofitting questions to the newest methods, instead think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{losos2011seeing}.
-    Finally, methods-users should never be afraid to question standard practice, sometimes it's just PCM folklore!
+    It is also important to consider whether the method will work on a particular dataset.
+    For example it is worth considering how many species are required for reasonable power (often methods require more species than are usually available, e.g.
+the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily be applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most empirical analyses have \textless 1000 tips), whether the method is influenced by polytomies (many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution \citealp{cooper2010body}), and if the method can use non-ultrametric trees (some current implementations of the OU model should not be used with non-ultrametric trees because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix, e.g.
+MOTMOT; \citealp{Thomas:2011aa}, and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citealp{slater2014correction}).
+    It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{losos2011seeing,westoby1995misinterpreting}.
+    Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore!
 
@@
 
-    Finally, an obvious solution to many of these problems is for methods-developers and methods-users to collaborate more. 
-    The benefits to methods-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
-    A big difficulty with collaboration is that it is often asymmetrical in terms of benefit; some researchers can put a lot of time and effort into running analyses for less experienced colleagues, yet at best they come out of the experience with a middle authored paper.
-    Until these author positions are considered equal to first and last author positions, there is little incentive to collaborate extensively.
+    Finally, an obvious solution to many of these problems is for methods-developers and end-users to collaborate more. 
+    The benefits to end-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
+    A big difficulty with collaboration is that it is often asymmetrical in terms of benefit; some researchers can put a lot of time and effort into running analyses for less experienced colleagues, yet at best they gain a middle authored paper.
+    Until these author positions are considered equal to first and last author positions, there is little incentive for methods-developers to collaborate extensively in this way.
 
   \subsection{Conclusion}
-
-%I guess I’d close by saying it’s our responsbilitiy as part of the methods developing and using community to find innovative ways to tackle these challenges.
+    We are currently in an exciting period for phylogenetic comparative methods research.
+    New methods are being published with increasing regularity, and we are also beginning to question older methods and classical ways of looking at comparative questions. 
+    In addition, the field is becoming more Open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups and even continents becoming more and more common.
+    However, while embracing these changes, we also need to ensure that we do not forget that our methods have assumptions, caveats and biases like every other. 
+    These need to be highlighted so they can be accounted for in empirical analyses, and we need to be more active at providing ways of assessing these issues when publishing new methods.
+    As members of the phylogenetic comparative methods community, we have a responsibility to find innovative ways to tackle these challenges.
 
@@
     \centering
-      \includegraphics[width=12cm]{Figures/PCMCitations.pdf}
+      %\includegraphics[width=12cm]{Figures/PCMCitations.pdf}
       \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).}

### 2015-10-11 Manuscript/dark.tex
First draft completed NC

@@
 \runninghead{Limitations of PCMs}
-\keywords{PCM, assumption,\\
-\textbf{Word count}: XXX (including references, tables and figure legends)}
+\keywords{PCM, assumption, bias, caveat, phylogenetic independent contrasts, Ornstein-Uhlenbeck, trait-dependent diversification\\
+\textbf{Word count}: 3508 (including references, tables and figure legends)}
 
@@
 Unfortunately, although these limitations are often well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
-Here we explore reasons for this communication gap between those developing new methods and those using them. 
+Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
 We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
-We also highlight problems with users jumping straight into software implementations of the methods (e.g.
+We also highlight problems with users jumping straight into software implementations of methods (e.g.
 in R) that may lack documentation of biases and assumptions that are mentioned in the original papers.
@@
 
-Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method - a fact that is well established in the literature \citep[e.g.][]{freckleton2009seven,boettiger2012your,losos2011seeing,blomberg2012independent}.
+Unfortunately, PCMs also have a ``dark side''; they make various assumptions and suffer from biases in exactly the same way as any other statistical method - a fact that is well established in the literature \citep[e.g.][]{freckleton2009seven,losos2011seeing,blomberg2012independent,boettiger2012your}.
 Increasingly however, assumptions and biases are inadequately assessed in empirical studies, leading to poor model fits and misinterpreted results (see examples below).
-Additionally, little consideration is given to whether using a PCM is really appropriate for the question at hand \citep{losos2011seeing,westoby1995misinterpreting}.
+Additionally, little consideration is given to whether using a PCM is really appropriate for the question at hand \citep{westoby1995misinterpreting,losos2011seeing}.
 
@@
 \subsection{Examples of the problem}
-\subsubsection{Trait-dependent diversification} % FINISH and get GAVIN TO CHECK!!!
+Before exploring the reasons behind the communication gap, we give three brief examples of commonly used PCMs that have assumptions, biases or caveats that are often inadequately assessed in empirical studies.
+Because the aim of this paper is to provide positive ways to move forward, rather than to admonish authors for past errors, we do not cite papers we feel have fallen into these traps (especially as we are guilty of making some of the same mistakes).
 
-Analyses of trait-dependent diversification are used to detect whether particular traits promote high rates of diversification, leading to some clades becoming more diverse than others. 
-These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{}. 
-Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSEE paper \citealp{maddison2007estimating} has been cited 384 times; Google Scholar search 9th October 2015)
-However, this method has several caveats. 
-Rabosky and Goldberg BUT all mentioned in Maddison and FitzJohn papers. Not unknown but definitely misused.
-
-a strong correlation
-between character states and diversification can be
-inferred from merely a single diversification rate
-shift within a phylogenetic tree, even if the shift is
-unrelated to the modeled character
-
-This illustrates one
-form of model inadequacy: speciation rate heterogeneity
-tied to a character not under study can drive
-mistaken inference of state-dependent diversification
+\subsubsection{Phylogenetic independent contrasts}
+The phylogenetic independent contrasts method uses phylogenetic information to account for the fact that species in a comparative analysis are related to each other and thus may share similarities because they inherit them from their ancestors, not because of independent evolution \citep{felsenstein1985phylogenies,harvey1991comparative}.
+This is the most commonly used phylogenetic comparative method (\citet{felsenstein1985phylogenies} has been cited nearly 6000 times; Google Scholar search 10th October 2015), and a great deal of literature exists on the assumptions underlying the method, and ways of testing whether these assumptions are met.
+The method has three major assumptions (although there are many more than this) \citep{diaz1996testing}, (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian constant variance model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
+The third assumption is stated in \citet{felsenstein1985phylogenies} though the other two are not explicitly mentioned. 
+However each assumption is explored in many subsequent papers \citep[e.g.][]{felsenstein1988phylogenies,grafen1989phylogenetic,harvey1991comparative,garland1992procedures,purvis1995comparative,diaz1996testing,hansen1996translating,martins1997phylogenies,freckleton2000phylogenetic,garland2000using,hansen2005assessing,freckleton2006detecting,rohle2006comment}.
+There are several ways of testing these assumptions, including looking for relationships among standardised contrasts and node heights \citep{grafen1989phylogenetic,freckleton2006detecting}, absolute values of standardised contrasts and their standard deviations \citep{garland1992procedures,diaz1996testing}, and heteroscedasticity in model residuals \citep{purvis1995mammal}.
+These tests are fairly easy to implement and are included as standard model diagnostic plots in CAIC and the \texttt{caper} R package \citep{purvis1995comparative,Orme:2013aa,R-Core-Team:2014aa}.
+However, the majority of studies using phylogenetic independent contrasts do not mention testing these assumptions (\citealp{freckleton2006detecting}; although they may have tested the assumptions and not recorded this).
+In addition, because phylogenetic independent contrasts are identical to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}, these models also have the same assumptions that are equally rarely addressed.
 
 \subsubsection{Ornstein-Uhlenbeck models of trait evolution}
-Most models of trait evolution are based on the Brownian constant variance model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
-The Ornstein-Uhlenbeck (OU) model is a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{Butler:2004aa,hansen1997stabilizing}.
+Most models of trait evolution are based on the Brownian constant variance model \citep{cavalli1967,felsenstein1973maximum}.
+The Ornstein-Uhlenbeck (OU) model is a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{hansen1997stabilizing,Butler:2004aa}.
 OU models have become increasingly popular as they tend to fit the data better than Brownian motion models, and have attractive biological interpretations \citep{cooper2016}.
-For example, fit to an OU model has been seen as evidence of evolutionary constraints, stabilizing selection, niche conservatism, and selective regimes \citep{}.
-However, the OU model has several well-known caveats \citep[see][]{ho2013asymptotic,ho2014intrinsic,boettiger2012your}. 
-First, it is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, particularly for small datasets that are commonly used in these analyses (the median number of taxa used for OU studies is 58; \citealp{cooper2016}).
-Second, very small amounts of error in datasets can result in an OU model being favoured over Brownian motion simply because OU can accommodate more variance towards the tips of the phylogeny, rather than due to any interesting biological process \citep{boettiger2012your,pennell2015model}.
-Finally, the literature describing the OU model is clear that a simple explanation of clade-wide stabilizing selection is unlikely to account for data fitting an OU model \citep[e.g.][]{hansen1997stabilizing,hansen2005assessing}, but implementations of the model often state that this is the case.
+For example, fit to an OU model has been seen as evidence of evolutionary constraints, stabilising selection, niche conservatism, and selective regimes \citep{Wiens:2010aa,beaulieu2012modeling,christin2013anatomical,mahler2013exceptional}.
+However, the OU model has several well-known caveats \citep[see][]{ives2010phylogenetic,boettiger2012your,hansen2012interpreting,ho2013asymptotic,ho2014intrinsic}. 
+For example, it is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, particularly for small datasets that are commonly used in these analyses (the median number of taxa used for OU studies is 58; \citealp{cooper2016}).
+Additionally, very small amounts of error in datasets can result in an OU model being favoured over Brownian motion simply because OU can accommodate more variance towards the tips of the phylogeny, rather than due to any interesting biological process \citep{boettiger2012your,pennell2015model}.
+Finally, the literature describing the OU model is clear that a simple explanation of clade-wide stabilising selection is unlikely to account for data fitting an OU model \citep[e.g.][]{hansen1997stabilizing,hansen2005assessing}, but users of the model often state that this is the case.
 Unfortunately, these limitations are rarely taken into account in empirical studies.
 
-\subsubsection{Phylogenetically independent contrasts} % FINISH!!!
-Even the most commonly used phylogenetic comparative method, phylogentically independent contrasts has weel-known assumptions that are rarely tested.
+\subsubsection{3. Trait-dependent diversification} % Gavin - I could do with your help here, not so sure on these methods but it seemed like a timely example.
+Analyses of trait-dependent diversification are used to detect whether particular traits promote high rates of diversification, leading to some clades becoming more diverse than others \citep{nee1994reconstructed}.
+These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{}. % add citations
+Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSEE paper \citet{maddison2007estimating} has been cited 384 times; Google Scholar search 10th October 2015).
+However, \citet{rabosky2015model} recently reevaluated the method and its assumptions and showed via simulations that a strong correlation between a trait and diversification rate can be inferred from a single diversification rate shift within a tree, even if the shift is unrelated to the trait of interest.
+They suggest that many examples of trait-dependent diversification actually reflect this rate heterogeneity in trees, and thus are biologically meaningless.
+Interestingly, these caveats are mentioned in earlier papers \citep{maddison2007estimating,fitzjohn2010quantitative,FitzJohn:2012aa} but were seemingly not widely understood given the shock reaction to \citet{rabosky2015model}.
 
@@
     However, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
-    For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation models \citep{maddison2007estimating,nee1994reconstructed}.
+    For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation models \citep{nee1994reconstructed,maddison2007estimating}.
 
@@
     Additionally there is a lot of literature to wade through before getting all of the required information. 
-    For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of the most commonly used PCM, phylogenetic independent contrasts \citep{felsenstein1985phylogenies} (including how they are related to phylogenetic generalised least squares models; \citealp{blomberg2012independent,rohle2006comment}).
-    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ pages, 23 papers and a book to read.
+    For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
+    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
     This seems like an excessive amount of reading to run a simple empirical analysis.
@@
 MATLAB code; \citealp{rohlf2001comparative}), and still others provided no way of implementing their methods at all.
-    This resulted in many frustrating hours (and days and months!) trying to implement any new method you wanted to use.\\
+    This resulted in many frustrating hours (and days and months) trying to implement any new method you wanted to use.\\
 
@@
     This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
-    The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).\\
-
-    %We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology.
-It is extremely frustrating reading/reviewing papers where the authors have clearly failed to carefully read or understand the primary literature on the method they are using.
-    %Unfortunately, when authors are chastised for making errors it is often just via an assertion that they have used the method ``incorrectly'' and to ``please see the following references''.
-    %This makes it very difficult to completely understand where they went wrong.
-    %More cynically one could accuse some methods-users of trying to retrofit a question so they can use the newest methods available.
-This is all driven by our ``publish or perish'' culture, and the pressure to produce as many papers as possible out of a dataset.
-    %The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have caused additional problems (though on balance, many things have been greatly improved).
+    The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).
 
@@
     Encouraging increased efforts at reproducibility would also help, by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
-    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).\\
+    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).
 
@@
 
-%One solution is to establish a discourse page for providing R help. 
-%Currently most methods-users either email the R-sig-phylo help list or the package developer for assistance. 
-%Both approaches are suboptimal. 
-%R-sig-phylo is impossible to search effectively, leading to the same questions being dealt with repeatedly. 
-%The same is true in email exchanges between methods-users and package developers with the added issue of this being ``dark advice'' that isn't shared. 
-%A discourse page would collate all of this information and reduce pressure on package developers by allowing advanced methods-users to help novice users with simple queries. 
-%We are in the process of setting up a discourse page for this purpose (it will be called ``phyloBootstrap'' because it will provide ``support'' for PCMs).
-At Evolution 2014 many methods-developers agreed to move their R package help to this forum.
-
-  \subsubsection{2. Be more cynical about methods}
+  \subsubsection{2. Be more cynical about existing methods}
     A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
@@
 MOTMOT; \citealp{Thomas:2011aa}, and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citealp{slater2014correction}).
-    It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{losos2011seeing,westoby1995misinterpreting}.
-    Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore!
-
-%\subsubsection{3. Collaboration}
-%A obvious solution to all these problems is for methods-developers and methods-users to collaborate more. 
-%The benefits to methods-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
-%A big difficulty with collaboration is that it is often asymmetrical in terms of benefit to methods-users and methods-developers. 
-%Methods-developers and advanced methods-users can often put a lot of time and effort into running analyses for less experienced colleagues, yet at best they come out of the experience with a middle authored paper.
-%Until these author positions are considered equal to first and last author positions, there is little incentive to collaborate extensively.
+    It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
+    Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore.
 
@@
     New methods are being published with increasing regularity, and we are also beginning to question older methods and classical ways of looking at comparative questions. 
-    In addition, the field is becoming more Open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups and even continents becoming more and more common.
+    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups and even continents becoming more common.
     However, while embracing these changes, we also need to ensure that we do not forget that our methods have assumptions, caveats and biases like every other. 

### 2015-10-16 Manuscript/dark.tex
Changed title
Typo in title
Final edits and GHT corrections

@@
 % Title page information
-\title{The ``Dark Side'' of Phylogenetic Comparative Methods}
+\title{Shedding Light on the ``Dark Side'' of Phylogenetic Comparative Methods}
 \author{
@@
   $^4$ Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia. \\
-  $^*$ Corresponding author: natalie.cooper@nhm.ac.uk; Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. Fax: +353 1 677 8094; Tel: +353 1 896 5083.\\
+  $^*$ Corresponding author: natalie.cooper@nhm.ac.uk; Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. Fax: +44 1 677 8094; Tel: +44 1 896 5083.\\
 }}
@@
 
-\runninghead{Limitations of PCMs}
+\runninghead{The dark side of phylogenetic comparative methods}
 \keywords{PCM, assumption, bias, caveat, phylogenetic independent contrasts, Ornstein-Uhlenbeck, trait-dependent diversification\\
-\textbf{Word count}: 3508 (including references, tables and figure legends)}
+\textbf{Word count}: 3532 (including references, tables and figure legends)}
 
@@
 \addtolength{\parskip}{.3em}
-
 % Abstract 350 words max
@@
 However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
-Unfortunately, although these limitations are often well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
+Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
 Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
@@
 We also highlight problems with users jumping straight into software implementations of methods (e.g.
-in R) that may lack documentation of biases and assumptions that are mentioned in the original papers.
+in R) that may lack documentation on biases and assumptions that are mentioned in the original papers.
 To help solve these problems we make a number of suggestions including providing
-blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, incentives for collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
+blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
 Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
@@
 
-A long time ago in a galaxy far, far away - also known as the 1980s - phylogenetic comparative methods (PCMs) were first developed. 
-These methods use information on the evolutionary relationships of organisms to compare species \citep{harvey1991comparative}, and were initially designed to deal with the statistical non-independence of species in comparative analyses (e.g.
+Phylogenetic comparative methods (PCMs) were initially developed in the 1980s to deal with the statistical non-independence of species in comparative analyses (e.g.
 \citealp{felsenstein1985phylogenies,grafen1989phylogenetic}).
@@
 We suggest that one cause of this problem is that although researchers developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to end-users.
-Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to to assess whether that model fit is reasonable. 
+Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to assess whether that model fit is reasonable. 
 Clearly more effort is needed to bridge the widening gap between those developing methods and end-users. 
@@
 This is the most commonly used phylogenetic comparative method (\citet{felsenstein1985phylogenies} has been cited nearly 6000 times; Google Scholar search 10th October 2015), and a great deal of literature exists on the assumptions underlying the method, and ways of testing whether these assumptions are met.
-The method has three major assumptions (although there are many more than this) \citep{diaz1996testing}, (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian constant variance model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
+The method has three major assumptions (\citealp{diaz1996testing}; although there are many more assumptions than these three), (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian constant variance model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
 The third assumption is stated in \citet{felsenstein1985phylogenies} though the other two are not explicitly mentioned. 
@@
 
-\subsubsection{Ornstein-Uhlenbeck models of trait evolution}
+\subsubsection{Ornstein-Uhlenbeck (single stationary peak) models of trait evolution}
 Most models of trait evolution are based on the Brownian constant variance model \citep{cavalli1967,felsenstein1973maximum}.
-The Ornstein-Uhlenbeck (OU) model is a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{hansen1997stabilizing,Butler:2004aa}.
+The Ornstein-Uhlenbeck (OU) model can be thought of as a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{hansen1997stabilizing,Butler:2004aa}.
 OU models have become increasingly popular as they tend to fit the data better than Brownian motion models, and have attractive biological interpretations \citep{cooper2016}.
@@
 
-\subsubsection{3. Trait-dependent diversification} % Gavin - I could do with your help here, not so sure on these methods but it seemed like a timely example.
+\subsubsection{3. Trait-dependent diversification} 
 Analyses of trait-dependent diversification are used to detect whether particular traits promote high rates of diversification, leading to some clades becoming more diverse than others \citep{nee1994reconstructed}.
-These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{}. % add citations
-Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSEE paper \citet{maddison2007estimating} has been cited 384 times; Google Scholar search 10th October 2015).
+These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{Goldberg22102010,Price01052012,Givnish201455,ELE:ELE12168}.
+Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSSE paper \citet{maddison2007estimating} has been cited 384 times; Google Scholar search 10th October 2015).
 However, \citet{rabosky2015model} recently reevaluated the method and its assumptions and showed via simulations that a strong correlation between a trait and diversification rate can be inferred from a single diversification rate shift within a tree, even if the shift is unrelated to the trait of interest.
@@
     These often start out as arbitrary starting points for data exploration with no justification for their use, but over time become the way the analysis is always performed.
-    For example, % Anyone got a good example here?
 
@@
     For example, Felsenstein published the phylogenetic independent contrasts method in 1985 \citep{felsenstein1985phylogenies}, but it was not until the early 1990s that thorough critiques of the method and its assumptions began to be published \citep[e.g.][]{garland1992procedures}.
-    This time-lag is shorter with more recent methods because simulations testing the method are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions).
-    However, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
+    This time-lag is shorter with more recent methods for a number of reasons.
+    First, historically theory and software papers were generally separated, whereas currently they are combined in the same papers making it easier to run simulations. 
+    Second, simulations testing methods are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions), and finally, there are now far more people in the field and thus more papers published annually.
+    However, even with this reduced time-lag, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
     For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation models \citep{nee1994reconstructed,maddison2007estimating}.
@@
     Additionally there is a lot of literature to wade through before getting all of the required information. 
-    For example, one of us (NC) recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
+    For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
     Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
@@
     More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' http://ftp.iitm.ac.in/cran/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous.
-    The number of R packages for PCMs has increased sharply since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
+    The number of R packages for PCMs has increased markedly since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
     Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
     \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
-    However, this has instead led to an increasing number of people (including the authors) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
-    This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
+    However, this has instead led to an increasing number of people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
+    This is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
     The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).
@@
 \subsection{How can we solve some of these problems?}
-  \subsubsection{1. Simplify, summarize and share}
+  \subsubsection{1. Simplify, summarise and share}
     Many of the problems above are caused by methods papers being difficult to read. 
@@
     A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
+    As the Manic Street Preachers put it, ``Cynicism is the only thing that keeps me sane''.
     At a minimum, it is worth reading the original papers describing a method, plus any recent updates, and looking carefully for assumptions and caveats that may affect the analyses at hand.
@@
     It is also important to consider whether the method will work on a particular dataset.
-    For example it is worth considering how many species are required for reasonable power (often methods require more species than are usually available, e.g.
-the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily be applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most empirical analyses have \textless 1000 tips), whether the method is influenced by polytomies (many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution \citealp{cooper2010body}), and if the method can use non-ultrametric trees (some current implementations of the OU model should not be used with non-ultrametric trees because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix, e.g.
-MOTMOT; \citealp{Thomas:2011aa}, and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citealp{slater2014correction}).
+    For example it is worth considering how many species are required for reasonable power. 
+    Often methods require more species than are usually  available.
+For example, the new trait-dependent diversification method of \citet{rabosky2015robust} is ``primarily applicable to phylogenies that include at least several thousand tips'' although the authors suggest that most empirical analyses have \textless 1000 tips.
+    Other considerations include whether the method is influenced by polytomies and whether the method is applicable to both ultrametric and non-ultrametric trees. 
+    Indeed, many methods arbitrarily resolve polytomies using zero length branches, thus polytomies can inflate rates of evolution, and bias models of evolution  \citep{cooper2010body}. 
+    Some current implementations of the OU model should not be used with non-ultrametric trees (e.g.
+MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citep{slater2014correction}.
+
     It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
@@
     Second we need to fund pure methods development, including incremental methods. 
-    Currently it is difficult to get funding for purely methods driven research; an empirical component is always needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
+    Currently it is difficult to get funding for purely methods driven research; an empirical component is generally needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
     This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.
@@
     Finally, an obvious solution to many of these problems is for methods-developers and end-users to collaborate more. 
-    The benefits to end-users are obvious, but methods-developers can also benefit by gaining extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released.
-    A big difficulty with collaboration is that it is often asymmetrical in terms of benefit; some researchers can put a lot of time and effort into running analyses for less experienced colleagues, yet at best they gain a middle authored paper.
-    Until these author positions are considered equal to first and last author positions, there is little incentive for methods-developers to collaborate extensively in this way.
+    Both parties can benefit greatly by collaboration. 
+    Some benefits are obvious, e.g.
+methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released, whereas end-users can work with the most cutting-edge methods and software.
+    Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.
+    There is a huge benefit in sharing ideas as well as products (data and code) and time, as long as the benefits to both parties are not heavily asymmetric.
 
@@
     New methods are being published with increasing regularity, and we are also beginning to question older methods and classical ways of looking at comparative questions. 
-    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups and even continents becoming more common.
-    However, while embracing these changes, we also need to ensure that we do not forget that our methods have assumptions, caveats and biases like every other. 
+    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups, and even continents, becoming more common.
+    However, while embracing these changes, we also need to ensure that we do not forget that PCMs have assumptions, caveats and biases like every other method. 
     These need to be highlighted so they can be accounted for in empirical analyses, and we need to be more active at providing ways of assessing these issues when publishing new methods.
@@
   \subsection{Acknowledgments}
-    This paper is based on a talk given by NC at the Methods in Ecology and Evolution 5th Anniversary Symposium.
+    This paper is based on a talk given at the Methods in Ecology and Evolution 5th Anniversary Symposium (NC).
     Thanks to the Society of Systematic Biologists for funding the symposium at Evolution 2014 where the ideas in this paper were first consolidated. 
@@
     \centering
-      %\includegraphics[width=12cm]{Figures/PCMCitations.pdf}
+      \includegraphics[width=12cm]{Figures/PCMCitations.pdf}
       \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).}
@@
     \centering
-      %\includegraphics[width=12cm]{Figures/PCMRpackages.pdf}
+      \includegraphics[width=12cm]{Figures/PCMRpackages.pdf}
       \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time from 1980-2014.

### 2015-12-02 Manuscript/response.tex
First draft of response

\documentclass[12pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[osf]{mathpazo}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\signature{Natalie Cooper \\ Gavin Thomas \\ Chris Venditti \\ Andrew Meade \\ Rob Freckleton}
\address{Department of Life Sciences \\ Natural History Museum \\ Cromwell Road \\ London, SW7 5BD \\ nhcooper12@gmail.com}
\longindentation=0pt
\begin{document}

\begin{letter}{}
\opening{Dear Dr O'Hara}

Please find enclosed our resubmission of our manuscript ``Shedding Light on the ``Dark Side'' of Phylogenetic Comparative Methods" (MEE-15-10-697). We include below details of how we have dealt with the editor's and reviewers' comments. The comments are in blue and our response is in black.

Thanks to Bob, Amy and Simon for all your helpful comments. This was a tricky paper to write, and we debated heavily about whether to write it at all. Finding the right tone was indeed very difficult, and we clearly haven't completely succeeded there. Also while we had started to "Do Something" about some of this around the time I first presented this, in the interim both Rich and I have moved countries and jobs (Rich out of academia) delaying some projects and derailing others completely!

\textcolor{blue}{\textbf{Bob O'Hara's comments}}\\

\textcolor{blue}{The reviewers and Associate Editor are positive about your work, however there are a few issues that need attention. Therefore, I invite you to respond to the reviewer(s)' comments and revise your manuscript. This might be closer to a Major Revisions in some ways: there are several issues to be addressed. At times the manuscript comes across as a bit whiny ("Why don't you lot do something?"). Tone can be difficult to get right, and there are some good suggestions (e.g. use Markdown to show your code), but after reading this I think I might avoid PCMs, as it seems there are a lot of pitfalls that I won't know about unless I work with someone who knows about them. Evidently it is not my destiny to come over to the Dark Side.}

We have tried to correct the tone where we can.

\textcolor{blue}{I think the authors' suggestions that Someone Needs to Do Something would be stronger if they Started to Do Something. For example, they mention the BES QSIG's FGE pages, so why not start up a page on PCMs? An intro to running PCMs would be a necessary start, plus perhaps a Checklist of Things to be Checked (And How to Check Them). If the authors could find a day or two to set these up, then I think this would give the manuscript a more positive tone by announcing it, and might (if you are really really lucky) encourage others to add their knowledge.}

I'm involved with the FGE pages, and was planning a few PCM posts

\textcolor{blue}{Minor Comments
Abstract: This should be in the proper BES-approved format (i.e. as a bulleted list).}

Fixed

\textcolor{blue}{l315-6: A proper citation is needed, I think. Because who wouldn't want to see this cited formally?}

\textcolor{blue}{\textbf{Referee 1: Amy Zanne}}\\

\textcolor{blue}{As an audience member at this symposium and PCM end user, I was made well aware of the nature of use and abuse of PCMs. I walked away from this symposium oddly hopeful though that through collaborations, discussions and a number of items detailed in this review, solutions to intractable problems and public awareness of misuse can be addressed. I enjoyed this review as it provided concrete identification of the problems facing us without pointing fingers at anyone in particular. It also provided tractable solutions for how we can move forward. I found it a nice way to pause and take stock of where we are and how we can better go in the future.}

\textcolor{blue}{My biggest suggestion is that there can be some tightening of the text in a few places and a better mapping in others to the headings. I’ll detail this specifically below.}

Thanks for the helpful and encouraging comments. :)

\textcolor{blue}{L179 I’d suggest that some of this folklore also comes from within the labs developing PCM. The reason Sam Price and I proposed the R hackathon at NESCent was because each of us had knowledge about PCM issues and solutions that were lab based from the groups where we came from who were developing methods. Some of this folklore is based on good practices, others not but the point is it is being passed down outside of the public literature.}

\textcolor{blue}{L224-225 I think you can just say long and tedious or something along those lines. Why is “>20 pages” too long and what gives me a headache may be different to what gives you a “headache”.}

\textcolor{blue}{L231 “This” is somewhat unclear. I’d make it clear that this review led to you needing to read more than 300 manuscript pages and a book to fully understand the issue.}

\textcolor{blue}{L234 “required” for what? I’m assuming to correctly run the analysis or something along those lines?}

\textcolor{blue}{L257 drop a “to”}

\textcolor{blue}{L279 Again “incomprehensible” seems to depend on who is reading it. I’d argue that most methods papers are written for other methods developers as those are the audiences that will be reviewing the paper. They may or may not find the papers incomprehensible. The papers are likely to be less accessible but important for end users.}

\textcolor{blue}{L284 I’d go further than saying “there’s no harm in doing”. In fact there’s a lot of benefit (carrots). Although as you later note, we currently have no sticks making us do this. This seems like the best way to get the newly developed method into the hands of end users, make it accessible, but I agree that it is more work.}

\textcolor{blue}{L287 “help” with what? I fully agree that reproducibility is important in science but I’m unclear how it helps with this particular issue of “simplify, summarise and share”. I can imagine some avenues but I think it would be helpful to spell them out for the reader as otherwise it seems off topic.}

\textcolor{blue}{L297 “bigger problem” for what? I’m guessing to come up to speed with understanding how the method works.}

\textcolor{blue}{L305 I’ve appreciated the Prometheus wiki ($http://prometheuswiki.publish.csiro.au/tiki-custom_home.php$) as a place for methods to be posted and reviewed, as well as providing an avenue for discussion of those methods. You have a number listed here so don’t need another per se but I think this is a nice and well-used resource that would serve as a nice model for PCMs.}

\textcolor{blue}{L311 I think there’s plenty of incentive but it’s the carrot kind (it gets the PCM in an accessible form into the hands of an end user, meaning it’s more likely to be used/reused) rather than the stick. Perhaps we do better with sticks than carrots, which is why Genbank works and more people are making public their data as journals require it?}

\textcolor{blue}{L313 To me this section reads as a series of suggestions rather than that we should be more cynical. I might broaden this subject heading. For instance, beginning with “It” on L323 this seems to be about best practices and things to keep in mind rather than being cynical. Basically you are outlining a nice checklist that we should keep in mind for most analyses (i.e., what’s the sample size needed for enough power, do we need a complete and fully bifurcated phylogeny, etc.). Such a list would make a nice table… There’s also of course assumptions/issues to consider that are very tied to specific PCMs. So, my take home is there are some things to keep in mind generally across most/all analyses and there are things to keep in mind specific to this analysis you are considering (the latter is harder to come by). There are a lot of “consider” and “for example” in this section.}

\textcolor{blue}{L341 It might also be useful to mention that reviewers and editors sometimes require authors to run PCMs even when these were not part of the original plan for the project. Sometimes these are useful suggestions that help to broaden the question, other times they aren’t appropriate. It would be good to empower authors to say no when it’s the latter.}

\textcolor{blue}{L346 Again this section is more than just “Incentives” it feels perhaps like “Solutions” or something like that.}

\textcolor{blue}{L359 I agree that assumptions are not always given or are hard to find. Perhaps add “lists and” following “include”}

\textcolor{blue}{L364 I fully agree that it’s hard to find funding for method development but it’s also hard to find funding for simple data collection. I think what drives finding are theoretical/hypothesis driven science questions.}

\textcolor{blue}{L366 I don’t understand what this sentence means as written “This unfairly prejudices funding bodies (and hiring committees)”. Define “this” perhaps. I’d take it that funding bodies and hiring committees are creating these “biases” and driving them forward. We certainly need to change how the field appreciates methods developers but it comes at all of these levels.}

\textcolor{blue}{L372 “get” should be “gets”}

\textcolor{blue}{L386 Insert “is” before “becoming”}

\textcolor{blue}{Reviewer 2: Simon Blomberg}

\textcolor{blue}{After reading the abstract of this paper, I was sceptical. However, upon reading it I found that it makes some reasonable points. First I will discuss the negative aspects.}

\textcolor{blue}{My main issue is with sections 2 (beginning line 218) and 3 (beginning line 238). The fact that the literature is too difficult and opaque for some readers is not the fault of the literature. These statistical methods are NOT easy, so researchers will need to do due diligence and take the time to learn this stuff properly (including caveats and assumptions, when they are known and described in the literature). In any statistical analysis, the buck stops with the authors. It is your responsibility to get it right. You can't complain that you are using methods that you don't understand. If that is the case you should go back and learn more about the methods. Complaining that there is too much maths and it gives you a headache (lines 224-225) is not a reason to castigate the authors of methods. Mathematics is the language of modern evolutionary biology and the sooner people get used to that, the better. Don't like maths? Don't do quantitative science. Complaining that the literature is too big and difficult and going ahead with an analysis anyway is just academic laziness. Don't want to get to grips with the literature? Collaborate with someone who knows it, like a statistician. The authors appear to be promoting a new section for such papers, as assumptions and caveats can be in the Intro, Methods, Results and/or Discussion (lines 235-237). Well how about this: read the papers properly! Again, laziness here.}

\textcolor{blue}{The authors complain that researchers often jump straight into using a method (line 259-263) without knowing its limitations. Well this is just a problem with due diligence again. It's not the method developer's fault. The authors complain that this seems to be particularly a problem with methods implemented in R. Well, R is free software. You have the source code. If you really want to know how a method works, study the source code! You can't do that for other, proprietary software. You have to trust the company/lab that produces the software that it does what is claimed on the box. Using free software like R is thus more scientific than using proprietary software. It \_allows\_ researchers to do due diligence.}

\textcolor{blue}{The authors claim that novice users may fail to understand a method because the accompanying paper is incomprehensible (lines 277-279). Well, that novice needs to do more work. Usually that involves reading around the topic, reading the source code, and maybe running some simulations. Students should be doing this all the time: it's called \_learning\_. It's their job. In fact it is every researcher's job.}

\textcolor{blue}{And now for the positive points.}

\textcolor{blue}{The authors ask for more and better documentation, including YouTube tutorials, blog posts, etc. This is to be commended. We always need more, better documentation. Like scientists who always need more, better data. Unfortunately, most phylogenetic comparative methods are developed by academic researchers who may not have the resources (time or money) to do all this. That they offer their work for free (in R) is a major gift to the scientific community. I have a suggestion: Don't like the (lack of) documentation? Write it yourself! You will learn way more than just blindly going ahead and using a method uncritically. If it's an R package, contact the author and say you would like to contribute to the documentation. I suspect that most R developers would jump at the chance to have someone provide more/better documentation. This could be directly editing the package help files and/or providing vignettes. You  might even be included as an author of the package. If you really have a lot to say about methods, write a book! There are plenty of books on R, including a couple that describe PCMs. Again, more/better books are welcome. Another solution would be to start giving workshops and seminars on a fee for service basis. The fees could be ploughed back into improving the software and documentation. Workshops on statistical analysis (of any kind) are very popular and you do not need to be a package's developer to deliver these. See  http://www.highstat.com/ for an example of how to make a business out of it. I'm not sure that the evolutionary biology community is big enough to support this, but it may be.}

\textcolor{blue}{The authors encourage more reproducible research and open source science (lines 287-290). This is a growing movement and will no doubt help raise the quality of research.}

\textcolor{blue}{The authors discuss editorial policies allowing incremental improvements and diagnostics for PCMs. This would also be of value. I also applaud the incentives suggested in lines 346-378. They would be most welcome.}

\textcolor{blue}{Minor criticisms:}

\textcolor{blue}{line 124-125. "Brownian constant variance model" is a complete contradiction in terms. Brownian motion has infinite total variation and quadratic variation increasing with time. If a process has constant variance, it's not Brownian motion!}

\textcolor{blue}{line 145-150. It may be that a particular data set cannot support more than one tree-wise set of parameters. But the authors are right. In general there is no reason to expect that a whole tree would be governed by just one process. Of course, you can fit more complicated models in different parts of the tree, but again the quality of the data may not make this feasible.}

\textcolor{blue}{line 339-340: It's just not true that, "non-ultrametric trees cause variance-covariance matrices to become non-tree like". You can jump back and forth between a tree representation and a matrix representation, and it is just not affected by ultrametricity. In the ape package, there is vcv.phylo to convert a tree to a matrix, and vcv2phylo to convert a matrix back to a tree. (I wrote the latter.)}

Let us know if you require any further information,

\closing{Yours sincerely,}

\end{letter}
\end{document}

### 2015-12-07 Manuscript/response.tex
Manics
Manics reference, remembering how to use git.
Ulrametric
Clarified point about non-ultrametric trees and OU transforms.

@@
 \textcolor{blue}{l315-6: A proper citation is needed, I think. Because who wouldn't want to see this cited formally?}
+
+Citation is something like: Manic Street Preachers. 1996. Mr Carbohydrate. A Design for Life [CD 1], track 2. Columbia Records.
 
@@
 
+We believe that this is a misunderstanding of our point and reflects a lack of clarity in our writing. We now write:
+
+"Some current implementations of the OU model should not be used with non-ultrametric trees (e.g.
+MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix.
+The problem is that where at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical.
+Worked examples and explanations are provided in \citep{slater2014correction}.
+Although this is not a problem with applying the OU model to non-ultrametric trees per se, it is an example of different implementations of a common model that some users may not be aware of.
+"
+
 Let us know if you require any further information,

### 2015-12-07 Manuscript/dark.tex
Ulrametric
Clarified point about non-ultrametric trees and OU transforms.

@@
     Some current implementations of the OU model should not be used with non-ultrametric trees (e.g.
-MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citep{slater2014correction}.
+MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix.
+The problem is that where at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical.
+Worked examples and explanations are provided in \citep{slater2014correction}.
+Although this is not a problem with applying the OU model to non-ultrametric trees per se, it is an example of different implementations of a common model that some users may not be aware of.
 

### 2015-12-07 Manuscript/response.tex
Addressing due diligence comments

@@
 
+We have removed reference to
+
 \textcolor{blue}{The authors complain that researchers often jump straight into using a method (line 259-263) without knowing its limitations.
@@
 
+We agree.
+However, we also argue that doing due diligence is a responsibility of both the users (in reading widely and deeply) and the developers (in writing clearly and stating assumptions, limitations etc).
+
 \textcolor{blue}{The authors claim that novice users may fail to understand a method because the accompanying paper is incomprehensible (lines 277-279).
@@
 In fact it is every researcher's job.}
+
+We absolutely agree that readers must do the work to understand methods but reiterate the point above - due diligence is a responsibility for developers as well as users.
+We have revised lines 277-279 accordingly.
 
@@
 
+This was an unfortunate error on our part. Of course the variance is not constant in the Brownian motion model. We have replaced the two instances of "constant variance" with "motion".
+
 \textcolor{blue}{line 145-150.

### 2015-12-07 Manuscript/dark.tex
Addressing due diligence comments

@@
 This is the most commonly used phylogenetic comparative method (\citet{felsenstein1985phylogenies} has been cited nearly 6000 times; Google Scholar search 10th October 2015), and a great deal of literature exists on the assumptions underlying the method, and ways of testing whether these assumptions are met.
-The method has three major assumptions (\citealp{diaz1996testing}; although there are many more assumptions than these three), (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian constant variance model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
+The method has three major assumptions (\citealp{diaz1996testing}; although there are many more assumptions than these three), (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian motion model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
 The third assumption is stated in \citet{felsenstein1985phylogenies} though the other two are not explicitly mentioned. 
@@
 \subsubsection{Ornstein-Uhlenbeck (single stationary peak) models of trait evolution}
-Most models of trait evolution are based on the Brownian constant variance model \citep{cavalli1967,felsenstein1973maximum}.
+Most models of trait evolution are based on the Brownian motion model \citep{cavalli1967,felsenstein1973maximum}.
 The Ornstein-Uhlenbeck (OU) model can be thought of as a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{hansen1997stabilizing,Butler:2004aa}.
@@
     A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
-    It is not unusual for papers to be full of equations, \textgreater 20 pages long, and written in a way that gives one a headache. 
-    Additionally there is a lot of literature to wade through before getting all of the required information. 
-    For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
+    It is not unusual for papers to be full of equations, \textgreater 20 pages long. 
+    Such detail and mathematical literacy is critical in describing methods and facilitates testing/assessment, implementation and future developments of the method.
+    In most cases equations could not simply be replaced with text.
+   However, from the end user perspective it may more helpful to understand the assumptions and caveats.
+   Within many comparative methods assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
+    They are rarely neatly corralled in one place.
+    When combined with the vast literature on comparative methods it becomes easy to miss pertinent details. 
+     We recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
     Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
-    This seems like an excessive amount of reading to run a simple empirical analysis.
-    Another issue is that it can sometimes be hard to find the information required. 
-    Assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
-    They are rarely neatly corralled in one place, making it easy to miss pertinent details.
+    Methods developers are not responsible for users conducting due diligence or reading the literature. 
+    The onus is, quite rightly in our opinion, on the reader/user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
+    However, simple steps, such as subheadings that point to caveats and assumptions could add clarity and limit method misuse.
 
@@
 MATLAB code; \citealp{rohlf2001comparative}), and still others provided no way of implementing their methods at all.
-    This resulted in many frustrating hours (and days and months) trying to implement any new method you wanted to use.\\
+    This resulted in many frustrating hours (and days and months) trying to implement any new method you wanted to use.
+    Writing your own implementation may be the best way to learn the intricacies of a method but is a major hurdle and would necessarily dissuade many potential users.  \\
 
@@
     Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
-    \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
+    The benefits of R are clear. 
+    It is available to all, has a wide and engaged support community, and perhaps most importantly the source code for new methods is accessible.
+    Users can, if they choose to, fully explore any new method by examining the source code and running their own simulations. 
+    Indeed, \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
     However, this has instead led to an increasing number of people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
-    This is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
+    This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
+    Manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
     The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).
@@
     This is not entirely the fault of those writing these papers; most journals have strict word limits so it is often easier to use a one line equation rather than a paragraph of description. 
-    However, it is hard to complain about novice users failing to understand a method when the paper is incomprehensible.
+    As we note above, the onus is on the reader to read extensively and to do due diligence.
+    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target readers, especially in describing assumptions and limitations.
     One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 

### 2015-12-08 Manuscript/response.tex
GHTs edits

@@
 
-Citation is something like: Manic Street Preachers. 1996. Mr Carbohydrate. A Design for Life [CD 1], track 2. Columbia Records.
+Exactly how a song should be cited is not entirely clear but we have added a the following citation: Manic Street Preachers. 1996. Mr Carbohydrate. A Design for Life [CD 1], track 2. Columbia Records.
 
@@
 
+We agree and make exactly these points on line XX: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles."
+
 \textcolor{blue}{L224-225 I think you can just say long and tedious or something along those lines.
@@
 
+We have removed reference to headaches and also added the point that often equations are necessary.
+
 \textcolor{blue}{L231 “This” is somewhat unclear. I’d make it clear that this review led to you needing to read more than 300 manuscript pages and a book to fully understand the issue.}
+
+We have expended on this to make the point that the difficulty relates to the problem of how much reading is necessary to have sufficient understanding.
 
@@
 
-We have removed reference to
+We have removed reference to headaches and revised these sections substantially.
+We agree that due diligence is key.
+However our point is that due diligence is not just the responsibility of readers/users but also of developers.
+It is not simply laziness on the readers side that leads to misunderstanding.
+Often papers, especially those with software, are presented as if the method is faultless.
+This is a symptom of how papers are published but can very easily mislead novice users.
+Clarity in writing is the responsibility of the write, thorough reading is the responsibility of the reader.
 
@@
 We agree.
-However, we also argue that doing due diligence is a responsibility of both the users (in reading widely and deeply) and the developers (in writing clearly and stating assumptions, limitations etc).
+However, as we note above we also argue that doing due diligence is a responsibility of both the users (in reading widely and deeply) and the developers (in writing clearly and stating assumptions, limitations etc).
 

### 2015-12-08 Manuscript/dark.tex
GHTs edits

@@
     They are rarely neatly corralled in one place.
-    When combined with the vast literature on comparative methods it becomes easy to miss pertinent details. 
-     We recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
+     The sheer volume of literature can become a barrier to understanding, even, and perhaps especially, for the best known methods.
+     For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
     Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
-    Methods developers are not responsible for users conducting due diligence or reading the literature. 
+    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may be easy for experienced users but not clear for new users.
+    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details.
+    Many of these problems relate to due diligence for both reader and developer. 
+    Methods developers are not responsible for making sure that all users read the literature. 
     The onus is, quite rightly in our opinion, on the reader/user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
-    However, simple steps, such as subheadings that point to caveats and assumptions could add clarity and limit method misuse.
+    However, simple steps, such as subheadings that point to caveats and assumptions, could add clarity and limit method misuse.
 
@@
   \subsubsection{1. Simplify, summarise and share}
-    Many of the problems above are caused by methods papers being difficult to read. 
-    This is not entirely the fault of those writing these papers; most journals have strict word limits so it is often easier to use a one line equation rather than a paragraph of description. 
+    Many of the problems above are the result of methods that are difficult to communicate in purely descriptive terms.
+    This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text.
     As we note above, the onus is on the reader to read extensively and to do due diligence.
@@
     Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there is no harm in doing this and hosting it on a personal website.
+    Some package developers provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. 
     The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
@@
     At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).
+    An important point here is that we do not see these possible solutions as entirely the responsibility of the method developer or package writer. 
+    The comparative methods community can help too and indeed often does via platforms such as r-sig-phylo.
+    However, the ability to find helpful replies to lists depends on the subject headings and details in the text.
+    A longer lasting solution would be to make published methods papers, and especially those that have a software focus, more readily updatable. 
+    It should not be necessary to publish a new paper for updates of method.
+    One example that gets around this problem is PLoS Currents where published papers can be updated relatively quickly with the full history of updated articles versioned.
 

### 2015-12-08 Manuscript/response.tex
Dealt with BOH and AZ's comments

@@
 \usepackage{amsmath,amssymb}
-\signature{Natalie Cooper \\ Gavin Thomas \\ Chris Venditti \\ Andrew Meade \\ Rob Freckleton}
-\address{Department of Life Sciences \\ Natural History Museum \\ Cromwell Road \\ London, SW7 5BD \\ nhcooper12@gmail.com}
+\signature{Natalie Cooper \\ Gavin Thomas \\ Richard FitzJohn}
+\address{Department of Life Sciences \\ Natural History Museum \\ Cromwell Road \\ London, SW7 5BD \\ natalie.cooper@nhm.ac.uk}
 \longindentation=0pt
@@
 
-We have tried to correct the tone where we can.
+We have tried to correct the tone where we can. It's very tricky!
 
@@
 
-I'm involved with the FGE pages, and was planning a few PCM posts
+I'm involved with the FGE pages, and was planning to add a few PCM pages to this.
+As mentioned above, we had planned to "Do Something" and got a little way towards planning this with other members of the community, but personal constraints have prevented us from getting any further.
+We do hope to get something set up in the future, so have added a comment to this effect.
+We completely agree that this would be stronger with that kind of support set up already.
 
@@
 
-Fixed
+Fixed.
 
@@
 
-\textcolor{blue}{My biggest suggestion is that there can be some tightening of the text in a few places and a better mapping in others to the headings. I’ll detail this specifically below.}
+\textcolor{blue}{My biggest suggestion is that there can be some tightening of the text in a few places and a better mapping in others to the headings. I'll detail this specifically below.}
 
-Thanks for the helpful and encouraging comments. :)
+Thanks for the helpful and encouraging comments.
 
-\textcolor{blue}{L179 I’d suggest that some of this folklore also comes from within the labs developing PCM.
+\textcolor{blue}{L179 I'd suggest that some of this folklore also comes from within the labs developing PCM.
 The reason Sam Price and I proposed the R hackathon at NESCent was because each of us had knowledge about PCM issues and solutions that were lab based from the groups where we came from who were developing methods.
@@
 
-We agree and make exactly these points on line XX: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles."
+***We agree and make exactly these points on line XX: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles." However to make this clearer we have specified that it "tends to be passed down from PIs to graduate students, and within labs developing these methods".
+We also have been more explicit that sometimes it is based on "best practice".
 
@@
 
-\textcolor{blue}{L231 “This” is somewhat unclear. I’d make it clear that this review led to you needing to read more than 300 manuscript pages and a book to fully understand the issue.}
+\textcolor{blue}{L231 “This” is somewhat unclear. I'd make it clear that this review led to you needing to read more than 300 manuscript pages and a book to fully understand the issue.}
 
-We have expended on this to make the point that the difficulty relates to the problem of how much reading is necessary to have sufficient understanding.
+We have expanded on this to make the point that the difficulty relates to the problem of how much reading is necessary to have sufficient understanding.
 
-\textcolor{blue}{L234 “required” for what? I’m assuming to correctly run the analysis or something along those lines?}
+\textcolor{blue}{L234 “required” for what? I'm assuming to correctly run the analysis or something along those lines?}
+
+This section has been modified in response to Simon's comments so hopefully it is now clearer that we mean to understand the method.
 
@@
 
+Fixed.
+
 \textcolor{blue}{L279 Again “incomprehensible” seems to depend on who is reading it.
-I’d argue that most methods papers are written for other methods developers as those are the audiences that will be reviewing the paper.
+I'd argue that most methods papers are written for other methods developers as those are the audiences that will be reviewing the paper.
 They may or may not find the papers incomprehensible.
@@
 
-\textcolor{blue}{L284 I’d go further than saying “there’s no harm in doing”.
-In fact there’s a lot of benefit (carrots).
+****This is a very good point
+
+\textcolor{blue}{L284 I'd go further than saying “there's no harm in doing”.
+In fact there's a lot of benefit (carrots).
 Although as you later note, we currently have no sticks making us do this.
@@
 
+Agreed, we have modified this line to say "there are great benefits to doing this"
+
 \textcolor{blue}{L287 “help” with what?
-I fully agree that reproducibility is important in science but I’m unclear how it helps with this particular issue of “simplify, summarise and share”.
+I fully agree that reproducibility is important in science but I'm unclear how it helps with this particular issue of “simplify, summarise and share”.
 I can imagine some avenues but I think it would be helpful to spell them out for the reader as otherwise it seems off topic.}
 
-\textcolor{blue}{L297 “bigger problem” for what? I’m guessing to come up to speed with understanding how the method works.}
+We have modified this to "Encouraging increased efforts at reproducibility would also help make methods more accessible to new users, by providing fully worked examples that can act as a starting point for their analyses.".
+Our point is that if someone uses a method and makes their analysis reproducible by sharing their code etc, new users can use this as a starting point for their own analyses.
 
-\textcolor{blue}{L305 I’ve appreciated the Prometheus wiki ($http://prometheuswiki.publish.csiro.au/tiki-custom_home.php$) as a place for methods to be posted and reviewed, as well as providing an avenue for discussion of those methods.
-You have a number listed here so don’t need another per se but I think this is a nice and well-used resource that would serve as a nice model for PCMs.}
+\textcolor{blue}{L297 “bigger problem” for what? I'm guessing to come up to speed with understanding how the method works.}
 
-\textcolor{blue}{L311 I think there’s plenty of incentive but it’s the carrot kind (it gets the PCM in an accessible form into the hands of an end user, meaning it’s more likely to be used/reused) rather than the stick.
+Sorry we weren't clear here.
+The paragraph above is dealing with recommendations for new papers, but there are loads of already published papers that don't have accompanying blog posts or code etc., that also need to be summarised somehow.
+We have changed this sentence to "Summarising the glut of literature that already exists for established methods is a more difficult problem to solve."
+
+\textcolor{blue}{L305 I've appreciated the Prometheus wiki ($http://prometheuswiki.publish.csiro.au/tiki-custom_home.php$) as a place for methods to be posted and reviewed, as well as providing an avenue for discussion of those methods.
+You have a number listed here so don't need another per se but I think this is a nice and well-used resource that would serve as a nice model for PCMs.}
+
+It's always good to include more examples so we have added the Prometheus wiki to our list. We weren't aware of this beforehand so thanks for pointing it out.
+
+\textcolor{blue}{L311 I think there's plenty of incentive but it's the carrot kind (it gets the PCM in an accessible form into the hands of an end user, meaning it's more likely to be used/reused) rather than the stick.
 Perhaps we do better with sticks than carrots, which is why Genbank works and more people are making public their data as journals require it?}
+
+We have removed this line, but agree that a GenBank like system would be great for data sharing in ecology and evolution.
 
@@
 For instance, beginning with “It” on L323 this seems to be about best practices and things to keep in mind rather than being cynical.
-Basically you are outlining a nice checklist that we should keep in mind for most analyses (i.e., what’s the sample size needed for enough power, do we need a complete and fully bifurcated phylogeny, etc.).
-Such a list would make a nice table… There’s also of course assumptions/issues to consider that are very tied to specific PCMs.
+Basically you are outlining a nice checklist that we should keep in mind for most analyses (i.e., what's the sample size needed for enough power, do we need a complete and fully bifurcated phylogeny, etc.).
+Such a list would make a nice table… There's also of course assumptions/issues to consider that are very tied to specific PCMs.
 So, my take home is there are some things to keep in mind generally across most/all analyses and there are things to keep in mind specific to this analysis you are considering (the latter is harder to come by).
@@
 
+We have changed the subheading to "Factors to consider before using a method" to more accurately reflect its content.
+We have also rephrased the section to reduce the number of considers and for examples.
+
 \textcolor{blue}{L341 It might also be useful to mention that reviewers and editors sometimes require authors to run PCMs even when these were not part of the original plan for the project.
-Sometimes these are useful suggestions that help to broaden the question, other times they aren’t appropriate.
-It would be good to empower authors to say no when it’s the latter.}
+Sometimes these are useful suggestions that help to broaden the question, other times they aren't appropriate.
+It would be good to empower authors to say no when it's the latter.}
+
+Agreed.
+This can be a big problem so we have may this point more explicitly: "In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions."
 
@@
 
+We have changed this to "Solutions and incentives" to better reflect the content.
+
 \textcolor{blue}{L359 I agree that assumptions are not always given or are hard to find. Perhaps add “lists and” following “include”}
 
-\textcolor{blue}{L364 I fully agree that it’s hard to find funding for method development but it’s also hard to find funding for simple data collection.
-I think what drives finding are theoretical/hypothesis driven science questions.}
+Added.
 
-\textcolor{blue}{L366 I don’t understand what this sentence means as written “This unfairly prejudices funding bodies (and hiring committees)”.
+\textcolor{blue}{L364 I fully agree that it's hard to find funding for method development but it's also hard to find funding for simple data collection.
+I think what drives funding are theoretical/hypothesis driven science questions.}
+
+Agreed.
+It's pretty hard these days to get funding for *any* kind of research!
+We have added a line "(although it is also hard to get funding for empirical projects)" to clarify that we're aware the situation is bad for everyone at present.
+
+\textcolor{blue}{L366 I don't understand what this sentence means as written “This unfairly prejudices funding bodies (and hiring committees)”.
 Define “this” perhaps.
-I’d take it that funding bodies and hiring committees are creating these “biases” and driving them forward.
+I'd take it that funding bodies and hiring committees are creating these “biases” and driving them forward.
 We certainly need to change how the field appreciates methods developers but it comes at all of these levels.}
+
+The point we were trying to make is that methods development is hard to get funded, thus your CV lacks big grants making it harder to secure faculty positions/tenure.
+however, we recognise that is was ambiguous, and also needs action at all levels.
+We have therefore deleted this line.
 
@@
 
+Fixed.
+
 \textcolor{blue}{L386 Insert “is” before “becoming”}
+
+Fixed.
 

### 2015-12-08 Manuscript/dark.tex
Dealt with BOH and AZ's comments

@@
 
-Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
+\begin{enumerate}
+  \item Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
   However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
-Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
+  \item Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
   Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
-We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
+  \item We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
   We also highlight problems with users jumping straight into software implementations of methods (e.g.
 in R) that may lack documentation on biases and assumptions that are mentioned in the original papers.
-To help solve these problems we make a number of suggestions including providing
+  \item To help solve these problems we make a number of suggestions including providing
   blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
   Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
+\end{enumerate}
 
@@
     Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
-    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
-    Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules. 
+    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing these methods (and it is occasionally closer to fiction than fact!).
+    Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules.
     Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
@@
     A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
-    It is not unusual for papers to be full of equations, \textgreater 20 pages long. 
-    Such detail and mathematical literacy is critical in describing methods and facilitates testing/assessment, implementation and future developments of the method.
+    It is not unusual for papers to be long, tedious and full of equations. 
+    Of course, such detail and mathematical literacy is critical in describing methods and facilitates testing/assessment, implementation and future developments of the method.
     In most cases equations could not simply be replaced with text.
    However, from the end user perspective it may more helpful to understand the assumptions and caveats.
-   Within many comparative methods assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
+   Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
     They are rarely neatly corralled in one place.
@@
     Users can, if they choose to, fully explore any new method by examining the source code and running their own simulations. 
-    Indeed, \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
+    Indeed, \citet{freckleton2009seven} suggested that the ability to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
     However, this has instead led to an increasing number of people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
@@
     One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 
-    Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there is no harm in doing this and hosting it on a personal website.
+    Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there are great benefits to doing this anyway and hosting it on a personal website.
     Some package developers provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. 
     The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
-    Encouraging increased efforts at reproducibility would also help, by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
+    Encouraging increased efforts at reproducibility would also help make methods more accessible to new users, by providing fully worked examples that can act as a starting point for their analyses.
+    Reproducibility can be achieved by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
     At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).
@@
 
-    The glut of literature that already exists for established methods is a bigger problem.
+    Summarising the glut of literature that already exists for established methods is a more difficult problem to solve.
     The Oxford Bibliographies Evolutionary biology pages have lists of key papers (http://www.oxfordbibliographies.com/obo/page/evolutionary-biology), but these are curated by just a few individuals and tend to contain a lot of papers.
@@
     The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
-    Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) is an excellent example of a similar approach for phylogenetics. 
+    Two excellent examples of a similar approach are Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) for phylogenetics, and the Prometheus wiki for protocols in plant physiology (http://prometheuswiki.publish.csiro.au/tiki-custom\_home.php).
     The British Ecological Society's Quantitative Ecology Special Interest Group is also in the process of creating a ``Field Guide for Ecologists'' (http://bes-qsig.github.io/fge/) that will fill a similar niche in ecological methods.
-    Unfortunately there is currently little incentive for people to contribute to such a project at the cost of time pursuing their own research.\\
+    We plan to establish a similar guide for PCMs in the near future with help from across the community.\\
+    %Unfortunately there is currently little incentive for people to contribute to such a project at the cost of time pursuing their own research.\\
 
-  \subsubsection{2. Be more cynical about existing methods}
+  \subsubsection{2. Factors to consider before using a method}
     A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
-    As the Manic Street Preachers put it, ``Cynicism is the only thing that keeps me sane''.
-    At a minimum, it is worth reading the original papers describing a method, plus any recent updates, and looking carefully for assumptions and caveats that may affect the analyses at hand.
+    As the Manic Street Preachers (\citeyear{manics}) put it, ``Cynicism is the only thing that keeps me sane''.
+    At a minimum, users should read the original papers describing a method, plus any recent updates, and look carefully for assumptions and caveats that may affect the analyses at hand.
     A good way to check a method is to simulate some data and see if the results are as expected \citep[e.g.][]{boettiger2012your}. 
     This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
-    It is also important to consider whether the method will work on a particular dataset.
-    For example it is worth considering how many species are required for reasonable power. 
+    It is also important to determine whether the method will work on a particular dataset.
+    One key consideration is how many species are required for reasonable power. 
     Often methods require more species than are usually available.
@@
     Worked examples and explanations are provided in \citep{slater2014correction}. 
-Although this is not a problem with applying the OU model to non-ultrametric trees per se, it is an example of different implementations of a common model that some users may not be aware of.
+    Although this is not a problem with applying the OU model to non-ultrametric trees \textit{per se}, it is an example of different implementations of a common model that some users may not be aware of.
 
     It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
+    In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions.
     Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore.
 
-  \subsubsection{3. Incentives}
+  \subsubsection{3. Solutions and incentives}
     It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
@@
     This is fairly standard in other fields, e.g. statistical phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
-    When novel methods are published, journals should encourage researchers to include ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.
+    When novel methods are published, journals should encourage researchers to include lists and ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.
 
     Second we need to fund pure methods development, including incremental methods. 
-    Currently it is difficult to get funding for purely methods driven research; an empirical component is generally needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
-    This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.
+    Currently it is difficult to get funding for purely methods driven research; an empirical component is generally needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse (although it is also hard to get funding for empirical projects).
+    %This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.
 
@@
     Some benefits are obvious, e.g.
-methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released, whereas end-users can work with the most cutting-edge methods and software.
+methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it gets released, whereas end-users can work with the most cutting-edge methods and software.
     Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.
@@
     New methods are being published with increasing regularity, and we are also beginning to question older methods and classical ways of looking at comparative questions. 
-    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups, and even continents, becoming more common.
+    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups, and even continents, is becoming more common.
     However, while embracing these changes, we also need to ensure that we do not forget that PCMs have assumptions, caveats and biases like every other method. 

### 2015-12-08 Manuscript/dark.tex
Dealing with SB's comments

@@
 \section{Abstract}
-
 \begin{enumerate}
@@
   \item To help solve these problems we make a number of suggestions including providing
-  blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
+  blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods and ways of detecting biases or testing model fit.
   Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
@@
     A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
-    It is not unusual for papers to be long, tedious and full of equations. 
-    Of course, such detail and mathematical literacy is critical in describing methods and facilitates testing/assessment, implementation and future developments of the method.
-    In most cases equations could not simply be replaced with text.
-   However, from the end user perspective it may more helpful to understand the assumptions and caveats.
-   Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
-    They are rarely neatly corralled in one place.
-     The sheer volume of literature can become a barrier to understanding, even, and perhaps especially, for the best known methods.
+    It is not unusual for papers to be long and full of equations. 
+    Of course, such detail is critical for describing methods and facilitates testing/assessment, implementation and future developments of the method.
+    Additionally, in most cases equations could not simply be replaced with text.
+    There is evidence, however, that heavy use of equations impedes understanding and communication of concepts in biology \citep{fawcett2012heavy}.\\
+
+    Another issue is that end-users need to understand the assumptions and caveats of methods.
+    Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper - they are rarely neatly corralled in one place.
+    The sheer volume of literature can thus become a barrier to understanding, even, and perhaps especially, for the best known methods.
     For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
-    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
-    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may be easy for experienced users but not clear for new users.
-    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details.
-    Many of these problems relate to due diligence for both reader and developer. 
-    Methods developers are not responsible for making sure that all users read the literature. 
-    The onus is, quite rightly in our opinion, on the reader/user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
-    However, simple steps, such as subheadings that point to caveats and assumptions, could add clarity and limit method misuse.
+    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read in order to fully understand the method.
+    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may not be clear to all users.\\
+
+    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details in PCM papers.
+    These problems relate to due diligence for both end-users and developers. 
+    Methods developers are not responsible for making sure that end-users read the literature. 
+    Instead the onus is on the end-user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
+    However, simple steps could be taken by methods developers, such as subheadings that point to caveats and assumptions, to add clarity and limit method misuse and misinterpretation.
 
@@
     Indeed, \citet{freckleton2009seven} suggested that the ability to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
-    However, this has instead led to an increasing number of people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
-    This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
-    Manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
-    The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).
+    However, the increasing use of R has instead led to more people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
+    %This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
+    This is not the fault of methods developers and users should conduct due diligence in understanding a method before using it. 
+    Unfortunately the problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
+    The APE book \citep[an excellent resource;][]{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see \texttt{CAIC} and \texttt{caper} documentation; \citealp{purvis1995comparative,Orme:2013aa}).
 
@@
     As we note above, the onus is on the reader to read extensively and to do due diligence.
-    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target readers, especially in describing assumptions and limitations.
-    One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 
+    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target readers, especially in describing assumptions and limitations.\\
+
+    One solution is to prepare an accompanying blog post or video for each new method, explaining it in less technical terms. 
     Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there are great benefits to doing this anyway and hosting it on a personal website.
-    Some package developers provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. 
     The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
+    Some package developers also provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. \\
+
     Encouraging increased efforts at reproducibility would also help make methods more accessible to new users, by providing fully worked examples that can act as a starting point for their analyses.
-    Reproducibility can be achieved by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
-    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).
+    Reproducibility can be achieved by including \texttt{knitR} \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
+    At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).\\
+
     An important point here is that we do not see these possible solutions as entirely the responsibility of the method developer or package writer. 
-    The comparative methods community can help too and indeed often does via platforms such as r-sig-phylo.
-    However, the ability to find helpful replies to lists depends on the subject headings and details in the text.
-    A longer lasting solution would be to make published methods papers, and especially those that have a software focus, more readily updatable. 
-    It should not be necessary to publish a new paper for updates of method.
-    One example that gets around this problem is PLoS Currents where published papers can be updated relatively quickly with the full history of updated articles versioned.
+    The comparative methods community can help too and indeed often does via platforms such as \texttt{r-sig-phylo}.
+    However, the ability to find helpful replies on a listserv depends on the subject headings and details in the text.
+    A longer lasting solution would be to make published methods papers, and especially those that have a software focus, more readily updatable, rather than requiring a new paper for each update.
+    One example that gets around this problem is PLoS Currents (currents.plos.org) where published papers can be updated relatively quickly with the full history of updated articles versioned.\\
 
@@
 MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix.
-    The problem is that where at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical. 
+    The problem is that where there is a pair of tips and at least one tip does not survive to the present, the expected covariances relating each of those two tips with any other tip in the tree are not identical.
     Worked examples and explanations are provided in \citep{slater2014correction}. 
@@
   \subsubsection{3. Solutions and incentives}
-    It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
+    It is important to recognise that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
     We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for such skilled researchers to remain in (or at least interact with) academia.

### 2015-12-08 Manuscript/response.tex
Response to SB's comments

@@
 This was a tricky paper to write, and we debated heavily about whether to write it at all.
-Finding the right tone was indeed very difficult, and we clearly haven't completely succeeded there.
+Finding the right tone was particularly difficult, and we clearly haven't completely succeeded there.
 Also while we had started to "Do Something" about some of this around the time I first presented this, in the interim both Rich and I have moved countries and jobs (Rich out of academia) delaying some projects and derailing others completely!
 
-\textcolor{blue}{\textbf{Bob O'Hara's comments}}\\
+\textcolor{blue}{\textbf{Bob O'Hara's comments}}
 
@@
 
-We have tried to correct the tone where we can. It's very tricky!
+We have tried to correct the tone where we can. In particular we have edited pages 11-13, page 14 paragraph 1 and page 15 paragraph 2 (line 333).
 
@@
 
-\textcolor{blue}{\textbf{Referee 1: Amy Zanne}}\\
+\textcolor{blue}{\textbf{Referee 1: Amy Zanne}}
 
@@
 
-***We agree and make exactly these points on line XX: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles." However to make this clearer we have specified that it "tends to be passed down from PIs to graduate students, and within labs developing these methods".
+We agree and make exactly these points on line 187-88: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles." However to make this clearer we have specified that it "tends to be passed down from PIs to graduate students, and within labs developing these methods".
 We also have been more explicit that sometimes it is based on "best practice".
@@
 
-****This is a very good point
+This is a very good point and we have modified this section to reflect this and in response to Simon's comments below.
 
@@
 We have changed the subheading to "Factors to consider before using a method" to more accurately reflect its content.
-We have also rephrased the section to reduce the number of considers and for examples.
+We have also rewritten parts of the section to reduce the number of considers and for examples.
 
@@
 Agreed.
-This can be a big problem so we have may this point more explicitly: "In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions."
+This can be a big problem so we have made this point more explicitly: "In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions."
 
@@
 
-The point we were trying to make is that methods development is hard to get funded, thus your CV lacks big grants making it harder to secure faculty positions/tenure.
-however, we recognise that is was ambiguous, and also needs action at all levels.
+The point we were trying to make is that methods development is hard to get funded, thus developer's CVs lacs big grants making it harder to secure Faculty positions/tenure.
+However, we recognise that is was ambiguous, and also needs action at all levels.
 We have therefore deleted this line.
@@
 \textcolor{blue}{After reading the abstract of this paper, I was sceptical. However, upon reading it I found that it makes some reasonable points. First I will discuss the negative aspects.}
+
+Thanks for these useful comments.
+We have tried to clear up the misunderstandings and issues outlined below.
+In particular we have edited pages 11-13, page 14 paragraph 1 and page 15 paragraph 2 (line 333).
 
@@
 This is a symptom of how papers are published but can very easily mislead novice users.
-Clarity in writing is the responsibility of the write, thorough reading is the responsibility of the reader.
+Clarity in writing is the responsibility of the writer, thorough reading is the responsibility of the reader.
+
+However, while we sympathise with the points made here and were not aiming to castigate authors (apologies if it comes across like that, we have tried to rephrase), it isn't possible for everyone who needs to use quantitative methods to be an expert in their use.
+Not every university has access to a statistician, and even those with one may not have an expert in PCMs.
+It is also virtually impossible to do any science these days that is not quantitative, so simply avoiding quantitative methods entirely is not an option.
+A lot of problems also come from situations where reviewers (or supervisors) tell people to use a specific method they may not have expertise in.
+This is a tricky issue we agree.
 
@@
 
-We agree.
+We agree, and were not in any way suggesting this was the fault of developers (apologies if that is how it sounded, we have changed the writing here).
 However, as we note above we also argue that doing due diligence is a responsibility of both the users (in reading widely and deeply) and the developers (in writing clearly and stating assumptions, limitations etc).
+We are all big fans and promoters of using R and non-proprietary software in general.
+But again, it is difficult to expect someone who has been asked to apply a method by a reviewer to understand source code for a complicated method.
+Particularly since they often require you to understand C or C++ as well as R.
+This is a particularly difficult expectation for graduate students.
 
@@
 
-We absolutely agree that readers must do the work to understand methods but reiterate the point above - due diligence is a responsibility for developers as well as users.
+We absolutely agree that readers must do the work to understand methods but reiterate the point above - due diligence is a responsibility for developers as well as users, and not everyone can be an expert in every single method they use.
 We have revised lines 277-279 accordingly.
@@
 
+We were not meaning to castigate developers here, and apologise if it came across like this.
+We have tried to rephrase.
+While we agree that people who understand the methods enough to use them but do not develop methods regularly, could help out by teaching courses (indeed all of the authors of the paper are involved in such ventures) and writing documentation (again something we have all been involved with), it is not feasible for new users to do this as they learn the method themselves.
+We are also aware of the time constraints faced by methods developers, and the great service they do to the community by providing their methods, this is why we mention incentives near the end of the paper.
+
 \textcolor{blue}{The authors encourage more reproducible research and open source science (lines 287-290). This is a growing movement and will no doubt help raise the quality of research.}
@@
 They would be most welcome.}
+
+Thanks for these positive comments.
 
@@
 
-This was an unfortunate error on our part. Of course the variance is not constant in the Brownian motion model. We have replaced the two instances of "constant variance" with "motion".
+Yes, of course the variance is not constant in the Brownian motion model.
+We have replaced the two instances of "constant variance" with "motion".
+Often however it is referred to as the "Brownian constant variance model" with "constant variance" implying that the expected variance of the changes is constant.
 
@@
 
+Agreed.
+
 \textcolor{blue}{line 339-340: It's just not true that, "non-ultrametric trees cause variance-covariance matrices to become non-tree like".
@@
 
-We believe that this is a misunderstanding of our point and reflects a lack of clarity in our writing. We now write:
+We believe that this is a misunderstanding of our point and reflects a lack of clarity in our writing. We now write (p17, line 388):
 
-"Some current implementations of the OU model should not be used with non-ultrametric trees (e.g.
-MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix.
-The problem is that where at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical.
-Worked examples and explanations are provided in \citep{slater2014correction}.
-Although this is not a problem with applying the OU model to non-ultrametric trees per se, it is an example of different implementations of a common model that some users may not be aware of.
-"
+``Some current implementations of the OU model should not be used with non-ultrametric trees (e.g.
+MOTMOT; Thomas et al.
+2011) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix.
+The problem is that where there is a pair of tips and at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical.
+Worked examples and explanations are provided in Slater (2014).
+Although this is not a problem with applying the OU model to non-ultrametric trees \textit{per se}, it is an example of different implementations of a common model that some users may not be aware of.''
 
-Let us know if you require any further information,
+Please let us know if you require any further information,
 

### 2015-12-09 Manuscript/dark.tex
Updated wordcount
Minor edits to manuscript pre-submission

@@
 \keywords{PCM, assumption, bias, caveat, phylogenetic independent contrasts, Ornstein-Uhlenbeck, trait-dependent diversification\\
-\textbf{Word count}: 3532 (including references, tables and figure legends)}
+\textbf{Word count}: 4108 (including references, tables and figure legends)}
 
@@
 The phylogenetic independent contrasts method uses phylogenetic information to account for the fact that species in a comparative analysis are related to each other and thus may share similarities because they inherit them from their ancestors, not because of independent evolution \citep{felsenstein1985phylogenies,harvey1991comparative}.
-This is the most commonly used phylogenetic comparative method (\citet{felsenstein1985phylogenies} has been cited nearly 6000 times; Google Scholar search 10th October 2015), and a great deal of literature exists on the assumptions underlying the method, and ways of testing whether these assumptions are met.
+This is the most commonly used phylogenetic comparative method (\citet{felsenstein1985phylogenies} has been cited over 6000 times; Google Scholar search 9th December 2015), and a great deal of literature exists on the assumptions underlying the method, and ways of testing whether these assumptions are met.
 The method has three major assumptions (\citealp{diaz1996testing}; although there are many more assumptions than these three), (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian motion model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
@@
 These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{Goldberg22102010,Price01052012,Givnish201455,ELE:ELE12168}.
-Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSSE paper \citet{maddison2007estimating} has been cited 384 times; Google Scholar search 10th October 2015).
+Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSSE paper \citet{maddison2007estimating} has been cited 394 times; Google Scholar search 9th December 2015).
 However, \citet{rabosky2015model} recently reevaluated the method and its assumptions and showed via simulations that a strong correlation between a trait and diversification rate can be inferred from a single diversification rate shift within a tree, even if the shift is unrelated to the trait of interest.
@@
     Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
-    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing these methods (and it is occasionally closer to fiction than fact!).
+    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing methods (and it is occasionally closer to fiction than fact!).
     Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules.
@@
     For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
-    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read in order to fully understand the method.
+    Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read in order to fully understand the method and its caveats.
     The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may not be clear to all users.\\
@@
     This resulted in many frustrating hours (and days and months) trying to implement any new method you wanted to use.
-    Writing your own implementation may be the best way to learn the intricacies of a method but is a major hurdle and would necessarily dissuade many potential users.  \\
+    Writing your own implementation may be the best way to learn the intricacies of a method but is a major hurdle and can dissuade many potential users.  \\
 
@@
     It is available to all, has a wide and engaged support community, and perhaps most importantly the source code for new methods is accessible.
-    Users can, if they choose to, fully explore any new method by examining the source code and running their own simulations. 
+    Users can fully explore any new method by examining the source code and running their own simulations. 
     Indeed, \citet{freckleton2009seven} suggested that the ability to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
@@
   \subsubsection{1. Simplify, summarise and share}
-    Many of the problems above are the result of methods that are difficult to communicate in purely descriptive terms.
+    Many of the problems above arise because PCMs are difficult to communicate in purely descriptive terms.
     This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text.
-    As we note above, the onus is on the reader to read extensively and to do due diligence.
-    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target readers, especially in describing assumptions and limitations.\\
+    As we note above, the onus is on the end-user to read extensively and to do due diligence.
+    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target users, especially in describing assumptions and limitations.\\
 
@@
     The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
-    Some package developers also provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. \\
+    Some package developers also provide blog updates (e.g. \texttt{phytools}: http://blog.phytools.org; \citealp{revell2012phytools}) that serve this function. \\
 
@@
     An important point here is that we do not see these possible solutions as entirely the responsibility of the method developer or package writer. 
-    The comparative methods community can help too and indeed often does via platforms such as \texttt{r-sig-phylo}.
+    The PCM community can help too and indeed often does via platforms such as \texttt{r-sig-phylo} (http://www.mail-archive.com/r-sig-phylo@r-project.org/).
     However, the ability to find helpful replies on a listserv depends on the subject headings and details in the text.
-    A longer lasting solution would be to make published methods papers, and especially those that have a software focus, more readily updatable, rather than requiring a new paper for each update.
-    One example that gets around this problem is PLoS Currents (currents.plos.org) where published papers can be updated relatively quickly with the full history of updated articles versioned.\\
+    A longer lasting solution would be to make published methods papers, especially those that have a software focus, more readily updatable, rather than requiring a new paper for each update.
+    One example that gets around this problem is PLoS Currents (http://currents.plos.org) where published papers can be updated relatively quickly with the full history of updated articles versioned.\\
 
@@
     The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
-    Two excellent examples of a similar approach are Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) for phylogenetics, and the Prometheus wiki for protocols in plant physiology (http://prometheuswiki.publish.csiro.au/tiki-custom\_home.php).
+    Two excellent examples of a similar approach are Erick Matsen's Phylobabble discourse page for phylogenetics (http://phylobabble.org/), and the Prometheus wiki for protocols in plant physiology (http://prometheuswiki.publish.csiro.au/tiki-custom\_home.php).
     The British Ecological Society's Quantitative Ecology Special Interest Group is also in the process of creating a ``Field Guide for Ecologists'' (http://bes-qsig.github.io/fge/) that will fill a similar niche in ecological methods.
@@
 
-\bibliographystyle{mee}
+\bibliographystyle{mee2}
 \bibliography{darkside}

### 2015-12-23 Manuscript/dark.tex
Add citation for data in figshare

--- 2015-12-08 Manuscript/dark.tex
@@
 
-\begin{enumerate}
-  \item Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
+Phylogenetic comparative methods are becoming increasingly popular for investigating evolutionary patterns and processes.
 However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
-  \item Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
+Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
 Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
-  \item We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
+We suggest that some important pieces of information are missing from the literature, and that others are difficult to extract from long, technical papers.
 We also highlight problems with users jumping straight into software implementations of methods (e.g.
 in R) that may lack documentation on biases and assumptions that are mentioned in the original papers.
-  \item To help solve these problems we make a number of suggestions including providing
+To help solve these problems we make a number of suggestions including providing
 blog posts or videos to explain new methods in less technical terms, encouraging reproducibility and code sharing, making wiki-style pages summarising the literature on popular methods, more careful consideration and testing of whether a method is appropriate for a given question/dataset, increased collaboration, and a shift from publishing purely novel methods to publishing improvements to existing methods, and ways of detecting biases or testing model fit.
 Many of these points are applicable across methods in ecology and evolution, not just phylogenetic comparative methods.
-\end{enumerate}
 
@@
 This is the most commonly used phylogenetic comparative method (\citet{felsenstein1985phylogenies} has been cited nearly 6000 times; Google Scholar search 10th October 2015), and a great deal of literature exists on the assumptions underlying the method, and ways of testing whether these assumptions are met.
-The method has three major assumptions (\citealp{diaz1996testing}; although there are many more assumptions than these three), (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian motion model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
+The method has three major assumptions (\citealp{diaz1996testing}; although there are many more assumptions than these three), (1) that the topology of the phylogeny is accurate; (2) that the branch lengths of the phylogeny are correct; and (3) that traits evolve in the manner of the Brownian constant variance model, a simple model of trait evolution where trait variance accrues as a linear function of time \citep{cavalli1967,felsenstein1973maximum}.
 The third assumption is stated in \citet{felsenstein1985phylogenies} though the other two are not explicitly mentioned. 
@@
 \subsubsection{Ornstein-Uhlenbeck (single stationary peak) models of trait evolution}
-Most models of trait evolution are based on the Brownian motion model \citep{cavalli1967,felsenstein1973maximum}.
+Most models of trait evolution are based on the Brownian constant variance model \citep{cavalli1967,felsenstein1973maximum}.
 The Ornstein-Uhlenbeck (OU) model can be thought of as a modification of the Brownian model with an additional parameter that measures the strength of return towards a theoretical optimum shared across a clade or subset of species \citep{hansen1997stabilizing,Butler:2004aa}.
@@
     Unfortunately, some of the information needed to properly apply PCMs is not found in the literature. 
-    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing these methods (and it is occasionally closer to fiction than fact!).
-    Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules.
+    We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students (and it is occasionally closer to fiction than fact!).
+    Sometimes the folklore includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules. 
     Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
@@
     A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
-    It is not unusual for papers to be long, tedious and full of equations. 
-    Of course, such detail and mathematical literacy is critical in describing methods and facilitates testing/assessment, implementation and future developments of the method.
-    In most cases equations could not simply be replaced with text.
-   However, from the end user perspective it may more helpful to understand the assumptions and caveats.
-   Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
-    They are rarely neatly corralled in one place.
-     The sheer volume of literature can become a barrier to understanding, even, and perhaps especially, for the best known methods.
+    It is not unusual for papers to be full of equations, \textgreater 20 pages long, and written in a way that gives one a headache. 
+    Additionally there is a lot of literature to wade through before getting all of the required information. 
     For example, we recently reviewed papers discussing the assumptions and limitations of phylogenetic independent contrasts \citep{felsenstein1985phylogenies}, including how they are related to phylogenetic generalised least squares models \citep{garland2000using,rohle2006comment,blomberg2012independent}.
     Even with prior knowledge of the key papers and authors to focus on, this resulted in $\approx 300$ manuscript pages and a book to read.
-    The volume of reading itself is perhaps not the key issue, rather it is assessing when you have reached sufficient understanding which may be easy for experienced users but not clear for new users.
-    The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details.
-    Many of these problems relate to due diligence for both reader and developer. 
-    Methods developers are not responsible for making sure that all users read the literature. 
-    The onus is, quite rightly in our opinion, on the reader/user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
-    However, simple steps, such as subheadings that point to caveats and assumptions, could add clarity and limit method misuse.
+    This seems like an excessive amount of reading to run a simple empirical analysis.
+    Another issue is that it can sometimes be hard to find the information required. 
+    Assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper. 
+    They are rarely neatly corralled in one place, making it easy to miss pertinent details.
 
@@
 MATLAB code; \citealp{rohlf2001comparative}), and still others provided no way of implementing their methods at all.
-    This resulted in many frustrating hours (and days and months) trying to implement any new method you wanted to use.
-    Writing your own implementation may be the best way to learn the intricacies of a method but is a major hurdle and would necessarily dissuade many potential users.  \\
+    This resulted in many frustrating hours (and days and months) trying to implement any new method you wanted to use.\\
 
@@
     Simultaneously, more people are able to use R thanks to workshops and changes in student training, thus when a new method is published it is now possible to take an R package ``off the shelf'' and use it to run the method immediately.
-    The benefits of R are clear. 
-    It is available to all, has a wide and engaged support community, and perhaps most importantly the source code for new methods is accessible.
-    Users can, if they choose to, fully explore any new method by examining the source code and running their own simulations. 
-    Indeed, \citet{freckleton2009seven} suggested that the ability to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
+    \citet{freckleton2009seven} suggested that the ability to to conduct PCMs in flexible computing environments such as R would improve our ability to implement methods correctly. 
     However, this has instead led to an increasing number of people (including the authors of this paper) jumping straight to the implementation of a method, without fully understanding what the method is doing, what its assumptions might be, or what the results mean in a biological context.
-    This is again an example of where users can and should conduct due diligence but also where developers and experienced users can help.
-    Manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
+    This is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
     The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citealp{purvis1995comparative,Orme:2013aa}).
@@
   \subsubsection{1. Simplify, summarise and share}
-    Many of the problems above are the result of methods that are difficult to communicate in purely descriptive terms.
-    This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text.
-    As we note above, the onus is on the reader to read extensively and to do due diligence.
-    Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target readers, especially in describing assumptions and limitations.
+    Many of the problems above are caused by methods papers being difficult to read. 
+    This is not entirely the fault of those writing these papers; most journals have strict word limits so it is often easier to use a one line equation rather than a paragraph of description. 
+    However, it is hard to complain about novice users failing to understand a method when the paper is incomprehensible.
     One solution is to prepare an accompanying blog post or video explaining the method in less technical terms. 
-    Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there are great benefits to doing this anyway and hosting it on a personal website.
-    Some package developers provide blog updates (e.g. phytools http://blog.phytools.org) that serve this function. 
+    Some journals already encourage this (including the British Ecological Society journals; see https://soundcloud.com/besjournals, https://www.youtube.com/user/MethodsEcolEvol), and even if they do not, there is no harm in doing this and hosting it on a personal website.
     The ability to share ideas with a non-technical audience is a key skill to develop, and may encourage more people to use the method.
-    Encouraging increased efforts at reproducibility would also help make methods more accessible to new users, by providing fully worked examples that can act as a starting point for their analyses.
-    Reproducibility can be achieved by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
+    Encouraging increased efforts at reproducibility would also help, by including knitR \citep{Xie:2015aa} reports in Supplementary Material showing exactly how each analysis was run \citep[e.g.][]{fitzjohn2014much}, or by requiring all analyses and code to be available on GitHub or Bitbucket.
     At the very least, having a list of the main assumptions, biases and caveats of the method somewhere obvious would reduce misuse and provide a place to point people to when they begin using a method (this will be difficult as there are often hidden assumptions in a method, and listing all possible assumptions and caveats may be unfeasible in some cases).
-    An important point here is that we do not see these possible solutions as entirely the responsibility of the method developer or package writer. 
-    The comparative methods community can help too and indeed often does via platforms such as r-sig-phylo.
-    However, the ability to find helpful replies to lists depends on the subject headings and details in the text.
-    A longer lasting solution would be to make published methods papers, and especially those that have a software focus, more readily updatable. 
-    It should not be necessary to publish a new paper for updates of method.
-    One example that gets around this problem is PLoS Currents where published papers can be updated relatively quickly with the full history of updated articles versioned.
 
-    Summarising the glut of literature that already exists for established methods is a more difficult problem to solve.
+    The glut of literature that already exists for established methods is a bigger problem.
     The Oxford Bibliographies Evolutionary biology pages have lists of key papers (http://www.oxfordbibliographies.com/obo/page/evolutionary-biology), but these are curated by just a few individuals and tend to contain a lot of papers.
@@
     The community would be responsible for peer reviewing these summaries to ensure all opinions are covered. 
-    Two excellent examples of a similar approach are Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) for phylogenetics, and the Prometheus wiki for protocols in plant physiology (http://prometheuswiki.publish.csiro.au/tiki-custom\_home.php).
+    Erick Matsen's Phylobabble discourse page (http://phylobabble.org/) is an excellent example of a similar approach for phylogenetics. 
     The British Ecological Society's Quantitative Ecology Special Interest Group is also in the process of creating a ``Field Guide for Ecologists'' (http://bes-qsig.github.io/fge/) that will fill a similar niche in ecological methods.
-    We plan to establish a similar guide for PCMs in the near future with help from across the community.\\
-    %Unfortunately there is currently little incentive for people to contribute to such a project at the cost of time pursuing their own research.\\
+    Unfortunately there is currently little incentive for people to contribute to such a project at the cost of time pursuing their own research.\\
 
-  \subsubsection{2. Factors to consider before using a method}
+  \subsubsection{2. Be more cynical about existing methods}
     A key skill to develop in science is cynicism, i.e., never take results from PCMs (or any other statistical analysis) at face value.
-    As the Manic Street Preachers (\citeyear{manics}) put it, ``Cynicism is the only thing that keeps me sane''.
-    At a minimum, users should read the original papers describing a method, plus any recent updates, and look carefully for assumptions and caveats that may affect the analyses at hand.
+    As the Manic Street Preachers put it, ``Cynicism is the only thing that keeps me sane''.
+    At a minimum, it is worth reading the original papers describing a method, plus any recent updates, and looking carefully for assumptions and caveats that may affect the analyses at hand.
     A good way to check a method is to simulate some data and see if the results are as expected \citep[e.g.][]{boettiger2012your}. 
     This can expose hidden assumptions or biases that have not been explored in the papers accompanying the method, or reveal a lack of understanding of the the mechanics of the method being used. 
-    It is also important to determine whether the method will work on a particular dataset.
-    One key consideration is how many species are required for reasonable power. 
+    It is also important to consider whether the method will work on a particular dataset.
+    For example it is worth considering how many species are required for reasonable power. 
     Often methods require more species than are usually  available.
@@
     Some current implementations of the OU model should not be used with non-ultrametric trees (e.g.
-MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix.
-    The problem is that where at least one tip does not survive to the present the expected covariances relating each of those two tips with any other tip in the tree are not identical. 
-    Worked examples and explanations are provided in \citep{slater2014correction}. 
-    Although this is not a problem with applying the OU model to non-ultrametric trees \textit{per se}, it is an example of different implementations of a common model that some users may not be aware of.
+MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix and non-ultrametric trees cause variance-covariance matrices to become non-tree like \citep{slater2014correction}.
 
     It is also important to avoid retrofitting questions to the newest methods, instead we should think carefully about the question, whether the method is appropriate for the question, and whether PCMs are needed at all \citep{westoby1995misinterpreting,losos2011seeing}.
-    In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions.
     Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore.
 
-  \subsubsection{3. Solutions and incentives}
+  \subsubsection{3. Incentives}
     It is important to recognize that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
@@
     This is fairly standard in other fields, e.g. statistical phylogenetics, and these kinds of papers are arguably more useful to the community than constantly publishing new methods.
-    When novel methods are published, journals should encourage researchers to include lists and ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.
+    When novel methods are published, journals should encourage researchers to include ways of testing the assumptions of their methods within the original publications and packages, and request simplified summaries to accompany technical papers.
 
     Second we need to fund pure methods development, including incremental methods. 
-    Currently it is difficult to get funding for purely methods driven research; an empirical component is generally needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse (although it is also hard to get funding for empirical projects).
-    %This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.
+    Currently it is difficult to get funding for purely methods driven research; an empirical component is generally needed and methods development is often seen as part of the bigger empirical picture, rather than the reverse.
+    This unfairly prejudices funding bodies (and hiring committees) away from this kind of research.
 
@@
     Some benefits are obvious, e.g.
-methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it gets released, whereas end-users can work with the most cutting-edge methods and software.
+methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it get released, whereas end-users can work with the most cutting-edge methods and software.
     Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.
@@
     New methods are being published with increasing regularity, and we are also beginning to question older methods and classical ways of looking at comparative questions. 
-    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups, and even continents, is becoming more common.
+    In addition, the field is becoming more open, with code being shared before analyses are even submitted for publication, and collaborative software development across groups, and even continents, becoming more common.
     However, while embracing these changes, we also need to ensure that we do not forget that PCMs have assumptions, caveats and biases like every other method. 
@@
   \subsection{Data Accessibility}
-    All data are available on GitHub at https://github.com/nhcooper123/pcm-darkside.
+    All data from Figures 1 and 2 are available on figshare: https://dx.doi.org/10.6084/m9.figshare.2057802.v1.
+R code for recreating the figures is available on GitHub: https://github.com/nhcooper123/pcm-darkside.
 
@@
       \includegraphics[width=12cm]{Figures/PCMCitations.pdf}
-      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).}
+      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).
+Data available from figshare \citep{coopermeedata}.}
       \label{PCMCitations}
@@
       \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time from 1980-2014.
-Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.}
+Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.
+Data available from figshare \citep{coopermeedata}.}
       \label{PCMRpackages}

### 2015-12-23 Manuscript/response.tex

@@
 
-We agree and make exactly these points on line 187-88: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles." However to make this clearer we have specified that it "tends to be passed down from PIs to graduate students, and within labs developing these methods".
+We agree and make exactly these points on line 187-88: "Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles." However to make this clearer we have specified that it "tends to be passed down from PIs to graduate students, and within labs developing methods".
 We also have been more explicit that sometimes it is based on "best practice".
@@
 
-The point we were trying to make is that methods development is hard to get funded, thus developer's CVs lacs big grants making it harder to secure Faculty positions/tenure.
+The point we were trying to make is that methods development is hard to get funded, thus developer's CVs lack big grants making it harder to secure Faculty positions/tenure.
 However, we recognise that is was ambiguous, and also needs action at all levels.

### 2015-12-23 Manuscript/dark2.tex
added figshare link

--- 2015-12-09 Manuscript/dark.tex
@@
       \includegraphics[width=12cm]{Figures/PCMCitations.pdf}
-      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).}
+      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).
+Data available from figshare \citep{coopermeedata}.
+}
       \label{PCMCitations}
@@
       \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time from 1980-2014.
-Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.}
+Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.
+Data available from figshare \citep{coopermeedata}.
+}
       \label{PCMRpackages}

### 2015-12-23 Manuscript/dark2.tex
Fixed BOH's comments

@@
   However, these methods are not infallible - they suffer from biases and make assumptions like all other statistical methods.
-  \item Unfortunately, although these limitations are generally well-known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
+  \item Unfortunately, although these limitations are generally well known in the phylogenetic comparative methods community, they are often inadequately assessed in empirical studies leading to misinterpreted results and poor model fits.
   Here we explore reasons for the communication gap dividing those developing new methods and those using them. 
@@
 
-We suggest that one cause of this problem is that although researchers developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to end-users.
+We suggest that one cause of this problem is that although researchers developing and implementing new methods are aware of the limitations of their methods and the assumptions that underly them, this information is not always being effectively transferred to end users.
 Additionally, the tools and approaches used to fit models are often far more user-friendly and better documented than the methods used to assess whether that model fit is reasonable. 
-Clearly more effort is needed to bridge the widening gap between those developing methods and end-users. 
+Clearly more effort is needed to bridge the widening gap between those developing methods and end users. 
 Here we explore the causes of this communication gap and suggest some potential solutions.
@@
 For example, fit to an OU model has been seen as evidence of evolutionary constraints, stabilising selection, niche conservatism, and selective regimes \citep{Wiens:2010aa,beaulieu2012modeling,christin2013anatomical,mahler2013exceptional}.
-However, the OU model has several well-known caveats \citep[see][]{ives2010phylogenetic,boettiger2012your,hansen2012interpreting,ho2013asymptotic,ho2014intrinsic}. 
+However, the OU model has several well known caveats \citep[see][]{ives2010phylogenetic,boettiger2012your,hansen2012interpreting,ho2013asymptotic,ho2014intrinsic}. 
 For example, it is frequently incorrectly favoured over simpler models when using Likelihood ratio tests, particularly for small datasets that are commonly used in these analyses (the median number of taxa used for OU studies is 58; \citealp{cooper2016}).
@@
 These kinds of analyses have been applied extensively in recent years to a variety of taxa and traits \citep[e.g.][]{Goldberg22102010,Price01052012,Givnish201455,ELE:ELE12168}.
-Most use the binary-state speciation and extinction model (BiSSE) and related methods (the original BiSSE paper \citet{maddison2007estimating} has been cited 394 times; Google Scholar search 9th December 2015).
+Most use the binary state speciation and extinction model (BiSSE) and related methods (the original BiSSE paper \citet{maddison2007estimating} has been cited 394 times; Google Scholar search 9th December 2015).
 However, \citet{rabosky2015model} recently reevaluated the method and its assumptions and showed via simulations that a strong correlation between a trait and diversification rate can be inferred from a single diversification rate shift within a tree, even if the shift is unrelated to the trait of interest.
@@
     We refer to this knowledge as ``PCM folklore'' because it tends to be passed down from PIs to graduate students, and within labs developing methods (and it is occasionally closer to fiction than fact!).
-    Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules-of-thumb; other times it is more opinion-based, but over time these opinions become rules.
+    Sometimes the folklore is based on best practice and includes tricks to get methods working, or useful rules of thumb; other times it is more opinion based, but over time these opinions become rules.
     Useful PCM folklore is often shared among developers, and among collaborating groups, but is not always shared outside of these circles. 
@@
     One example of PCM folklore is that species with Studentized residuals $\pm 3$ are often omitted from regressions of phylogenetic independent contrasts, to avoid highly influential outliers affecting the results \citep[e.g.][]{GEB:GEB355}.
-    The rationale for this comes from \citet{jones1997optimum}, however, the $\pm 3$ cut-off is arbitrary, and barely mentioned in the original paper, but has become a rule-of-thumb for running these analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
+    The rationale for this comes from \citet{jones1997optimum}, however, the $\pm 3$ cut-off is arbitrary, and barely mentioned in the original paper, but has become a rule of thumb for running these analyses (the paper has been cited \textgreater 100 times, mostly as a justification of this procedure).
     Another example is in the defaults of programs that perform PCMs. 
@@
 
-    Other information about the limitations of a method may be absent from the literature due to the time-lag between a new method being published and others having time to test it. 
+    Other information about the limitations of a method may be absent from the literature due to the time lag between a new method being published and others having time to test it. 
     For example, Felsenstein published the phylogenetic independent contrasts method in 1985 \citep{felsenstein1985phylogenies}, but it was not until the early 1990s that thorough critiques of the method and its assumptions began to be published \citep[e.g.][]{garland1992procedures}.
-    This time-lag is shorter with more recent methods for a number of reasons.
+    This time lag is shorter with more recent methods for a number of reasons.
     First, historically theory and software papers were generally separated, whereas currently they are combined in the same papers making it easier to run simulations. 
-    Second, simulations testing methods are now required by journals (although simulations only show that the method behaves appropriately under ideal conditions), and finally, there are now far more people in the field and thus more papers published annually.
-    However, even with this reduced time-lag, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
+    Second, simulations testing methods are now required by journals (although simulations usually only show that the method behaves appropriately under ideal conditions), and finally, there are now far more people in the field and thus more papers published annually.
+    However, even with this reduced time lag, we suspect there are still many hidden assumptions and biases in all PCMs, even established methods, that have yet to be properly explored in the literature.
     For example, see Maddison \& FitzJohn's (\citeyear{maddison2014unsolved}) recent critique of Pagel's (\citeyear{pagel1994detecting}) correlated evolution method, and Rabosky \& Goldberg's (\citeyear{rabosky2015model}) discussion of trait-dependent speciation models \citep{nee1994reconstructed,maddison2007estimating}.
@@
     Although some information is not found in the literature (see above), the majority of assumptions and biases of PCMs are documented somewhere. 
-    A big issue for novice methods-users (and often for advanced users too) is that this information can be extremely technical and dense.
+    A big issue for novice methods users (and often for advanced users too) is that this information can be extremely technical and dense.
     It is not unusual for papers to be long and full of equations. 
@@
 
-    Another issue is that end-users need to understand the assumptions and caveats of methods.
+    Another issue is that end users need to understand the assumptions and caveats of methods.
     Within many comparative methods papers, assumptions and caveats can be found in the Introduction, Methods, Results and/or Discussion of a paper - they are rarely neatly corralled in one place.
@@
     The combined effects of a vast literature and sometimes opaque assumptions make it easy to miss pertinent details in PCM papers.
-    These problems relate to due diligence for both end-users and developers. 
-    Methods developers are not responsible for making sure that end-users read the literature. 
-    Instead the onus is on the end-user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
+    These problems relate to due diligence for both end users and developers. 
+    Methods developers are not responsible for making sure that end users read the literature. 
+    Instead the onus is on the end user to ensure that they have a clear understanding of the methods and caveats prior to using them. 
     However, simple steps could be taken by methods developers, such as subheadings that point to caveats and assumptions, to add clarity and limit method misuse and misinterpretation.
@@
 
-    More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' http://ftp.iitm.ac.in/cran/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous.
+    More recently however, the community has moved towards mostly implementing methods in R (\citealp{R-Core-Team:2014aa}; for a list of packages see Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' https://cran.r-project.org/web/views/Phylogenetics.html), and code sharing has become almost ubiquitous.
     The number of R packages for PCMs has increased markedly since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
@@
     Unfortunately the problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
-    The APE book \citep[an excellent resource;][]{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see \texttt{CAIC} and \texttt{caper} documentation; \citealp{purvis1995comparative,Orme:2013aa}).
+    The APE book \citep[an excellent resource;][]{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well established and very easily implemented in R (for a counter example see \texttt{CAIC} and \texttt{caper} documentation; \citealp{purvis1995comparative,Orme:2013aa}).
 
@@
     This is not entirely the fault of those writing these papers; most journals have strict word limits and a one line equation is generally a more succinct and precise descriptor of a method than a paragraph of text.
-    As we note above, the onus is on the end-user to read extensively and to do due diligence.
+    As we note above, the onus is on the end user to read extensively and to do due diligence.
     Nonetheless, due diligence is a responsibility at all levels of research and that includes ensuring clarity to all target users, especially in describing assumptions and limitations.\\
@@
     Some current implementations of the OU model should not be used with non-ultrametric trees (e.g.
-MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance co-variance matrix.
+MOTMOT; \citealp{Thomas:2011aa}) because they are based on transforming the tree directly, rather than transforming the variance covariance matrix.
     The problem is that where there is a pair of tips and at least one tip does not survive to the present, the expected covariances relating each of those two tips with any other tip in the tree are not identical.
@@
     In some cases, editors and reviewers may suggest using PCMs where they are not appropriate, and users should feel confident in rejecting these suggestions.
-    Finally, end-users should never be afraid to question standard practice, sometimes it is just PCM folklore.
+    Finally, end users should never be afraid to question standard practice, sometimes it is just PCM folklore.
 
   \subsubsection{3. Solutions and incentives}
-    It is important to recognise that our ability to do rigorous quantitative science often relies on highly-skilled methods-developers, especially as evolutionary biology becomes ever more computationally intensive.
+    It is important to recognise that our ability to do rigorous quantitative science often relies on highly-skilled methods developers, especially as evolutionary biology becomes ever more computationally intensive.
     We cannot afford to lose these people to industry, nor can we afford to pay industry wages, thus we need to make it worthwhile for such skilled researchers to remain in (or at least interact with) academia.
@@
 
-    Finally, an obvious solution to many of these problems is for methods-developers and end-users to collaborate more. 
+    Finally, an obvious solution to many of these problems is for methods developers and end users to collaborate more. 
     Both parties can benefit greatly by collaboration. 
     Some benefits are obvious, e.g.
-methods-developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it gets released, whereas end-users can work with the most cutting-edge methods and software.
+methods developers can gain extra datasets to test their ideas on and people who will discover corner cases and bugs in their software before it gets released, whereas end users can work with the most cutting edge methods and software.
     Most methods are not designed in a vacuum, they have a specific purpose usually based on predictions from theory, experimentation or observation.

### 2015-12-23 Manuscript/dark.tex
Final edits prepublication

--- 2015-12-09 Manuscript/dark.tex
@@
       \includegraphics[width=12cm]{Figures/PCMCitations.pdf}
-      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).}
+      \caption{The number of papers containing the phrase ``phylogenetic comparative'' published each year from 1980-2014 (Google Scholar search April 13th 2015).
+Data are available from figshare \citep{coopermeedata}.
+}
       \label{PCMCitations}
@@
       \caption{The cumulative total number of R packages for phylogenetics and phylogenetic comparative methods through time from 1980-2014.
-Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.}
+Source: Brian O'Meara's ``CRAN Task View: Phylogenetics, Especially Comparative Methods'' version 2015-01-21.
+Data are available from figshare \citep{coopermeedata}.
+}
       \label{PCMRpackages}
